### 2016-11-08 CreateData/showData4.Rmd
cosmetics

---
title: "showData 4"
author: "Göran Broström"
date: "November 5, 2016"
output: pdf_document
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
options(stringsAsFactors = FALSE)
```

# Introduction

Now *showData4.Rmd* is updated to include *fob90* and also let the
follow-up for *fob85* be *five years* and the follow-up for *fob90* be *ten
years*. We also add, as an update to *showData3*, guys aged 20 or more to the 
*FoB data*. Note that results (data files) are saved in **data4**!

This is an updated version of *showData2.Rmd*, where we use the newly-found
knowledge that the yearly *lisa* data relates to *December 31* the given
year, while the *FoB* data relate to the date *November 1* the given
year. These facts have the following implications for our plan: For *FoB*
data, we follow individuals from their appearance in a FoB until the next
FoB five years later, even if they do not appear there. The exception is if
they die within five years after the first appearance. So, *fob60* induces
a time interval *from November 1, 1960 to October 31, 1965*, and so on. To
include all 40+ in this interval, we must select all that are *born before
November 1, 1925*, that is, those who have reached the age 35 at November
1, 1960. Since we do not have birth dates but only birth year and quarter,
we decide to use as birthdate the *midpoint of the birth quarter*.

The following FoBs are treated similarly, until we reach the Fob
in 1985. Since the first *lisa* date is December 31, 1986, this "fob
interval" starts at November 1, 1985, and ends at December 31, 1986, that
is, it is 14 months long.

# Fixed data

This time we start with the *inddata* file and treat it as a *fixed-data*
file (no time-varying info). 
We find the birth dates (see above), death dates (for those who are observed die), and the "date last seen".
We also remove individuals that were born 1975 or later because they never get the chance to reach forty before *December 31, 2014*, which is our stop date.

Note that 37 individuals in *inddata* have no information at all, so they are removed.

```{r frominddata}
load("~/Forskning/Data/inddata.rda")
load("~/Forskning/Data/sex.rda")
indx <- match(inddata$LINNEID, sex$LINNEID)
inddata$KON <- sex$KON[indx]
per <- inddata[!is.na(inddata$KON) & (inddata$FODELSEAR < 1975), ]
per$birthdate <- per$FODELSEAR + 0.125  +  (per$FODELSEKVARTAL - 1) / 4
```

Now we put on *death date*, *death cause*, and *KOMMUN of death*, from *socstyr*, on *per*:

```{r styronper}
load("~/Forskning/Data/socstyr.rda")
indx <- match(per$LINNEID, socstyr$LINNEID)
per$DODSDAT <- socstyr$DODSDAT[indx]
fixd <- !is.na(per$DODSDAT) & (per$DODSDAT < 1960010)
per$DODSDAT[fixd] <- per$DODSDAT[fixd] * 100 + 15
per$deathdate <- as.numeric(skum::tillTid(per$DODSDAT))
per$dodsKOMMUN <- socstyr$KOMMUN[indx]
per$dodsALDER <- socstyr$ALDER[indx]
per$ULORSAK <- socstyr$ULORSAK[indx]
save(per, file = "data4/per.rda")
```

# The FoBs

Through the *FoB*s we have snapshots of the population on *November 1* the years *1960(5)1990*, giving the dates (decimal form).

```{r days}
ar <- seq(1960, 1990, by = 5)
days <- as.numeric(skum::tillTid(paste(ar, "1101", sep = "")))
days
```

We have to limit the data collection to the *KOMMUN*er that are contained
in our area, that is

```{r kommuner}
ourkom <- c(2406, 2414, 2417, 2418, 2480, 2482)
ourkom <- c(ourkom, 2411, 2412, 2413, 2416, 2415, 2461, 2407, 2402, 2408)
```

Note that this is per year.

## FoB 1960

Persons in the fob and in our KOMMUNer.

```{r fob60}
load("~/Forskning/Data/fob.rda")
fob60 <- fob[fob$AR == 1960 & 
                 fob$KOMMUN %in% ourkom &
                 fob$LINNEID %in% per$LINNEID, ]
```

They need to be of the right age (born on November 1, 1940 or earlier). We
get the birthdates from *inddata*.

```{r mat60}
library(eha)
indx <- match(fob60$LINNEID, per$LINNEID)
fob60$birthdate <- per$birthdate[indx]
lim <- as.numeric(toTime("1940-11-01"))
fob60 <- fob60[fob60$birthdate < lim, ]
indx <- match(fob60$LINNEID, per$LINNEID) # NOTE: Necessary!
fob60$enter <- lim + 20 - fob60$birthdate
dd <- per$deathdate[indx]
fob60$event <- !(is.na(dd)) & (dd <= lim + 25) # NOTE: 40 --> 25!
fob60$exit <- fob60$enter + 5
fob60$exit[fob60$event] <- dd[fob60$event] - fob60$birthdate[fob60$event]
head(fob60)
```

## FoB 1965

Persons in the fob and in our KOMMUNer.

```{r fob65}
fob65 <- fob[fob$AR == 1965 & 
                 fob$KOMMUN %in% ourkom &
                 fob$LINNEID %in% per$LINNEID, ]
```

They need to be of the right age (born on November 1, 1945 or earlier). We
get the birthdates from *inddata*.

```{r mat65}
library(eha)
indx <- match(fob65$LINNEID, per$LINNEID)
fob65$birthdate <- per$birthdate[indx]
lim <- as.numeric(toTime("1945-11-01"))
fob65 <- fob65[fob65$birthdate < lim, ]
indx <- match(fob65$LINNEID, per$LINNEID) # NOTE: Necessary!
fob65$enter <- lim + 20 - fob65$birthdate
dd <- per$deathdate[indx]
fob65$event <- !(is.na(dd)) & (dd <= lim + 25) ## NOTE: 40 --> 25!
fob65$exit <- fob65$enter + 5
fob65$exit[fob65$event] <- dd[fob65$event] - fob65$birthdate[fob65$event]
fob65 <- fob65[fob65$exit > fob65$enter, ] # Shouldn't be necessary!
head(fob65)
```

It should be noted that three persons present in the 1965 FoB actually died 1964 or earlier.

## FoB 1970

Persons in the fob and in our KOMMUNer.

```{r fob70}
fob70 <- fob[fob$AR == 1970 & 
                 fob$KOMMUN %in% ourkom &
                 fob$LINNEID %in% per$LINNEID, ]
```

They need to be of the right age (born on November 1, 1950 or earlier). We
get the birthdates from *inddata*.

```{r mat70}
library(eha)
indx <- match(fob70$LINNEID, per$LINNEID)
fob70$birthdate <- per$birthdate[indx]
lim <- as.numeric(toTime("1950-11-01"))
fob70 <- fob70[fob70$birthdate <= lim, ]
indx <- match(fob70$LINNEID, per$LINNEID) # NOTE: Necessary!
fob70$enter <- lim + 20 - fob70$birthdate
dd <- per$deathdate[indx]
fob70$event <- !(is.na(dd)) & (dd <= lim + 25)
fob70$exit <- fob70$enter + 5
fob70$exit[fob70$event] <- dd[fob70$event] - fob70$birthdate[fob70$event]
fob70 <- fob70[fob70$exit > fob70$enter, ] # Shouldn't be necessary!
head(fob70)
cat("lim = ", lim, "\n")
```

It should be noted that one person present in the 1970 FoB actually died 1964.

## FoB 1975

Persons in the fob and in our KOMMUNer.

```{r fob75}
fob75 <- fob[fob$AR == 1975 & 
                 fob$KOMMUN %in% ourkom &
                 fob$LINNEID %in% per$LINNEID, ]
```

They need to be of the right age (born on November 1, 1955 or earlier). We
get the birthdates from *inddata*.

```{r mat75}
library(eha)
indx <- match(fob75$LINNEID, per$LINNEID)
fob75$birthdate <- per$birthdate[indx]
lim <- as.numeric(toTime("1955-11-01"))
fob75 <- fob75[fob75$birthdate < lim, ]
indx <- match(fob75$LINNEID, per$LINNEID) # NOTE: Necessary!
fob75$enter <- lim + 20 - fob75$birthdate
dd <- per$deathdate[indx]
fob75$event <- !(is.na(dd)) & (dd <= lim + 25)
fob75$exit <- fob75$enter + 5
fob75$exit[fob75$event] <- dd[fob75$event] - fob75$birthdate[fob75$event]
fob75 <- fob75[fob75$exit > fob75$enter, ] # Shouldn't be necessary!
head(fob75)
##cat("lim = ", lim, "\n")
```

It should be noted that one person present in the 1975 FoB actually died 1964 (same person as earlier).

## FoB 1980

Persons in the fob and in our KOMMUNer.

```{r fob80}
fob80 <- fob[fob$AR == 1980 & 
                 fob$KOMMUN %in% ourkom &
                 fob$LINNEID %in% per$LINNEID, ]
```

They need to be of the right age (born on November 1, 1960 or earlier). We
get the birthdates from *inddata*.

```{r mat80}
library(eha)
indx <- match(fob80$LINNEID, per$LINNEID)
fob80$birthdate <- per$birthdate[indx]
lim <- as.numeric(toTime("1960-11-01"))
fob80 <- fob80[fob80$birthdate < lim, ]
indx <- match(fob80$LINNEID, per$LINNEID) # NOTE: Necessary!
fob80$enter <- lim + 20 - fob80$birthdate
dd <- per$deathdate[indx]
fob80$event <- !(is.na(dd)) & (dd <= lim + 25)
fob80$exit <- fob80$enter + 5
fob80$exit[fob80$event] <- dd[fob80$event] - fob80$birthdate[fob80$event]
fob80 <- fob80[fob80$exit > fob80$enter, ] # Shouldn't be necessary!
head(fob80)
##cat("lim = ", lim, "\n")
```

It should be noted that one person present in the 1980 FoB actually died 1964 (same person as earlier).

## FoB 1985

This is no longer the last FoB we are using: The follow-up time this time
*is* five years (as before), but five years, from November 1, 1985 to October 31, 1990, when the *lisa* data fills in.

Persons in the fob and in our KOMMUNer.

```{r fob85}
fob85 <- fob[fob$AR == 1985 & 
                 fob$KOMMUN %in% ourkom &
                 fob$LINNEID %in% per$LINNEID, ]
```

They need to be of the right age (born on November 1, 1965 or earlier). We
get the birthdates from *inddata*.

```{r mat85}
library(eha)
indx <- match(fob85$LINNEID, per$LINNEID)
fob85$birthdate <- per$birthdate[indx]
lim <- as.numeric(toTime("1965-11-01"))
fob85 <- fob85[fob85$birthdate < lim, ]
indx <- match(fob85$LINNEID, per$LINNEID) # NOTE: Necessary!
fob85$enter <- lim + 20 - fob85$birthdate
##fob85$enter <- as.numeric(toTime("1985-11-01")) - fob85$birthdate
dd <- per$deathdate[indx]
fob85$event <- !(is.na(dd)) & (dd <= lim + 25)
##fob85$event <- !(is.na(dd)) & (dd < 1987) 
fob85$exit <- fob85$enter + 5  ## Not any longer: 1 year and 2 months.
fob85$exit[fob85$event] <- dd[fob85$event] - fob85$birthdate[fob85$event]
fob85 <- fob85[fob85$exit > fob85$enter, ] # Shouldn't be necessary!
head(fob85)
##cat("lim = ", lim, "\n")
```

It should be noted that one person present in the 1985 FoB actually died 1964 (same person as earlier).

## FoB 1990

This is no longer the last FoB we are using: The follow-up time this time
*is* ten years (as before), from November 1, 1990 to December 31, 2001, when
the *lisa* data is filling in.

**Update 2016-11-05:** Draw it to December 31, 2001, i.e. add 14 months! Reason: 
No hisclass in LISA before 1 Jan 2002!

Persons in the fob and in our KOMMUNer.

```{r fob90}
fob90 <- fob[fob$AR == 1990 & 
                 fob$KOMMUN %in% ourkom &
                 fob$LINNEID %in% per$LINNEID, ]
```

They need to be of the right age (born on November 1, 1970 or earlier). We
get the birthdates from *inddata*.

```{r mat90}
library(eha)
indx <- match(fob90$LINNEID, per$LINNEID)
fob90$birthdate <- per$birthdate[indx]
lim <- as.numeric(toTime("1970-11-01"))
fob90 <- fob90[fob90$birthdate < lim, ]
indx <- match(fob90$LINNEID, per$LINNEID) # NOTE: Necessary!
fob90$enter <- lim + 20 - fob90$birthdate
##fob90$enter <- as.numeric(toTime("1985-11-01")) - fob85$birthdate
dd <- per$deathdate[indx]
fob90$event <- !(is.na(dd)) & (dd <= lim + 31 + 61 / 365)
##fob90$event <- !(is.na(dd)) & (dd < 1987) 
fob90$exit <- fob90$enter + 10 + 1 + 61 / 365 ## 10 + 1 + nov, dec: 1 year and 2 months.
fob90$exit[fob90$event] <- dd[fob90$event] - fob90$birthdate[fob90$event]
fob90 <- fob90[fob90$exit > fob90$enter, ] # Shouldn't be necessary!
head(fob90)
##cat("lim = ", lim, "\n")
```

It should be noted that one person present in the 1985 FoB actually died 1964 (same person as earlier).

This ends the data collection from the FoBs. Remains to put all together.

## Putting the data from all FoBs in one data frame

We put the event histories described by the FoBs (with FoB90!) into one data frame.

```{r fobgrab}
fobdata <- rbind(fob60, fob65, fob70, fob75, fob80, fob85, fob90)
fobdata <- fobdata[order(fobdata$LINNEID, fobdata$enter), ]
##fobdata <- age.window(fobdata, c(40, 110))
rownames(fobdata) <- 1:NROW(fobdata)
save(fobdata, file = "data4/fobdata.rda")
```

# The Lisa data

The *lisa* data is slightly simpler because it is yearly with cutoff dates 12-31/01-01 each new-year. It starts off at 1987-01-01 (lisa1986) and ends at 2013-12-31 (lisa2013). So the time interval a *lisa* covers is *AR + 1*! But note that the *socstyr* data only cover deaths *before January 1, 2014, so *lisa2013* is in effect useless to us.

```{r lisadat}
load("~/Forskning/Data/lisa.rda")
lisa <- lisa[lisa$KOMMUN %in% ourkom, ]
indx <- match(lisa$LINNEID, per$LINNEID)
lisa$birthdate <- per$birthdate[indx]
```

Those with *missing birthdate* in *lisa* are those born 1975 or later. They can be removed.

```{r removeyoung}
lisa <- lisa[!is.na(lisa$birthdate), ]
```

So, in *lisa*, all enter in *AR + 1*, at the exact age of *(AR + 1 - birthdate)*, and exit at the exact age *enter + 1* if they do not die in *the year AR + 1*, otherwise at their death date, or at *deathdate - birthdate*.

```{r fixlisaintervals}
lisa$enter <- lisa$AR + 1 - lisa$birthdate
indx <- match(lisa$LINNEID, per$LINNEID) # To be safe ...
lisa$deathdate <- per$deathdate[indx]
lisa$event <- !is.na(lisa$deathdate) & lisa$deathdate > lisa$AR + 1 & lisa$deathdate <= lisa$AR + 2
lisa$exit <- lisa$enter + 1 # First tip.
lisa$exit[lisa$event] <- with(lisa, deathdate[event] - birthdate[event])
```

Now some sorting  and cleaning.

```{r cleanandsort}
##lisadata <- age.window(lisa, c(40, 110))
lisadata <- lisa[order(lisa$LINNEID, lisa$enter), ]
lisadata <- cal.window(lisadata, c(1987, 2014)) # 'Remove lisa2013'
rownames(lisadata) <- 1:NROW(lisadata)
save(lisadata, file = "data4/lisadata.rda")
```

Now, read *checkData* (maybe not).