### 2016-11-14 slides.Rnw
first commit

\documentclass[t,a4paper,11pt]{beamer}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{natbib}

%\usepackage[figuresonly]{endfloat}

%\bibliographystyle{apalike}

\usepackage{color}

\newcommand{\emp}[1]{\textcolor{blue}{#1}}

\usetheme{Singapore}% best choice.
  \setbeamercovered{transparent}%

\newcommand{\R}{{\bf R}}
\newcommand{\code}[1]{{\tt #1}}

\title{Social inequality in mortality among adults and elderly in 
  northern  Sweden 1851--2013\footnote{'Socoeconomic inequalities and mortality I: Europe'. 41st Annual Meeting of the Social Science History Association, 
    Chicago,
    November 17--20, 2016.}
}

\date{November 19, 2016}

\author{Göran Broström and Sören Edvinsson}

\begin{document}

\maketitle

<<setup, include = FALSE>>=
library(knitr)
opts_chunk$set(engine='R',dev='pdf',fig.path='figs/slide-',tidy=FALSE,fig.height=5.5,fig.width=8,strip.white=TRUE,cache=TRUE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
library(knitr)
library(eha)
library(xtable)
library(skum)
@

\begin{frame}{Aims and questions}

\begin{enumerate}
\item Has \emp{inequality in mortality} between \emp{social classes}
  increased in the adult and elderly 
  population?

%%\item Is the mortality transition initiated by a certain social class?

\item Are there any \emp{gender differences} in the effect of social position?

\item Is social position \emp{equally important} among the \emp{retired} population as
  for those in \emp{working age}?
\end{enumerate}

\end{frame}

\begin{frame}{The Skellefteå and Umeå regions in Sweden}

<<umpop>>=
ume1900 <- with(obs[obs$region == "ume", ],
                sum(birthdate + enter < 1900 &
                    birthdate + exit > 1900))
ume1950 <- with(obs[obs$region == "ume", ],
                sum(birthdate + enter < 1950 &
                    birthdate + exit > 1950))
@

\includegraphics[height=8cm]{myfigs/skum17}\includegraphics[height=8cm]{myfigs/sweden}

\end{frame}

\begin{frame}{The Skellefteå region}

\begin{itemize}  
\item Selection of
parishes surrounding the town of Skellefteå. 
\item Town founded in 1843 but with a
very small population during the first decades.
\item The majority of the
population in rural villages and hamlets, getting their
livelihood from agricultural production. 
\item 20th century: \emp{industrialisation}, population \emp{increase in the town},
\emp{diversified economy}. 
\end{itemize}

\end{frame}

\begin{frame}{The Umeå region}

\end{frame}

\begin{frame}[fragile]{Data sources}

\begin{itemize}
\item \emp{Swedish church book} records
\item digitized by the \emp{Demographic Data Base} at Umeå University, 
\begin{itemize}  
   \item \emp{\sc Popum} and 
   \item \emp{\sc Poplink}
\end{itemize}    
\item  \url{http://www.cedar.umu.se/}
\end{itemize}

\end{frame}

\begin{frame}{Software for analysis}

\begin{itemize}
\item  \R : \emp{\url{https://www.R-project.org}}.
  \begin{itemize}
    \item \emp{RStudio}: \url{https://www.rstudio.com}.
  \end{itemize}    
\item \R\ \emp{packages} (\emp{\url{https://cran.R-project.org}}):
  \begin{itemize}
  \item \emp{skum} (Broström, 2016) data from \emp{DDB}.
  \item \emp{rmarkdown} (RStudio Team, 2016).
  \item \emp{knitr} (Xie, 2016).
  \item \emp{eha} (Broström, 2015).
  \item \emp{xtable} (Dahl, 2016).
\end{itemize}            
\end{itemize}

\end{frame}

\begin{frame}{Variables}

\begin{description}

\item[what] is this?

\end{description}

\end{frame}

\begin{frame}{\sc Hisclass}

\begin{enumerate}
\item \'Elite, {\sc Hisclass} \emp{1 and 2}: Higher managers and higher professionals.
\item Lower managers, {\sc Hisclass} \emp{3}.
\item Farmers and fishermen, {\sc Hisclass} \emp{8}.
\item Lower professionals, {\sc Hisclass} \emp{4, 5 and 6}: Lower professionals, lower clerical, foremen.
\item Medium skilled workers, {\sc Hisclass} \emp{7}.
\item Lower skilled workers, {\sc Hisclass} \emp{9, 10, 11 and 12}: Lower and unskilled workers (including farm workers)
%%\item Lower skilled farm workers, {\sc Hisclass} 10 and 12, Lower and unskilled farm workers
\end{enumerate}

\end{frame}

\begin{frame}{{\sc Poplink} data, 1901--1950}

<<getdata, results='asis'>>=
library(xtable)
library(knitr)
load("data/adult.rda")
adult <- cal.window(adult, c(1901, 1951))
source("R/addPer.R")
adult <- addPer(adult, cuts = seq(1901, 1951, by = 10))
adult <- rc(adult)
## NEW hisclass, 6 + 7 --> 6:
adult$hisclass[!is.na(adult$hisclass) & (adult$hisclass == 7)] <- 6
@

\end{frame}

\begin{frame}{{\sc Fob} data, 1961--2000}

<<ssykonfob>>=
load("fobb.rda")
@

\end{frame}

\begin{frame}{{\sc Lisa} data, 2002--2013}

<<lisadata>>=
load("liss.rda")
source("R/addPer.R")
liss$id <- liss$LINNEID
liss <- addPer(liss, name = "period", cuts = seq(2002, 2014, by = 6))
liss <- liss[!is.na(liss$hisclass), ]
liss$hisclass <- cut(liss$hisclass, c(0, 2.5, 4.5, 5.5, 8), 
                     labels = c("1+2", "3+4", "5", "6"))
liss <- liss[with(liss, !is.na(civst) &
                      !is.na(hisclass)), ]
liss$civst <- factor(liss$civst, 
                     labels = c("unmarried", "married", "prev. married"))
liss$urban <- liss$FORSAMLING %in% c(248001, 248002, 284004, 284005, 
                                     248011, 248201, 248204)
@

\end{frame}

\begin{frame}{Results, ages 40--59}

<<colorc>>=
## Colors for plotting strata:
scol <- c(rep("blue", 2), rep("green", 2), rep("red", 2))
## Common for all groups:
labb1 <- levels(adult$period)
labb2 <- levels(fobb$period)
labb <- c(labb1, labb2)
nper1 <- length(labb1)
nper2 <- length(labb2)
nper <- nper1 + nper2
fobb$urban <- fobb$FORSAMLING %in% c(248001, 248002, 284004, 284005, 248011, 248201, 248204)
civst <- numeric(NROW(fobb))
civst[fobb$CIVIL %in% c(0, 5)] <- 1 # Ogift
civst[fobb$CIVIL %in% c(2, 7)] <- 2 # Gift, sambo
civst[fobb$CIVIL %in% c(1, 3, 4, 6, 8, 9)] <- 3 # skild, anka
fobb$civst <- factor(civst, labels = c("unmarried", "married", "prev. married"))
source("R/getPH.R")
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--59, women."
scap1w <-  "Cumulative hazards, ages 40--59, women."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--59, men."
scap1m <-  "Cumulative hazards, ages 40--59, men."
##
cap12 <- "Total hazard of dying before age 60 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom)."
scap12 <- "Total hazard by HISCLASS, ages 40--59."    
cap12log <- "Total hazard of dying before age 60 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). Log scale."
scap12log <- "Total hazard by HISCLASS, ages 40--59 (log scale)."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 60--89, women."
scap2w <-  "Cumulative hazards, ages 60--89, women."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 60--89, men."
scap2m <-  "Cumulative hazards, ages 60--89, men."
##
cap22 <- "Total hazard of dying before age 90 for a 60 year old person by HISCLASS and decade, women (top) and men (bottom)."
scap22 <- "Total hazard by HISCLASS, ages 60--89."

##
adult4059 <- age.window(adult, c(40, 60))
fobb4059 <- age.window(fobb, c(40, 60))
## End 'common'
@

\end{frame}

\begin{frame}{Ages 40--59, women}

<<hazcalw>>=
resw <- getPH(adult4059[adult4059$sex == "female", ],
                fobb4059[fobb4059$KON == "kvinna", ],
                nper1, nper2, labb)
resm <- getPH(adult4059[adult4059$sex == "male", ],
                fobb4059[fobb4059$KON == "man", ],
                nper1, nper2, labb)
@

<<hazplotw, fig.height = 6>>=
source("R/plotHaz.ph.R")
plotHaz.ph(resw, main = labb, printLegend = FALSE)
@

\end{frame}

\begin{frame}{Ages 40--59, men}

<<hazplotm, fig.height = 6>>=
plotHaz.ph(resm, main = labb, printLegend = FALSE)
@

\end{frame}

\begin{frame}{Total hazards in ages 40--59, women}

<<cumsw, fig.height = 5.5>>=
source("R/plotED.ph.R")
woutp <- plotED.ph(resw, scol = scol, logScale = FALSE, ylab = "H(60|40)")
@

\end{frame}

\begin{frame}{Total hazards in ages 40--59, men}

<<cumsm, fig.height = 5.5>>=
moutp <- plotED.ph(resm, scol = scol, logScale = FALSE, ylab = "H(60|40)")
@

\end{frame}

\begin{frame}{Total hazards in ages 40--59, women: Log scale}

<<cumslogw>>=
wloutp <- plotED.ph(resw, scol = scol, logScale = TRUE, ylab = "H(60|40)",
main = "Women")
@

\end{frame}

\begin{frame}{Total hazards in ages 40--59, men: Log scale}
<<cumslogm>>=
mloutp <- plotED.ph(resm, scol = scol, logScale = TRUE, ylab = "H(60|40)",
main = "Men")
@

\end{frame}

\begin{frame}{{\sc Lisa} data, 2002--2013, women 40--59}

<<lissyoung>>=
liss4059 <- age.window(liss, c(40, 59))
fitw1 <- phreg(Surv(enter-40, exit-40, event) ~ strata(hisclass) + 
                   civst + urban, dist = "gompertz", 
               data = liss4059[liss4059$KON == "kvinna", ])
fitm1 <- phreg(Surv(enter-40, exit-40, event) ~ strata(hisclass) + 
                   civst + urban, dist = "gompertz", 
               data = liss4059[liss4059$KON == "man", ])
dcol <- c("blue", "darkgreen", "darkgreen", "red")
@

<<plotliss1>>=
plot(fitw1, fn = "cum", col = dcol, main = "Women", lwd = rep(2, 4))
@

\end{frame}

\begin{frame}{{\sc Lisa} data, 2002--2013, men 40--59}

<<plotliss1m>>=  
plot(fitm1, fn = "cum", col = dcol, main = "Men", lwd = rep(2, 4))
@

\end{frame}

\begin{frame}{Cox regression, women 2002--2013, ages 40--59}

\scriptsize

<<tablisswom1, results = "asis">>=
fitwo1 <- coxreg(Surv(enter, exit, event) ~ hisclass + civst + urban + period, 
                 data = liss4059[liss4059$KON == "kvinna", ])
dr <- drop1(fitwo1, test = "Chisq")
ltx(fitwo1, dr = dr)
@

\end{frame}

\begin{frame}{Cox regression, men 2002--2013, ages 40--59}

\scriptsize

<<tablissman1, results = "asis">>=
fitma1 <- coxreg(Surv(enter, exit, event) ~ hisclass + civst + urban + period, 
                 data = liss4059[liss4059$KON == "man", ])
dr <- drop1(fitma1, test = "Chisq")
ltx(fitma1, dr = dr)
@

\end{frame}

\begin{frame}{Results, ages 60--89}

\end{frame}

\begin{frame}{Ages 60--89, women}

<<getske6089>>=
adult6089 <- age.window(adult, c(60, 90))
fobb6089 <- age.window(fobb, c(60, 90))
resw <- getPH(adult6089[adult6089$sex == "female", ],
                fobb6089[fobb6089$KON == "kvinna", ],
                nper1, nper2, labb, from = 60)
resm <- getPH(adult6089[adult6089$sex == "male", ],
                fobb6089[fobb6089$KON == "man", ],
                nper1, nper2, labb, from = 60)
plotHaz.ph(resw, main = labb)

@

\end{frame}

\begin{frame}{Ages 60--89, men}

<<hazplotmold>>=
plotHaz.ph(resm, main = labb)
@

\end{frame}

\begin{frame}{Total hazards in ages 60--89, women}

<<cumswold>>=
woutp <- plotED.ph(resw, scol = scol, logScale = FALSE, ylab = "H(90|60)")  
@

\end{frame}

\begin{frame}{Total hazards in ages 60--89, men}

<<cumsmold>>=
moutp <- plotED.ph(resm, scol = scol, logScale = FALSE, ylab = "H(90|60)")
@

\end{frame}

\begin{frame}{{\sc Lisa} data, 2002--2013, women 60--89}

<<lissold>>=
liss6089 <- age.window(liss, c(60, 90))
fitw2 <- phreg(Surv(enter-60, exit-60, event) ~ strata(hisclass) + 
                   civst + urban, dist = "gompertz", 
               data = liss6089[liss6089$KON == "kvinna", ])
fitm2 <- phreg(Surv(enter-60, exit-60, event) ~ strata(hisclass) + 
                   civst + urban, dist = "gompertz", 
               data = liss6089[liss6089$KON == "man", ])
dcol <- c("blue", "darkgreen", "red", "red")
plot(fitw2, fn = "cum", col = dcol)

@

\end{frame}

\begin{frame}{Cox regression, women 2002--2013, ages 60--89}

  \scriptsize
<<tablisswom2, results = "asis", eval = TRUE>>=
fitwo2 <- coxreg(Surv(enter, exit, event) ~ hisclass + civst + urban + period, 
                 data = liss6089[liss6089$KON == "kvinna", ])
dr <- drop1(fitwo2, test = "Chisq")
ltx(fitwo2, dr = dr)
@

\end{frame}

\begin{frame}{{\sc Lisa} data, 2002--2013, men 60--89}

<<lissoldmen>>=
plot(fitm2, fn = "cum", col = dcol)
@

\end{frame}

\begin{frame}{Cox regression, men 2002--2013, ages 60--89}

  \scriptsize
<<tablissman2, results = "asis", eval = TRUE>>=
fitma2 <- coxreg(Surv(enter, exit, event) ~ hisclass + civst + urban + period, 
                 data = liss6089[liss6089$KON == "man", ])
dr <- drop1(fitma2, test = "Chisq")
ltx(fitma2, dr = dr)
@

\end{frame}

\begin{frame}{Conclusion}

%\scriptsize

\begin{itemize}
\item \emp{Improved} survival characterised all {\sc Hisclass}es.
\item There are \emp{gender differences} in the social pattern
\item No difference in the social pattern in the two studied age groups.
\item The \emp{upper} class(es) gradually take over as the \emp{most healthy} group.
\end{itemize}

%We started this paper by raising the issue of the development of social
%inequalities. When studying this question in history we often get surprised
%of its complexity. In a society characterised by large economic inequality, 
%we would expect that this would result in corresponding mortality
%differences, i.e. higher social classes having an advantage. We find the
%appearance of a consistent social pattern in mortality, to the advantage of
%the higher classes, but surprisingly late in our history. It is only in the
%recent decades that this pattern dominates. Improved survival characterised
%however all {\sc Hisclass}es. We can however confirm that there are gender
% differences in the social pattern. While there is no clear social gradient
% among men, the results at least partly indicate advantages for the higher
% social classes for women already early in the 20th century. Females had
% more to gain from a high social position. Previous studies of the mortality
% transition have made similar observations. Among adults, gender has often
% been of more significant importance than social class. The reason for this
% difference and for the late appearance of the modern social pattern among
% men is still an enigma that require further studies, even though we have
% suggested the possible importance of behavioural factors and changes in
% ideals connected to different social classes. This should stimulate us to
% further disentangle the pathways for how social position influence health
% and mortality in different societies. Finally, we do not find any clear
% differences in the social pattern in the two studied age groups.

\end{frame}

\end{document}