### 2016-11-15 slides.Rnw
add stuff

\documentclass[t,a4paper,11pt]{beamer}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{natbib}

%\usepackage[figuresonly]{endfloat}

%\bibliographystyle{apalike}

\usepackage{color}

\newcommand{\emp}[1]{\textcolor{blue}{#1}}

\usetheme{Singapore}% best choice.
  \setbeamercovered{transparent}%

\newcommand{\R}{{\bf R}}
\newcommand{\code}[1]{{\tt #1}}

\title{Social inequality in mortality among adults and elderly in 
  northern  Sweden 1851--2013\footnote{'Socoeconomic inequalities and mortality I: Europe'. 41st Annual Meeting of the Social Science History Association, 
    Chicago,
    November 17--20, 2016.}
}

\date{November 19, 2016}

\author{Göran Broström and Sören Edvinsson}

\begin{document}

\maketitle

<<setup, include = FALSE>>=
library(knitr)
opts_chunk$set(engine='R',dev='pdf',fig.path='figs/slide-',tidy=FALSE,fig.height=5.5,fig.width=8,strip.white=TRUE,cache=TRUE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
library(knitr)
library(eha)
library(xtable)
library(skum)
@

\begin{frame}{Aims and questions}

\begin{enumerate}
\item Has \emp{inequality in mortality} between \emp{social classes}
  increased in the adult and elderly 
  population?

%%\item Is the mortality transition initiated by a certain social class?

\item Are there any \emp{gender differences} in the effect of social position?

\item Is social position \emp{equally important} among the \emp{retired} population as
  for those in \emp{working age}?
\item Is there a difference between classes in mortality from
  \emp{cardiovascular} diseases and \emp{cancers} respectively? 
\item Are there large differences in survival depending on level of
  \emp{education} and \emp{income} categories?

\end{enumerate}

\end{frame}

\begin{frame}{The Skellefteå and Umeå regions in Sweden}

<<umpop>>=
ume1900 <- with(obs[obs$region == "ume", ],
                sum(birthdate + enter < 1900 &
                    birthdate + exit > 1900))
ume1950 <- with(obs[obs$region == "ume", ],
                sum(birthdate + enter < 1950 &
                    birthdate + exit > 1950))
@

\includegraphics[height=8cm]{myfigs/skum17}\includegraphics[height=8cm]{myfigs/sweden}

\end{frame}

\begin{frame}{The Skellefteå region}

\begin{itemize}  
\item Selection of
parishes surrounding the town of Skellefteå. 
\item Town founded in 1843 but with a
very small population during the first decades.
\item The majority of the
population in rural villages and hamlets, getting their
livelihood from agricultural production. 
\item 20th century: \emp{industrialisation}, population \emp{increase in the town},
\emp{diversified economy}. 
\end{itemize}

\end{frame}

\begin{frame}{The Umeå region}

\begin{itemize}
\item \emp{Town} founded in \emp{1622}
\begin{itemize}
\item  Substantially \emp{larger} than Skellefteå
\item \emp{administrative} centre of the county Västerbotten
\item \emp{Military} regiments.
\item \emp{Schools}
\item \emp{University} from \emp{1963--65}.  
\end{itemize}  
\item The rural part similar to the one in the Skellefteå region.
\end{itemize}

\end{frame}

\begin{frame}[fragile]{Data sources}

\begin{itemize}
\item \emp{Swedish church book} records (1851--1950).
  \begin{itemize}
\item digitized by the \emp{Demographic Data Base} at Umeå University, 
\begin{itemize}  
   \item \emp{\sc Popum} and 
   \item \emp{\sc Poplink}
\end{itemize}    
\item  \url{http://www.cedar.umu.se/}
\end{itemize}
\item The \emp{Linnaeus} data base (1960--2013)
  \begin{itemize}
\item Census data 1960, 1965, \ldots, 1990
\item Population registers ({\sc Lisa}, 1986--2013)
  \begin{itemize}
    \item \emp{Income}
    \item \emp{Education}  
  \end{itemize}    
 \item Death information from \emp{National Board of Health and Welfare}
   \begin{itemize}
\item Death dates from 1 January 1961 to 31 December 2013.
  \item \emp{Causes of death} ({\sc Ulorsak}, ``main cause of death'')
   \end{itemize}
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}{The sampling frame}

<<lexis2>>=
plot(c(1851, 1851, 1951, 1951, 1851), c(40, 90, 90, 40, 40), 
     type = "l", xlim = c(1851, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 90))
axis(1, at = c(1851, 1901, 1951, 1961, 1981, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1961, 1961, 2014, 2014, 1961), c(40, 90, 90, 40, 40), 
      col = "blue", lwd = 2)
text(1900, 65, "PopLink", cex = 2, col = "red")
text(1900, 55, "79593 persons")
text(1900, 50, "32801 deaths")
text(1985, 65, "Linnaeus", cex = 2, col = "blue")
text(1985, 55, "156676 persons")
text(1985, 50, "42879 deaths")
@ 
\end{frame}

\begin{frame}{The Linnaeus data}

\begin{columns}
\begin{column}{0.6\textwidth}
<<picklexus,fig.width=6,fig.height = 5.9>>=

load("data/vb.rda")
vbd <- vb[vb$event == 1 & vb$dataset != "poplink", c("birthdate", "exit", "sex")]
x <- vbd$birthdate + vbd$exit
y <- vbd$exit
##par(mfrow = c(1, 2))
plot(x[vbd$sex == "male"], y[vbd$sex == "male"], type = "p", 
     pch = ".", col = "blue", xlab = "Year", ylab = "Age", las = 2,
     axes = FALSE)
axis(1, at = c(seq(1960, 1990, by = 5), 2002, 2007, 2014), las = 2)
axis(2, at = c(40, 65, 90))
axis(4, at = c(40, 65, 90))
for (i in seq(1960, 1990, by = 5) + (365 - 61) / 365){
    abline(v = i, lwd = 2)
}
for (i in 2002:2013){
    abline(v = i, lwd = 2, col = "black")
}
abline(h = c(40, 65, 90), lty = 2)
points(x[vbd$sex == "female"], y[vbd$sex == "female"], type = "p", 
     pch = ".", col = "red")
box()
@

\end{column}

\begin{column}{0.4\textwidth}
\scriptsize

\vspace{3\baselineskip}

\begin{tabular}{l|rr}
      & Number & Deaths \\ \hline
\textcolor{red}{Women} & 70\,249 & \textcolor{red}{13\,888} \\
\emp{Men} & 86\,427 & \emp{28\,991} \\ \hline
Total  &156\,676 & 42\,879
\end{tabular}

\vspace{\baselineskip}

\emp{\underline{Note}}: Persons with \emp{missing {\sc Hisclass}} (occupation) are \emp{excluded}.
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Software for analysis}

\begin{itemize}
\item  \R : \emp{\url{https://www.R-project.org}}.
  \begin{itemize}
    \item \emp{RStudio}: \url{https://www.rstudio.com}.
  \end{itemize}    
\item \R\ \emp{packages} (\emp{\url{https://cran.R-project.org}}):
  \begin{itemize}
  \item \emp{skum} (Broström, 2016) data from \emp{DDB}.
  \item \emp{rmarkdown} (RStudio Team, 2016).
  \item \emp{knitr} (Xie, 2016).
  \item \emp{eha} (Broström, 2016).
  \item \emp{xtable} (Dahl, 2016).
\end{itemize}            
\end{itemize}

\end{frame}

\begin{frame}{Variables}

\begin{description}

\item[sex]
\item[birth date]
\item[death date]
\item[cause of death]
\item[parish]  
\item[\sc Hisclass] grouped.
\item[income] 1990--2005, grouped by quartiles.
\item[education] 1990--2005. Seven categories.

\end{description}

\end{frame}

\begin{frame}{\sc Hisclass}

\begin{enumerate}
\item \emp{\tt elite}, HISCLASS 1 and 2. Higher managers and professionals.
\item \emp{\tt lowMan}, HISCLASS 3 and 8. Lower managers and farmers.
\item \emp{\tt lowWhiteC}, HISCLASS 4, 5 and 6. Lower white collar.
\item \emp{\tt worker}, HISCLASS 7, 9, 10, 11 and 12. Workers of different skills including farm workers.
\end{enumerate}

\end{frame}

\begin{frame}{Proportional hazards for women?}

\includegraphics[width=11cm,height=8cm]{figure/hazplotw-1}

\end{frame}

\begin{frame}{Proportional hazards for men?}

\includegraphics[width=11cm,height=8cm]{figure/hazplotm-1}

\end{frame}

\begin{frame}[fragile]{Interactions}

\begin{verbatim}
fit <- coxph(Surv(enter - 40, exit - 40, event) ~ 
    period * hisclass * sex + urban, data = vb)
drop1(fit, test = "Chisq")
\end{verbatim}

\begin{center}
\begin{tabular}{lrrrr}
  \hline
 & Df & AIC & LRT & Pr($>$Chi) \\ 
  \hline
$<$none$>$ &  & 1631901.32 &  &  \\ 
  urban & 1 & 1631946 & 47.04 & $7.0\times 10^{-12}$ \\ 
  \emp{period:hisclass:sex} & 27 & 1631957 & 110.11 & \emp{$5.3\times 10^{-12}$} \\ 
   \hline
\end{tabular}
\end{center}

\begin{itemize}

\item This model is described by \emp{80 coefficients}.

\item We show them in \emp{graphs}.
\end{itemize}

\end{frame}

\begin{frame}{No proportional hazards (wrt \sc Hisclass)!}

\begin{itemize}
\item Fit each {\sc Hisclass} \emp{separately}, for each time period.
\item Use the \emp{cumulative hazard} at 65 (or 90) for comparison.
\end{itemize}

So, we fit separate PH models for each combination of

\begin{itemize}
  \item \emp{\tt sex}
  \item \emp{\tt period}: 1851--75, 1876--1900, 1901--10, etc.
  \item \emp{\tt hisclass}
  \item \emp{\tt age group}: 40--64, 65--89.
  \item \emp{\tt cause of death}
\end{itemize}    
\end{frame}

\begin{frame}{The data table}

<<lexis3, fig.height = 4>>=
plot(c(1851, 1851, 1951, 1951, 1851), c(40, 90, 90, 40, 40), 
     type = "l", xlim = c(1851, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(40, 90))
axis(1, at = c(1851, 1901, 1951, 1961, 1981, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1961, 1961, 2014, 2014, 1961), c(40, 90, 90, 40, 40), 
      col = "blue", lwd = 2)
for (i in c(1851, 1876, 1901, 1926, 1971, 1981, 1991, 2002, 2007)){
    lines(c(i, i), c(40, 90), lty = 2)
}
for (i in c(45, 50, 55, 60, 65, 70, 75, 80, 85)){
    lines(c(1851, 1951), c(i, i), lty = 2)
    lines(c(1961, 2014), c(i, i), lty = 2)
}
@

\emp{Poisson regression}:

In \emp{each cell}, and for \emp{each} combination of \emp{covariate level}s, calculate 
\begin{itemize}
\item  the number of \emp{deaths} and 
\item the total \emp{exposure time}.
\end{itemize}
\end{frame}

\begin{frame}{Women and men, age 40--64}

  \begin{center}
\includegraphics[height = 7.5cm]{figure/cumsw-1}
  \end{center}
\end{frame}

\begin{frame}{Relative differences, ages 40--64 and 65--89}

  %\begin{center}
\includegraphics[height = 7.5cm]{figure/cumswstand-1}\includegraphics[height=7.5cm]{figure/cumswold-1}
   %\end{center}
\end{frame}

\begin{frame}{Cancer mortality, ages 40-64 and 65-89}

\includegraphics[height=7.5cm]{figure/12cumsw-1}\includegraphics[height=7.5cm]{figure/12cumswold-1}

\end{frame}

\begin{frame}{Cardiovascular mortality}

\includegraphics[height=7.5cm]{figure/41cumsw-1}\includegraphics[height=7.5cm]{figure/41cumswold-1}

\end{frame}

\begin{frame}{Education and income, 1990--2005, all deaths}

\includegraphics[height=3.5cm]{figure/liswomfig-1} Women, 40--64.

\includegraphics[height=3.5cm]{figure/lismanfig-1} Men, 40--64.

\end{frame}

\begin{frame}{Education and income, 1990--2005, all deaths}

\includegraphics[height=3.5cm]{figure/liswomfig2-1} Women, 65--89.

\includegraphics[height=3.5cm]{figure/lismanfig2-1} Men, 65--89.

\end{frame}

\begin{frame}{Education and income, 1990--2005, cancer}

\includegraphics[height=3.5cm]{figure/12liswomfig-1} Women, 40--64.

\includegraphics[height=3.5cm]{figure/12lismanfig-1} Men, 40--64.

\end{frame}

\begin{frame}{Education and income, 1990--2005, cancer}

\includegraphics[height=3.5cm]{figure/12liswomfig2-1} Women, 65--89.

\includegraphics[height=3.5cm]{figure/12lismanfig2-1} Men, 65--89.

\end{frame}

\begin{frame}{Education and income, 1990--2005, cardiovascular}

\includegraphics[height=3.5cm]{figure/41liswomfig-1} Women, 40--64.

\includegraphics[height=3.5cm]{figure/41lismanfig-1} Men, 40--64.

\end{frame}

\begin{frame}{Education and income, 1990--2005, cardiovascular}

\includegraphics[height=3.5cm]{figure/41liswomfig2-1} Women, 65--89.

\includegraphics[height=3.5cm]{figure/41lismanfig2-1} Men, 65--89.

\end{frame}

%%% Stop here!

\begin{frame}{Conclusion}

%\scriptsize

\begin{itemize}
\item \emp{Improved} survival characterised all {\sc Hisclass}es.
\item There are \emp{gender differences} in the social pattern
\item No difference in the social pattern in the two studied age groups.
\item The \emp{upper} class(es) gradually take over as the \emp{most healthy} group.
\end{itemize}

\end{frame}

\end{document}