### 2017-05-24 icass.Rnw
Updating text in introduction and more

\documentclass[a4paper,11pt,titlepage]{article}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{epstopdf}
\usepackage{url}

\usepackage{natbib}

\bibliographystyle{plainnat}

%\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\newcommand{\R}{{\bf R}}

%\usepackage[tablesonly]{endfloat}
%\usepackage[figuresonly]{endfloat}
\usepackage{endfloat}

\renewcommand{\efloatseparator}{\vfill}

% patch \efloat@iwrite to use \protected@write
\makeatletter
\renewcommand\efloat@iwrite[1]{%
   \immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother

\author{Sören Edvinsson\footnote{email: \texttt{soren.edvinsson@umu.se}} and
Göran Broström\footnote{email: \texttt{goran.brostrom@umu.se}}
     \\
Centre for Demographic and Ageing Research ({\sc Cedar})\\
Umeå University, SE-90187 Umeå, Sweden}

\title{Social inequality in mortality among
  adults and elderly in northern Sweden 1851--2013\footnote{Paper presented
    at the
    session 'Health and Mortality in Arctic Populations'. The Ninth International
    Congress on Arctic Social Sciences,
    Umeå,
    June 8--12, 2017}}

\date{May 24, 2017}

\begin{document}

\maketitle

<<setup, include = FALSE>>=
library(knitr)
opts_chunk$set(engine='R',dev='pdf',fig.path='figs/slide-',tidy=FALSE,fig.height=5.5,fig.width=8,strip.white=TRUE,cache=TRUE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
library(knitr)
library(eha)
library(xtable)
library(skum)
@

<<readalldata>>=
## Load remote data:
con <- url("http://capa.ddb.umu.se/data/vb0.rda")
##load("data/vb0.rda")
load(con)
close(con)
vb0$hisclass[is.na(vb0$hisclass)] <- 4
vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "middle", "worker", "NA"))
##
con <- url("http://capa.ddb.umu.se/data/vb.rda")
##load("data/vb.rda")
load(con)
close(con)
##
con <- url("http://capa.ddb.umu.se/data/lisaold.rda")
##load("data/lisaold.rda")
load(con)
close(con)
##
con <- url("http://capa.ddb.umu.se/data/loTab.rda")
load(con)
close(con)
loTab$period <- factor(loTab$period)
##
vb0Tab <- aggregate(vb0[, c("event", "exposure")],
                   vb0[, c("sex", "age", "period", "cohort",
                          "urban", "civst", "hisclass")],
                   FUN = sum)
vb0Tab$age <- factor(vb0Tab$age)
@

\begin{abstract}
A long-term perspective on the development of social
inequalities in mortality in the adult and elderly population is taken in
this paper.
The area under investigation consists of the Skellefteå and Umeå regions in
the north of Sweden, and the time period is 1851--2013. The outcome is all
cause mortality and cause specific mortality for the categories
cardiovascular diseases and cancers. The analysis of inequality is based on
Hisclass.

The main findings are that a consistent social gradient implying better
survival for classes with more access to economic and other resources, only
become evident in the later half of the 20th century, with lower classes
having the highest mortality. The class patterns in mortality differed
however depending on gender. Women of higher classes were comparatively
favoured in most of the studied periods, while higher class men were
disadvantaged when it comes to survival at least until the middle of the
20th century. The modern pattern of clear socio-economic differences
appeared surprisingly late. Additional analyses based on education and
income categories confirmed large inequality in mortality
between different groups during the most recent decades, but also demonstrated more distinct differences between categories with better survival the more income and the higher education level people have. Both cardiovascular diseases and cancers conform with the pattern already
observed when it comes to all-cause mortality.

\end{abstract}

\tableofcontents

\newpage

%%{\huge Här ska 'intro.tex' in !!!!!!!!!!!!!!!!!!!!!}

%%\subsection{Combining the data sets and tabulate}

\section{Introduction}

One of the most central aspects of survival on individual level is social class and access to economic and other resources. Even in present welfare societies, social position is a strong determinant when it comes to health and mortality and the impact seems to be increasing as well \citep{kuetal04, frlu07}. It has been suggested that “… social conditions have been, are and will continue to be irreducible determinants of health outcomes and therefore deserves their appellation as ‘fundamental causes’ of disease and death” \citep{liph95}. For a long time, the general view has been that socio-economic health inequalities were high in historical societies, probably higher than in modern societies. This can be considered as a reasonable assumption since these societies were in most cases characterised by very large differences between groups and classes. Knowing that access to resources provides advantages in all ways, the health advantage of higher classes ought to be obvious. \citet{ant67} and \citet{smi83} suggest that social inequality in health has passed through different historical phases. According to them, social health differences were comparatively small during the pre-transitional phase when space was a strong determinant for the spread of disease. Differences then increased during the transitional phase when mortality declined and wealthy groups used their resources to gain better health. Finally, health differences decreased again in modern low-mortality settings when instead health-related behaviour became important, resulting in marginal differences. \citet{omran82} state, in the third proposition in his theory of the epidemiological transition, that even if the class differential in mortality was maintained during the transition, the decline set in earlier and was faster among privileged groups.

Recent studies, investigating social inequality in health and mortality with micro-data, has however questioned the assumed pattern. These studies suggest that the modern pattern might be a fairly recent development, although the pattern differs depending on age, gender or historical context. When it comes to the adult and elderly population results from different studies have diverged, some reporting substantial differences but many, particularly those based on Swedish data, have found surprisingly small socio-economic differences or differences in the opposite direction than the expected during the first stages of the transition \citep{bevp11, bedr11, edli11,edbr12}, sharply contrasting against what we know from present-day society.

Another interesting aspect of the developent of social inequality in mortality, is its relation to economic inequality. \citet{wipi09} argue that income inequality has an independent effect on mortality, separated from the direct effect of the actual different access to economic resources. They argue that unequal societies perform less well when it comes to health than equal ones. Wilkinson and Pickett find a positive association between inequality and health (as well as other social conditions) at country level in the present-day economically developed world. This has initiated a vital scholarly debate and the topic has been extensively studied (see for example \citet{suka04, wado00}. How the association looked like in previous periods is unknown. This is of particular interest since the levels of inequality were very different from what we find during most of the recent decades. Even if it not a necessary implication, we would still assume that poorer groups were those most disadvantaged of large inequality in itself.Data are uncertain for the 19th century, but at least during the late 19th and early 20th century both income and wealth distribution was strongly skewed in Sweden. From a couple of decades into the 20th century, income and wealth inequality continuously diminished to reach the lowest point at around 1980. Thereafter, economic inequality has increased substantially until the present, although far from being as high as a century ago \citep{rowa08, rowa09}. The Swedish development resembles that of other European and North American countries in its basic features \citep{rowa14}.

It thus appears as if a more consistent social health divide appeared at some point during the mortality transition, at least in Sweden. Solid empirical evidence about the process is however lacking and studies focusing on the issue are still few. There is a need for more reliable studies from different geographical and historical settings as well as according to age and gender in order to better understand the role of socio-economic conditions for health and survival over time.

% This paper is a follow-up of an earlier paper \citep{segb16b}, presented at
% the ESHD Conference in Leuven, Belgium, 2016. Here we make a quick look back
% in time for the Skellefteå region (1851--1900), and we also consider causes
% of death from the early 20th century. The {\sc Lisa} data (Swedish population registers) allow
% us to study the effect of income and education on mortality in a more
% direct way for the period 1990--2005, and to compare it with the results of the class-based analysis.

\section{Aims and questions}

In the present paper, we take a long-term perspective on the development of social inequalities in mortality in the adult and elderly population. The areas under investigation are the Skellefteå and Umeå regions in the north of Sweden, and the time period is 1851--2013 for Skellefteå and 1901--2013 for Umeå. The main questions are:

\begin{enumerate}
\item Has inequality in mortality between social classes increased among the adult and elderly
  population, primarily benefiting the highest social positions and how is it connected to the development of economic inequality in society?

%%\item Is the mortality transition initiated by a certain social class?

\item Are there any gender differences in the effect of social position?

\item Is social position equally important among the retired population as for those in working age?
\item Is there a difference between classes in mortality from cardiovascular diseases and cancers respectively?
\item Do class inequalities in survival correspond to those depending on level of education and income categories for the period 1990 onwards?
\end{enumerate}

\section{The Skellefteå and Umeå regions}

% Figures~\ref{fig:mapp} and~\ref{fig:SKEUME}
% show the study areas and their position in Sweden.

% \begin{figure}
% \includegraphics[height=19cm]{myfigs/skum}
% \caption{The Skellefteå and Umeå regions.}
% \label{fig:SKEUME}
% \end{figure}

\begin{figure}
\includegraphics[height=19cm]{myfigs/sweden}
\caption{The Skellefteå and Umeå regions in Sweden.}
\label{fig:mapp}
\end{figure}

The Skellefteå and Umeå regions (Figure~\ref{fig:mapp}) are part of the county of Västerbotten in the north of Sweden along the coast of the Gulf of Bothnia. This was a remote part of the country, where communications with the rest of Sweden were restricted until the late 20th century, dominated by agriculture and characterise by a vulnerable economy. The county was vulnerable to harvest failures and several severe famines struck the regions during the 1800s. One living long in memory was the famine of 1867–-1868 \citep{segb14}. During the long winters, sea communication was hindered due to the Gulf of Bothnia being frozen. Communications during winter used the winter roads. The first ships during the year came to this area in May or even as late as June in some cases \citep{fah56}. Towards the end of the 19th century, the Swedish railway system reached this part of Sweden, thereby facilitating contacts with the rest of Sweden, improving the economy and mitigating the effects of harvest failures. The inland was particularly remote. During the 19th century, our regions became increasingly integrated in the same epidemiological pattern as the rest of Sweden. Diseases reached to places that earlier were spared.

The Skellefteå region consists of a selection of parishes surrounding the town of Skellefteå, founded in 1845 but with a very small population during the 19th century. The data from the period after 1960 cover the Skellefteå and Malå municipalities, the same area as for the earlier period but with the addition of two more parishes. The majority of the 19th century population lived in rural villages and hamlets, getting their livelihood from agricultural production. During the 20th century, industrialisation took place, partly by the opening of mines in the inlands. This also led to a population increase both in the town and in the rural parts, and a much more diversified economy.

Mortality was fairly low in comparison with other parts of the country and the fertility transition was late. The Skellefteå population size as defined in our data sets was 16\,473 on January 1, 1850, 41\,352 on January 1, 1900, 61\,938 on January 1, 1950 and 76\,723 at the end of the century.

The Umeå region in this study consists of Umeå urban and rural parish until 1950, and from 1960 onwards of Umeå municipality where another three parishes are included. This region has a somewhat different character from that of Skellefteå, even though agriculture dominated the economy as well for a long time. Umeå town had a small population but was substantially larger than Skellefteå town during almost the whole studied period. It was the administrative centre in the county of Västerbotten and schools and military regiments were placed here. Consequently, officials, teachers and the military were much more common in Umeå than in Skellefteå. Agriculture dominated the rural part, there were some foundries, some industries for example in forestry and small-scale production. Umeå had a much more diversified occupational structure than Skellefteå. The population size as defined in our data sets was 18\,970 on January 1, 1900, 32\,900 on January 1, 1950 and 103\,970 when the century ended.

\section{Data and variables}

The extension in time of data at the Demographic Data Base, Umeå University (DDB) by including the period between 1900 and 1950s makes it possible to investigate the long-term development of social inequality in mortality, in the present case from 1851 to 2013. The databases provide us with micro-data on a sufficiently large population and with substantial social diversity from the 20th century for the Skellefteå and Umeå regions \citep{wee16}.The data for the present study comes from two large population databases at the DDB, Umeå University. The early period is covered by
the database Poplink, the digitisation of historical parish registers for
the Skellefteå and Umeå region \citep{wee16}. Poplink is based on linked parish records, allowing
us to reconstruct life biographies on people as long as they remained in
the region. The records are linked within but not between the regions. Data
from Poplink are accessed for the period 1851--1950s.

The other large data-set is extracted from the Linnaeus database
\citep{malm10}. Linnaeus is based on different linked population registers from
1960 to 2013 and is used within the ageing program at CEDAR, Umeå
University. The study period from 1960 to 2000 are constructed from censuses every fifth
year 1960--1990, with additional information on deaths from National Board of Health and Welfare. For the period 2002--2013 we use the information from the yearly population registers (LISA) together with death information from National Board of Health and Welfare.

Information on individuals are anonymised and the two databases are not
linked. This stops us from following individuals present
in both databases throughout their lives, thus they are treated separately. This
also makes it impossible to add information on individuals in the Linnaeus
database from what we potentially could find in Poplink, for example family
background or occupations earlier in life.

In the data-set analysed here, all individuals 40 years and older ever being resident in either
of the regions are included. The data file contains the variables on sex,
birth date, death date, first and last date of observation and type of
entrance/exit.Total number of persons are 84.599 leading to 35.001 deaths in Poplink and 197.864 persons leading to 66297 deaths in the Linnaeus database. (SKA VI HA FIGUREN MED THE SAMPLING FRAME HÄR?)

\subsection{Presence periods}

Differences in available information in the two datasets as well as within the Linnaeus data makes it necessary to apply diverging approaches when it comes to identification of presence periods within our
regions. The Poplink data provides us with exact dates or at least year of
start and exit of the presence of all individuals in our data, allowing us to have full and continuous control of the de jure population. This is however not the case with the Linnaeus database. For the period 1960--2000 we use presence in the censuses 1960--1990 together with the yearly population registers 1986--2000 and information on deaths from the National Board of Health and Welfare to identify the observation periods in our analyses. We have however no exact information on what happened in-between the censuses, for example the exact date of exit for persons that moved between two censuses. We have solved this by considering everyone as resident in our regions during the complete five-year period between censuses if the person is missing in the later census. This overestimates their presence in our regions, but should not bias the results when it comes to socioeconomic differences. We furthermore miss those that both moved in and moved out or vice versa between censuses. Exit dates are exact if they refer to death, and in these cases we consider them as constantly present from the last census to the death data even if the death took place in another part of Sweden.

For the last period, 2002--2013, we use the yearly population registers (also known as LISA data from Statistics Sweden), together with information on deaths from National Board of Health and Welfare. This data allows good control of the presence periods in the studied regions.

\subsection{Social class}

In this paper we analyse mortality differences according to social class based on the classification scheme HISCLASS \citep{lema11}. The only type of socio-economic information available both in Linnaeus and Poplink is occupations. Thus, analysing according to social class is apart from being a scientific choice, also a practical one. Using occupational data is the only and obvious alternative for making a longitudinal study of socio-economic mortality differences with consistent classification schemes. Social position is treated as a time-dependent variable.

The availability to information on occupations varies over time and between sources. Occupations in Poplink are coded into HISCO \citep{lema02} and the original DDB coding system. The censuses 1960--1990 include occupational codes based on the system used in Sweden at that time, NYK (Nordisk Yrkesklassificering). The yearly population registers (LISA) do not report occupations until 2001, thus there is a window of 11 years from 1991 to 2001 where we lack information. In the censuses, we have solved this by extending the periods for which we consider the observed social class to be valid. Thus, the social class in the census 1990 is considered to be kept until 2000. This strategy is also used for other cases of missing social class. Furthermore, we consider the social position after retirement (in Sweden usually at age 65) to be the same as the last during working age.

The use of different data sets, forces us to apply different approaches towards defining the social position. In the Poplink data, some titles are not precise enough to identify the size of the enterprise, or the division between skills, semi-skilled and unskilled workers. We consider however the validity of the information to be sufficient for identification of social position. Another problem is that Poplink usually only provides the occupation only for the head of household, thus under-estimating female working-force participation as well as that of adult children residing with their parents \citep{vik10}. In this study and according to the research questions in the present paper, we have chosen to categorise wife according to the position of the head of household, usually the husband/father, assuming that the family shares the same socio-economic position. This strategy makes the problem of under-recording less problematic.

Female labour is much better covered in the Linnaeus database. A complication is however that it is  difficult to define households in the same way as in Poplink. Extra-marital cohabitation, connected to the Second Demographic Transition \citep{van87, van03}, became common and female labour force participation developed as the norm in Sweden during this period. The occupation of the husband as signifier of social class became, if not obsolete at least, less relevant. All included persons are signified by their individual occupation for this period. Still, the results for the different periods are not completely comparable and makes the interpretation of the development for women more shaky.

The classification scheme we use is  based on HISCLASS \citep{lema11}, a class-based system. The scheme uses HISCO for assigning categories. HISCO (Historical International Standard Classifications of Occupations) \citet{lema02}  is a classification system of historical occupations, categorizing occupations according to tasks and skills. We have merged the original 12 classes in HISCLASS, based on position on the labour market, ending up with the following three and an additional fourth category with unknown class:

\begin{enumerate}
\item {\tt elite}, HISCLASS 1 and 2. Higher managers and professionals.
\item {\tt middle}, HISCLASS 3, 4, 5, 6 and 8. Lower managers and farmers.
\item {\tt worker}, HISCLASS 7, 9, 10, 11 and 12. Workers of different skills including farm workers.
%%\item Lower skilled farm workers, HISCLASS 10 and 12, Lower and unskilled farm workers
\end{enumerate}

For the Poplink data, we have used a script re-coding HISCO into HISCLASS,
provided to us by Jonas Helgertz, Lund University. The procedure for
occupations in the Linnaeus database is a little more complicated. The
occupations in the censuses (NYK) are first transferred to the system
later in use, SSYK (the Swedish version of ISCO88). Thereafter the
SSYK, both those transferred from the censuses and those given in the
population registers 2001-2013, can be re-coded into HISCLASS.

Figure~\ref{fig:hiscskum} shows the distribution (per cent) of exposure
time over class  for our selected periods. The category of workers is
fairly stable over time. The highest class has
increased substantially, although from a very low level in the 19th
century. The increase of academics in connection to the establishment of
Umeå university explains much of this increase. Within the middle group, there has been a transfer from a majority of farmer during the 19th century to lower white collar from the middle of the 20th century.

<<hiscskum, fig.cap = "Exposure by hisclass (from the top: elite, middle, workers, missing) and time period", fig.height = 4.5>>=
xx <- with(vb0Tab, tapply(exposure, list(period = period, hisclass = hisclass), sum))
yy <- round(prop.table(xx, 1) * 100, 1)
barplot(t(yy)[4:1, ], col = c("red", "blue", "green", "black"), las = 2, ylab = "Per cent")
@

Those that we first observe in the first included censuses often lack stated occupations. This is especially the case for women and elderly being retired. The proportion with missing values ({\tt NA}) included is presented in Table~\ref{tab:nana}. It was very low
during the periods until 1950, between 4 and 8 \% lacked information
allowing determination of class. This increased considerably during the first two decades of data brom the Linnaeus database. During the 1960s, every second person lacks class category
and almost a third have missing information during the 1970s. As a consequence, we only report the results for men 40--64 for the 1960s and 1970s. For the most recent periods, the information is fairly complete.

<<nana, results = 'asis'>>=
library(xtable)
##load("data/vb0.rda")
##vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "lowMan", "lowWhiteC", "worker"))
x <- with(vb0, table(period, hisclass, useNA = "ifany"))
x <- round(prop.table(x, 1) * 100)
xtable(x, caption = "Hisclass by time period, per cent.", label = "tab:nana", digits = 0)
@

\subsection{Marital status}

Marital status is treated as a time-dependent variable in the data file. The statuses used are unmarried, married and dissolved marriages (without distinguishing between widowed and divorces). The status unmarried is often not explicitly noted in the source, which makes the status unknown in the database Poplink for people for people moving into our parishes. If no explicit status is given in the sources and partner is missing, the status has been set to unmarried.

Regarding the distribution of marital status over time, see Figure~\ref{fig:civstskum}.The figure indicate an under-registration of those married and and over-registration of previously married during the 1960s. (WHY?)

<<civstskum, fig.cap = "Exposure by civil status (from the top: previously married, married and never married) and time period", fig.height = 4.5>>=
xx <- with(vb0Tab, tapply(exposure, list(period = period, civst = civst), sum))
yy <- round(prop.table(xx, 1) * 100, 1)
barplot(t(yy)[3:1, ], col = c("red", "blue", "black"), las = 2, ylab = "Per cent")
@

\subsection{Cause of death}
In the present study, we analyse mortality in the two most common disease
groups in the ages we are investigating -- cardiovascular diseases and
cancers. National compulsory reporting of cause of death was introduced in 1911, thus our analysis of the two disease groups starts from that year. Classification systems have however changed. Sweden did not join ICD until 1952, but the historical causes of death in the Poplink database have been coded according to the ICD10 system. In the present study, we have identified the codes relevant for cardiovascular diseases and cancers in the ICD 8, 9 and 10.

The distribution of causes of death is shown in Figure~\ref{fig:causeskum}. Notice the high proportion of other causes in the first two periods. This category include cases with unknown or unspecific causes which has implications for the cause-specific analyses (see later discussion). The total dominance of our two disease groups is apparent.

<<causeskum, fig.cap = "Causes of death by time period (top: cancer, middle: cardiovascular, bottom: other).", fig.height = 4.5>>=
vb01 <- vb0[vb0$event == 1, ]
cancer <- tapply(vb01$cancer, vb01$period, sum)
heart <- tapply(vb01$heart, vb01$period, sum)
death <- tapply(vb01$period, vb01$period, length)
other <- death - cancer - heart
ttt <- rbind(cancer, heart, other)
ttt <- prop.table(ttt, 2) * 100
barplot(ttt[3:1, -(1:2)], col = 3:1, las = 2, ylab = "Per cent")
@

\subsection{Periods for analysis}
We have made separate analyses for the different periods. The century 1851--1950 is divided into four periods of 25 years each. The nineteenth century periods represent a mainly pre-industrial society dominated by agriculture and with restricted welfare provisions. During the years 1901--1950, industrialisation had started and the urban environments increased in population size. The modern welfare state began to develop during this time. For the period 1960 onwards, there are separate analyses on decades except for the two last periods. Sweden had developed into a wealthy welfare society with low income inequality in international comparison, particularly until the 1990s. A financial crisis struck Sweden in the early 1990s, which among other things increased unemployment and changed parts of the welfare system to not being as generous as before. The start of the last period (2008--2013) coincides with the international financial crisis.

\subsection{Income}
Income is based on disposable individual income from the LISA database at Statistics Sweden, grouped into quartiles. It is only analysed from 1990 to 2005.

\subsection{Education}
Education is here analysed only for the period 1990 to 2005. Years of education are given in 4 categories and is based on level of education according to categories in the LISA database at Statistics Sweden. It ranges from basic education (less than nine years schooling) at the lowest level, to high academic degree (PhD and similar) at the highest level. Which are the new categories?)

\section{Models}

The proportional hazards model in a survival analysis context is used,
allowing for adjustment for civil status and rural/urban environment. Since
the central explanatory variable, {\sc Hisclass}, do not follow the
proportionality property, the analyses are stratified with respect to the
variable, and the main results are presented graphically.

Another practical problem is the size of that data set:
It contains no less than 2\,432\,871 records of 236\,011 individuals, of
which 68\,244 are observed to die.
This makes it very inconvenient to use the nowadays so popular semi-parametric
technique known as
\emph{Cox regression} \citep{cox72,gb02}. Instead we use a parametric model
that allows for reduction of the data
to \emph{sufficient statistics} \citep{fisher22}. The idea is to keep all
relevant covariates \emph{categorical}, and then create a table with all
possible combinations of levels of the covariates. For each such
combination, the total amount of exposure and the number of deaths are
calculated.  We have two sexes, three civil statuses, ten periods, two
urban/rural groups, four hisclasses at age 40,  four hisclasses at age 60,
ten age groups, and 22 cohorts, in all 422\,400 possible
combinations. However, in this scheme there are many structural zeroes
(impossible combinations) and some random zeros, so the actual number of
combinations we end up with is 24\,326. The data set is thus reduced in
size by 99 per cent.

The parametric model we use for the baseline hazard function is piecewise
constant on five-year intervals, so we arrive at the following proportional
hazards regression model:

\begin{equation}\label{eq:prop}
h(t; \lambda, \boldsymbol{\beta}, \textbf{x}) =
\lambda(t)e^{\boldsymbol{\beta} \mathbf{x}}, \; \; 40 < t \le 90,
\end{equation}
where

\begin{equation*}
\begin{split}
\mathbf{x} &= (x_1, \ldots, x_n) \text{\ is a vector of covariates,} \\
\boldsymbol{\beta} &= (\beta_1, \ldots, \beta_n) \text{\ is the corresponding vector of regression
parameters.} \\
\end{split}
\end{equation*}
and
\begin{equation*}
\lambda(t) = \alpha_i,\; 40 + 5(i - 1) < t \le 40 + 5i, \; i = 1, \ldots, 10.
\end{equation*}

Since all components of $\mathbf{x}$ are categorical covariates, the
estimation of \eqref{eq:prop} can be performed in two ways, first by the
original data set and a survival analysis routine handling a piecewise
constant hazard, or second, by the reduced data set (table) and
\emph{Poisson regression}. The second way is faster than the first by a
factor of size at least 100, often much more, depending on hardware, and
the point is that the two ways give exactly the same results. This is
because the table is a \emph{sufficient statistic} for \eqref{eq:prop} and
the original data set.

The analyses are performed in the \R\ environment for statistical computing
and graphics \citep{cran}, especially using the package {\tt eha} \citep{gb12,eha}
The proportional hazards model allow us to scrutinise the differences and
the development in more detail. We divide the analysis into two age groups,
40--64 and 65--89 years of age, the first age group representing people that were still mainly in the workforce, and the second those that mainly were retired. The model controls for period, marital
status and whether the individual resides in the urban or rural part. There
are no serious signs of deviation from proportionality assumption, except
regarding {\sc Hisclass}.

We analyse both regions together but perform separate analyses for men and
women and for the age groups 40--64 and 65--89. The model in the analyses
include social class, sex and marital status. Each decade
during the 20th century gets separate analyses, but we then combine results
in order to show the development of the social patterning of mortality
during the studied period. We present the results for the period 2001--2013
separately. We suggest a heterogeneity index in order to investigate a
possible homogenisation of mortality between social classes.

The important explanatory variable, {\sc Hisclass}, is included in the
proportional hazards models as a \emph{stratification} variable. The main
reason for this choice is that the effect of {\sc Hisclass} on survival is
non-proportional, that is, it varies with age. With the reduced data set
and Poisson regression, this is simply achieved by introducing an
interaction term between {\tt age} and {\tt hisclass}. Then, the main
results are
presented \emph{graphically}. There is some  lack of information in the higher
ages around 1960--1980 (see Table~\ref{tab:nana}), because that kind of information is simply missing
to a large extent.

\section{Results, all causes of death}

<<colorc>>=
## Colors for plotting strata:
scol <- c("blue", rep("darkgreen", 2), "red")
## Common for all groups:

labb <- levels(vb0Tab$period)
nper <- length(labb)
#source("R/getPHglm.R")
## End 'common'
@

Models are fitted separately for each time period, from  1851 to
2013, with the exception of the period 1951--1960. We stratify on HISCLASS
and control for the covariates {\tt sex} and {\tt   civst}. The cumulative
hazards at the end of the time period (65 if 40--64, 90 if 65--89) are the
main target in our analysis: These numbers are used in the process of
graphically illustrating the development over time.

%creating an \emph{heterogeneity index} (a better name is needed:
%\emph{health disparity index}, \emph{Social Inequality Mortality Index, (SIMI)}?
%Something catchy.).

\subsection{Ages 40--64}

%The distribution over time periods.
<<getske4060>>=
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, women."
scap1w <-  "Cumulative hazards, ages 40--64, women."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, men."
scap1m <-  "Cumulative hazards, ages 40--64, men."
##
cap0 <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom)."
scap0 <- "Total hazard by HISCLASS, ages 40--64."
cap12 <- "Total hazard of dying before age 65 for a 40 year old person standardized by HISCLASS and decade, women (top) and men (bottom)."
scap12 <- "Total hazard standardized by HISCLASS, ages 40--64."
cap12log <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades."
scap12log <- "Total hazard by HISCLASS, ages 40--64, last years."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, women."
scap2w <-  "Cumulative hazards, ages 65--89, women."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, men."
scap2m <-  "Cumulative hazards, ages 65--89, men."
##
cap22 <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom)."
scap22 <- "Total hazard by HISCLASS, ages 65--89."
cap22log <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades."
scap22log <- "Total hazard by HISCLASS, ages 65--89, last years."

##
vb4064 <- vb0Tab[as.numeric(as.character(vb0Tab$age)) < 64.5, ]
vb4064$age <- factor(vb4064$age)
##vb4064$hisclass <- vb4064$hisclass40
@

We start by looking at cumulative hazards by {\sc Hisclass} over the last four time
periods, in order to illustrate the process. They are adjusted for {\tt civil status}. We present separate analyses  for women and men, see
Figures~\ref{fig:hazplotw} and~\ref{fig:hazplotm}.

%\subsubsection{Women}

<<hazcalw>>=
source("R/getPHglm2.R")
##nper <- 4
nper <- length(labb)
resw <- getPHglm2(vb4064[vb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb4064[vb4064$sex == "male", ],
            nper, labb)
@

<<hazplotw, fig.cap = cap1w, fig.scap = scap1w, fig.height = 7>>=
source("R/plotHaz.glm.R")

plotHaz.glm(resw[-(1:6)], main = labb, col = scol, layout = c(2, 2))
@

<<hazplotm, fig.cap = cap1m, fig.scap = scap1m, fig.height = 7>>=
plotHaz.glm(resm[-(1:6)], main = labb, col = scol, layout = c(2, 2))
@

From Figures~\ref{fig:hazplotw} and \ref{fig:hazplotm} it is obvious that
the idea of proportional hazards is not reasonable. We therefore use another approach.

The "TMR", that is, the summed hazards over the age interval 40--64 is shown in
Figure~\ref{fig:cumsw}.

<<cumsw, fig.cap = cap0, fig.scap = scap0, fig.height = 10>>=
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, ylab = "H(65|40)",
                main = "Women 40-64")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, ylab = "H(65|40)",
                main = "Men 40-64")##, exclude = 5:6)
par(oldpar)
@

Figure~\ref{fig:cumsw} shows mortality levels during the different periods. Obviously, and not surprisingly, mortality is decreasing over time, and we can already identify some interesting social patterns in mortality which will be commented more on below. Apart from
the social pattern that will be commented on below, there are a couple of
observations to be made from the figure. Women have consistently
better survival than men. This is a well-known phenomenon in Swedish
demographic history. Male mortality has almost always been higher than
female in all age groups \citep{wil99, suwi07}. We also observe a
substantial decline in mortality for both sexes, resulting in smaller
absolute differences between social classes. The compression that is
visible for later decades is partly explained by the decreasing general
mortality. It is not
obvious that differences between social classes have diminished. This
conclusion has been confirmed by separate analyses (results not shown). There is no
clear indication of homogenisation of mortality levels between social
classes.

However, we are mainly interested in illustrating the change in relative position among the HISCLASSES over time rather than the absolute levels, so we \emph{standardize} the levels in each time period. It is illustrated in
Figure~\ref{fig:cumswstand}. In the rest of the paper, we only present these relative mortality levels for the different classes.

<<cumswstand, fig.cap = cap12, fig.scap = scap12, fig.height = 10>>=
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(65|40)", main = "Women 40-64")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(65|40)", main = "Men 40-64")##, exclude = 5:6)
par(oldpar)
@

%%<<cumsm, fig.cap = "Total haz of deaths in ages 40--64 per individual by HISCLASS and decade, men.">>=
%%@

%% HHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHH

Notice that we have not analysed the 1950s due to
the fact that we only have data for some parishes at that time,
consequently having no corresponding population for that
decade. Furthermore, we need to be cautious when interpreting the results for the 1960s in particular, but also for the 1970s due to much missing information on HISCLASS for these decades, something that have restricted the size of the population observed. The changing economy
also have consequences for the analysis, particularly when it comes to the
agricultural sector. While farmers were common occupations during the first
half of the 20th century, these occupations has constantly decreased in
numbers thereafter. This means that the second HISCLASS group has somewhat different character during the last periods in the figures.

The social pattern of mortality among women is quite much in accordance with previous
analyses from the present research group on other regions in northern
Sweden and for other periods \citep{edli11,edbr12,segb16}. The highest
social class usually have comparatively low mortality, while working
class women have high. This is especially the case during the latest
decades of the century and until 2013. We once again remind that the results from the 1960s must be interpreted with caution due to the large proportion of missing cases.

A different pattern appears among men. The groups we would expect to have
the lowest mortality, if we assume that access to different resources and
having a high status determines survival, instead have the highest during the nineteenth and
much of the 20th century. This is also in accordance with our previous studies
\citep{edli11, edbr12, segb16}. Workers on the other hand have comparatively
low mortality until the 1960s. This changes however towards the end the 20th century. From
the 1980s, there is a clear advantage when it comes to survival for higher
managers and professionals. The expected social gradient in mortality has developed. This is also what we usually find in analyses
of social health inequalities in present-day Sweden as well as in most
other countries. In the regions we are studying, this is however a quite
recent phenomenon.

\subsubsection*{The effect of income and education, 1990--2005}

A common alternative to a class analysis is to use income or education for investigating socio-economic health differences. Comparing different approaches are furthermore valuable in order to understand why some are advantaged and others disadvantaged. For the time being, we have to restrict such a comparative analysis to the last decades.

The analysis for the period 1990--2005 is based on a separate analysis with income and education as explanatory variables and with controls of period and marital status.

<<lissyoung>>=
library(eha)
library(xtable)
liss4064 <- loTab[loTab$age < 64.5, ]
@

The statistical significance of the covariates are shown in Table~\ref{tab:liswom}.
<<liswom, results = 'asis'>>=
fitw1 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss4064[liss4064$sex == "female", ])
##cat("Women:\n")
drw <- drop1(fitw1, test = "Chisq")
xtable(drw, caption = "Effect of covariates for women 40-64, 1990-2005.", label = "tab:liswom")
##round(summary(fitw1)$coef[-(1:13), ], 4)
@

The estimated effect sizes for women are shown in Figure \ref{fig:liswomfig}.

<<liswomfig, fig.cap = "Effect sizes, women 40-64, 1990-2005.", fig.height = 4>>=
source("R/plotWEduc.R")
plotWEduc(fitw1)
@

And for men:

<<lismen, results='asis'>>=
fitm1 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss4064[liss4064$sex == "male", ])
##cat("\nMen:\n")
drm <- drop1(fitm1, test = "Chisq")
xtable(drm, caption = "Effect of covariates for men 40-64, 1990-2005.")
##round(summary(fitm1)$coef[-(1:13), ], 4)
dcol <- c("blue", "darkgreen", "darkgreen", "red")
@

<<lismanfig, fig.cap = "Effect sizes, men 40-64, 1990-2005.", fig.height = 4>>=
#source("R/plotWEduc.R")
plotWEduc(fitm1)
@

<<plotliss1, fig.cap = "Cumulative hazards, ages 40--64, 2002--2013.", fig.height = 4>>=
#oldpar <- par(mfrow = c(1, 2))
#plot(fitw1, fn = "cum", col = dcol, main = "Women")
#plot(fitm1, fn = "cum", col = dcol, main = "Men")
#par(oldpar)
##summary(fitw1)
##summary(fitm1)
@

%See Table~\ref{tab:tablisswom1} the PH regression for women.
<<tablisswom1, results = "asis">>=
#fitwo1 <- coxreg(Surv(enter, exit, event) ~ DISPINK + CIV + SUN2000NIVA,
#                 data = liss4064[liss4064$sex == "female", ])
#dr <- drop1(fitwo1, test = "Chisq")
#ltx(fitwo1, dr = dr, caption = "PH regression, women aged 40--64, 2002--2013.",
#    label = "tab:tablisswom1")
@

%See Table~\ref{tab:tablissman1} the PH regression for men.
<<tablissman1, results = "asis">>=
#fitma1 <- coxreg(Surv(enter, exit, event) ~ DISPINK + CIV + SUN2000NIVA,
#                 data = liss4064[liss4064$sex == "male", ])
#dr <- drop1(fitma1, test = "Chisq")
#ltx(fitma1, dr = dr, caption = "PH regression, men aged 40--64, 2002--2013.",
 #   label = "tab:tablissman1")
@
Both education and income were strongly significant, as well as marital status as expected. Those with the two lowest education level hade the highest mortality, while those with highest level of education had by far the best survival. The differences must be considered large, and it follows almost a continuous improvement according to increasing levels. The same can be said concerning income. Higher incomes result in lower mortality. This is also what could be expected, and conforms to what we know of social health differences in present-day Sweden. Notice however that we cannot definitely establish a causal relationship between income and mortality in particular. Having low incomes may be caused by bad health.

\subsection{Ages 65--89}

%The distribution over time periods.
<<getske6089>>=
vb6589 <- vb0Tab[as.numeric(as.character(vb0Tab$age)) >= 64.5, ]
vb6589$age <- factor(vb6589$age) # NOTE:: Change to '6589'!
##vb6589$hisclass <- vb6589$hisclass60
@

We now turn to the elderly, i.e. those that mostly have left the workforce.
Corresponding to the previous analysis, we use the value of the cumulative hazards at
duration 25 (age 90), starting from age 65 as a summary measure.

The general pattern and development of mortality (not shown here) do not differ that much from the younger age group, of course representing a much higher mortality level. Mortality among women was, as expected, lower than among
men. We find a clear decline in mortality, but this decline did not
set in definitely until sometimes after the Second World War (results not shown here), i.e. later than for the age group 40--64 years. This corresponds to the
Swedish development where we find a turning point at this time with better
survival among elderly, to a large extent caused by lower mortality in
cardiovascular diseases. The results for the 1960s are however uncertain as
discussed above, especially for women due to the difficulty to determine HISCLASS for the majority of them.

Concerning the social pattern, the pattern is a little more difficult to interpret, compared to the younger age group (the new analyses differ somewhat from the ones in \citet{segb16}). For women the results are mixed. We do not find the same clear advantage for elite women during period until 1950, and neither do we find any large disadvantage for working class elderly women. In some periods, the social differences were small, and the elite even had substantially higher mortality during the early 20th century. Notice however that this group was rather small at the time. During the recent decades, the pattern is not very clear, but for the two last periods, the elite have the lowest mortality. It is however not possible to identify a clear social gradient in mortality.

For men we find indications of the same intriguing change among elderly as for
the adult population 40--64 years. Mortality was in fact substantially
higher among men from the elite  but the disadvantage is mainly apparent during the 19th century. Differences are fairly small during the first half of the 20th century. Only during the second half of the 20th century, we find periods with much better survival for the highest social class while retired workers had fairly high mortality. But even in recent times, the pattern is not completely clear. During the last periods, we do not find a social gradient on HISCLASSES among elderly men (even in this case, the new analyses differ somewhat from the ones in \citet{segb16}). This is quite unexpected, since other studies have shown clear socio-economic differences in Sweden during recent decades.

<<cumswold, fig.cap = cap22, fig.scap = scap22, fig.height = 10>>=
##source("R/plotED.R")
resw <- getPHglm2(vb6589[vb6589$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb6589[vb6589$sex == "male", ],
            nper, labb)

oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(90|65)", main = "Women 65-89")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(90|65)", main = "Men 65-89")##, exclude = 5:6)
par(oldpar)
@

\subsubsection*{The effect of income and education, 1990--2005}

The fact that we find rather unclear associations between HISCLASSES and survival, makes alternative approaches to the question of socio-economic health differences interesting. As for the younger age group, we have included an analysis for the period 1990--2005 with income and education as explanatory variables.

<<lissyoung2>>=
library(eha)
library(xtable)
##load("data/loTab.rda")
##loTab$period <- factor(loTab$period)
liss6589 <- loTab[loTab$age > 64.5, ]
@

The statistical significance of the covariates are shown in
Table~\ref{tab:liswom2}
<<liswom2, results = 'asis'>>=
fitw2 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss6589[liss6589$sex == "female", ])
##cat("Women:\n")
drw2 <- drop1(fitw2, test = "Chisq")
xtable(drw2, caption = "Effect of covariates for women 65-89, 1990-2005.", label = "tab:liswom2")
##round(summary(fitw1)$coef[-(1:13), ], 4)
@

The estimated effect sizes are shown in Figure \ref{fig:liswomfig2}

<<liswomfig2, fig.cap = "Effect sizes, women 65-89, 1990-2005.", fig.height = 4>>=
#source("R/plotWEduc.R")
plotWEduc(fitw2)
@

And for men:

<<lismen2, results='asis'>>=
fitm2 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss6589[liss6589$sex == "male", ])
##cat("\nMen:\n")
drm2 <- drop1(fitm2, test = "Chisq")
xtable(drm2, caption = "Effect of covariates for men 65-89, 1990-2005.")
##round(summary(fitm1)$coef[-(1:13), ], 4)
dcol <- c("blue", "darkgreen", "darkgreen", "red")
@

<<lismanfig2, fig.cap = "Effect sizes, men 65-89, 1990-2005.", fig.height = 4>>=
#source("R/plotWEduc.R")
plotWEduc(fitm2)
@

These figures show clear associations both with income and education and for women as well as for men. More years of education resulted in lower mortality. For women the association with additional levels became however smaller at the higher levels, the effect is more visible at the steps above the basic levels. In a similar fashion, increased incomes is associated with improved survival both for men and for women. Apparently, education and income is better to catch up socio-economic differences compared to hisclass, at least for this period.

\section{Results, cardiovascular death}

We now turn to a special analysis of mortality for the two dominating causes of death among the adult and elderly population. We start with cardiovascular diseases. Many studies have shown that socio-economic conditions are strongly associated with mortality in this disease group. Belonging to working class, having less education and low income increases the mortality risks. We would therefore expect to find a clear social pattern in the present analysis.

First, a general observation is that the analysis of causes of death and their development are more difficult to interpret, partly because the number of cases are fewer but also that the quality and coverage of reported death causes differ. Thus, many deaths during the 19th century were not diagnosed at all, and there was still a substantial part of all death in the early twentieth century where death cause is designated as unknown or dying from old age, despite reporting causes of death had become mandatory. For the period 1901-1925, the proportion of deaths from cardiovascular diseases increased from 25 to 43 per cent for the age group 40 and above, but at the same time unknown cause and death of old age decreased from 28 percent to 18 per cent. Are there many cardiovascular deaths disguised there?
<<causes41>>=
cvb <- vb
cvb$event <- cvb$heart ## Can this be done better?
cvb <- cvb[!(cvb$period %in% c("1851-1875", "1876-1900")), ]
cvb$period <- factor(cvb$period)

# load("data/vb.rda")
# ##vb <- cal.window(vb, c(1961, 2001))
# ##x <- substr(vb$ULORSAK, 1, 2)
# vb$event <- vb$event & vb$heart
# vb$exposure <- with(vb, exit - enter)
# vb$cohort <- factor(vb$cohort)
# vb$period <- factor(vb$period)
cvbTab <- aggregate(cvb[, c("event", "exposure")],
                    cvb[, c("sex", "age", "period", "cohort",
                           "urban", "civst", "hisclass")],
                    FUN = sum)
 cvbTab$hisclass <- factor(cvbTab$hisclass)
 cvbTab$age <- factor(cvbTab$age)
@

<<41colorc>>=
## Colors for plotting strata:
scol <- c("blue", rep("darkgreen", 2), "red")
## Common for all groups:

labb <- levels(cvbTab$period)
nper <- length(labb)
##source("R/getPHglm.R")
## End 'common'
@

As before, models are fitted separately for each ten-year time period, from  1851 to
2013. We stratify on HISCLASS
and use the covariates {\tt sex} and {\tt   civst}. The cumulative
hazards at the end of the time period (65 if 40--64, 90 if 65--89) are the
main target in our analysis: These numbers are used in the process of
graphically illustrating the development over time.

%creating an \emph{heterogeneity index} (a better name is needed:
%\emph{health disparity index}, \emph{Social Inequality Mortality Index, (SIMI)}?
%Something catchy.).

\subsection{Ages 40--64}

%The distribution over time periods.
<<41getske4060>>=
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, women, cardiovascular death."
scap1w <-  "Cumulative hazards, ages 40--64, women, cardiovascular death."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, men, cardiovascular death."
scap1m <-  "Cumulative hazards, ages 40--64, men, cardiovascular death."
##
cap12 <- "Total hazard of dying before age 65 for a 40 year old person standardized by HISCLASS and decade, women (top) and men (bottom), cardiovascular death."
scap12 <- "Total hazard by HISCLASS, ages 40--64, cardiovascular death."
cap12log <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cardiovascular death."
scap12log <- "Total hazard by HISCLASS, ages 40--64, last years, cardiovascular death."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, women, cardiovascular death."
scap2w <-  "Cumulative hazards, ages 65--89, women, cardiovascular death."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, men, cardiovascular death."
scap2m <-  "Cumulative hazards, ages 65--89, men, cardiovascular death."
##
cap22 <- "Total hazard of dying before age 90 for a 65 year old person standardized by HISCLASS and decade, women (top) and men (bottom), cardiovascular death."
scap22 <- "Total hazard standardized by HISCLASS, ages 65--89, cardiovascular death."
cap22log <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cardiovascular death."
scap22log <- "Total hazard by HISCLASS, ages 65--89, last years, cardiovascular death."

##
cvb4064 <- cvbTab[as.numeric(as.character(cvbTab$age)) < 64.5, ]
cvb4064$age <- factor(cvb4064$age)
@

We start by looking at cumulative hazards by {\sc Hisclass} over time
periods. They are adjusted for {\tt civil status}. We present separate analyses for women and men.

%\subsubsection{Women}

<<41hazcalw>>=
source("R/getPHglm2.R")
resw <- getPHglm2(cvb4064[cvb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(cvb4064[cvb4064$sex == "male", ],
            nper, labb)
@

We will use as a summary measure the value of the cumulative hazards at
age 65, given survival to age 40.

<<41cumsw, fig.cap = cap12, fig.scap = scap12, fig.height = 10>>=
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE, ##exclude = 1:2,
                    ylab = "TMR(65|40)",  main = "Women 40-64")
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE, ##exclude = 1:2,
                    ylab = "TMR(65|40)", main = "Men 40-64")
par(oldpar)
@

%%<<cumsm, fig.cap = "Total haz of deaths in ages 40--64 per individual by HISCLASS and decade, men.">>=
%%@

Figure~\ref{fig:cumsw} shows mortality levels during different periods from 1901 onwards. A couple of observations can be made. Women from the highest social class had almost consistently lower mortality from cardiovascular diseases compared to others.This is apparent even for the periods with deficient cause of death data. The pattern concerning the other groups are less clear, but if we compare the two largest groups, the workers and the lower manager/farmer group, the differences were rather small during most of the time. A new pattern has however developed from 1990 onwards. Now working class women have substantially higher cardiovascular compared to others.

For men, we find a different pattern, a pattern however that to a large extent is consistent with what we observed concerning all-cause mortality. The elite group had much higher cardiovascular diseases up until the middle of the 20th century, but instead have the lowest mortality from 1980 onwards. Working class men had a fairly low cardiovascular mortality, but in the same way as among women, the levels were fairly similar to the one of the other large group, lower manager/farmers. And in the same way as for women at this age, working class me have by far the highest mortality from this cause of death during the last decades. A social hierarchy in cardiovascular mortality seems to have developed both for men and women.

%% HHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHH

% \subsubsection*{The effect of income and education, 1990--2005}
% 
% We have also performed the analysis on the alternative approaches on socio-economic differences for the causes of death. It is based on a separate analysis with income and education as explanatory variables.
% 
% <<41lissyoung>>=
% library(eha)
% library(xtable)
% ##load("data/loTab.rda")
% loTab$period <- factor(loTab$period)
% liss4064 <- loTab[loTab$age < 64.5, ]
% @
% 
% 
% The statistical significance of the covariates are shown in
% Table~\ref{tab:41liswom}
% <<41liswom, results = 'asis'>>=
% fitw1 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss4064[liss4064$sex == "female", ])
% ##cat("Women:\n")
% drw <- drop1(fitw1, test = "Chisq")
% xtable(drw, caption = "Effect of covariates for women 40-64, 1990-2005, cardiovascular death.", label = "tab:41liswom")
% ##round(summary(fitw1)$coef[-(1:13), ], 4)
% @
% 
% The estimated effect sizes are shown in Figure \ref{fig:41liswomfig}
% 
% <<41liswomfig, fig.cap = "Effect sizes, women 40-64, 1990-2005, cardiovascular death.", fig.height = 4>>=
% source("R/plotWEduc.R")
% plotWEduc(fitw1)
% @
% 
% 
% And for men:
% 
% <<41lismen, results='asis'>>=
% fitm1 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss4064[liss4064$sex == "male", ])
% ##cat("\nMen:\n")
% drm <- drop1(fitm1, test = "Chisq")
% xtable(drm, caption = "Effect of covariates for men 40-64, 1990-2005, cardiovascular death.", label = "tab:41lismen")
% ##round(summary(fitm1)$coef[-(1:13), ], 4)
% dcol <- c("blue", "darkgreen", "darkgreen", "red")
% @
% 
% <<41lismanfig, fig.cap = "Effect sizes, men 40-64, 1990-2005, cardiovascular death.", fig.height = 4>>=
% #source("R/plotWEduc.R")
% plotWEduc(fitm1)
% @
% 
% 
% <<41plotliss1, fig.cap = "Cumulative hazards, ages 40--64, 2002--2013, cardiovascular death.", fig.height = 4>>=
% #oldpar <- par(mfrow = c(1, 2))
% #plot(fitw1, fn = "cum", col = dcol, main = "Women")
% #plot(fitm1, fn = "cum", col = dcol, main = "Men")
% #par(oldpar)
% ##summary(fitw1)
% ##summary(fitm1)
% @
% 
% %See Table~\ref{tab:tablisswom1} the PH regression for women.
% <<41tablisswom1, results = "asis">>=
% #fitwo1 <- coxreg(Surv(enter, exit, event) ~ DISPINK + CIV + SUN2000NIVA,
% #                 data = liss4064[liss4064$sex == "female", ])
% #dr <- drop1(fitwo1, test = "Chisq")
% #ltx(fitwo1, dr = dr, caption = "PH regression, women aged 40--64, 2002--2013.",
% #    label = "tab:tablisswom1")
% @
% 
% %See Table~\ref{tab:tablissman1} the PH regression for men.
% <<41tablissman1, results = "asis">>=
% #fitma1 <- coxreg(Surv(enter, exit, event) ~ DISPINK + CIV + SUN2000NIVA,
% #                 data = liss4064[liss4064$sex == "male", ])
% #dr <- drop1(fitma1, test = "Chisq")
% #ltx(fitma1, dr = dr, caption = "PH regression, men aged 40--64, 2002--2013.",
%  #   label = "tab:tablissman1")
% @
% 
% The analysis of cardiovascular mortality according to education and income conforms very well to what we find for all-cause mortality, and we will therefore not delve into detail on this issue. More education as well as higher income is advantageous both for women and men for avoiding deaths related to the circulatory system. This is also well in line with our analysis based on HISCLASS above.

\subsection{Ages 65--89}

%The distribution over time periods.
<<41getske6089>>=
cvb6589 <- cvbTab[as.numeric(as.character(cvbTab$age)) >= 64.5, ]
cvb6589$age <- factor(cvb6589$age) # NOTE:: Now change to '6589'!
@

In this section we focus on cardiovascular mortality among those that mostly have left the workforce, the age group 65--89 years. In accordance with the analyses for the age group 40--64, the model adjusts for {\tt civil status} and with separate analyses for women and men. We use as a summary measure the value of the cumulative hazards at
duration 25 (age 90), starting from age 65.

%\subsubsection{Women}

<<41hazcalwold>>=
resw <- getPHglm2(cvb6589[cvb6589$sex == "female", ],
              nper, labb)
resm <- getPHglm2(cvb6589[cvb6589$sex == "male", ],
                nper, labb)
@

Concerning the social pattern, we find rather mixed results but with some similarities to the ones
for the age group 40--64. For women mortality in this disease group were somewhat higher among the elite and the lower manager/farmer group than the others until only the last decades. Elderly from a working class background have not been disadvantaged for most of the 20th century. The social class pattern has changed somewhat during the last periods and more in line to what is expected. The differences are however not that large. Cardiovascular disease mortality among elderly women do not present a consistent class pattern. Also for men, we find high cardiovascular mortality during the period 1926--1950 and the 1980s for the elite, but also for lower-managers/farmers during the first half of the 20th century. As for women, we find no clear class pattern during the last periods and the differences are quite small.The diffuse pattern of class differences in cardiovascular mortality is surprising. It is therefore of interest to find out what other aspects of social differentiation tell us.

Notice that the results for the 1960s must be interpreted with caution due to much missing data on HISCLASS among the elderly.

<<41cumswold, fig.cap = cap22, fig.scap = scap22, fig.height = 10>>=
##source("R/plotED.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "H(90|65)",  main = "Women 65-89")
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "H(90|65)", main = "Men 65-89")
par(oldpar)
@

% \subsubsection*{The effect of income and education on cardiovascular death, 1990--2005}
% 
% The analysis of cardiovascular mortality for the period 1990--2005 is in this section based on a separate analysis with
% income and education as explanatory variables. The statistical significance of the covariates are shown in Table~\ref{tab:41liswom2} for women and Table~\ref{tab:41lismen2} for men. It confirms that most of the variables are strongly significant except
% 
% <<41lissyoung2>>=
% library(eha)
% library(xtable)
% ##load("data/loTab.rda")
% ##loTab$period <- factor(loTab$period)
% liss6589 <- loTab[loTab$age > 64.5, ]
% @
% 
% 
% 
% <<41liswom2, results = 'asis'>>=
% fitw2 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss6589[liss6589$sex == "female", ])
% ##cat("Women:\n")
% drw2 <- drop1(fitw2, test = "Chisq")
% xtable(drw2, caption = "Effect of covariates for women 65-89, 1990-2005, cardiovascular death.", label = "tab:41liswom2")
% ##round(summary(fitw1)$coef[-(1:13), ], 4)
% @
% 
% 
% <<41lismen2, results='asis'>>=
% fitm2 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss6589[liss6589$sex == "male", ])
% ##cat("\nMen:\n")
% drm2 <- drop1(fitm2, test = "Chisq")
% xtable(drm2, caption = "Effect of covariates for men 65-89, 1990-2005, cardiovascular death.", label = "tab:41lismen2")
% ##round(summary(fitm1)$coef[-(1:13), ], 4)
% dcol <- c("blue", "darkgreen", "darkgreen", "red")
% @
% 
% The estimated effect sizes are shown in Figure \ref{fig:liswomfig}
% 
% <<41liswomfig2, fig.cap = "Effect sizes, women 65-89, 1990-2005, cardiovascular death.", fig.height = 4>>=
% #source("R/plotWEduc.R")
% plotWEduc(fitw2)
% @
% 
% 
% And for men:
% 
% 
% 
% <<41lismanfig2, fig.cap = "Effect sizes, men 65-89, 1990-2005, cardiovascular death.", fig.height = 4>>=
% #source("R/plotWEduc.R")
% plotWEduc(fitm2)
% @
% 
% The pattern is similar to what we find for all-cause mortality. The higher educational level, the lower cardiovascular mortality, and more income leads to better survival. This association is clear and consistent for women as well as for men. A minor difference between men and women is that mortality in the higher levels of education do not lead to lower mortality among women while that is the case among men.

\section{Results, cancer death}

While we assume clear social differences in cardiovascular diseases, this is not necessarily an obvious assumption when it comes to cancers. Different forms of cancers have very different origin and different pathways to disease and death. The social pattern of higher mortality in classes with less power and accordingly less life chances are considered to be less clear when it comes to cancers. In the present analysis we include cancer deaths 1901--2013, but with an unregistered and not analysed period between 1951--1960. There is a risk that cancer mortality is under-estimated during the first half of the 20th century, illustrated above by the large proportion of unknown or vague causes at that time. We believe however that cancers were better diagnosed than cardiovascular deaths. This is supported by the stability in proportion of cancer deaths between the two periods, 12,5 per cent 1901--1925 and 13,5 per cent for the period 1926--1950.

<<causes12>>=
cvb <- vb
cvb$event <- cvb$cancer ## Can this be done better?
cvb <- cvb[!(cvb$period %in% c("1851-1875", "1876-1900")), ]
cvb$period <- factor(cvb$period)

# load("data/vb.rda")
# ##vb <- cal.window(vb, c(1961, 2001))
# ##x <- substr(vb$ULORSAK, 1, 2)
# vb$event <- vb$event & vb$heart
# vb$exposure <- with(vb, exit - enter)
# vb$cohort <- factor(vb$cohort)
# vb$period <- factor(vb$period)
cvbTab <- aggregate(cvb[, c("event", "exposure")],
                    cvb[, c("sex", "age", "period", "cohort",
                           "urban", "civst", "hisclass")],
                    FUN = sum)
 cvbTab$hisclass <- factor(cvbTab$hisclass)
 cvbTab$age <- factor(cvbTab$age)
@

<<12colorc>>=
## Colours for plotting strata:
scol <- c("blue", rep("darkgreen", 2), "red")
## Common for all groups:

labb <- levels(cvbTab$period)
nper <- length(labb)
source("R/getPHglm.R")
## End 'common'
@

Models are fitted separately for each period, from  1851 to
2013. We stratify on HISCLASS
and use the covariates {\tt sex} and {\tt   civst}. The cumulative
hazards at the end of the period (65 if 40--64, 90 if 65--89) are the
main target in our analysis: These numbers are used in the process of
graphically illustrating the development over time.

%creating an \emph{heterogeneity index} (a better name is needed:
%\emph{health disparity index}, \emph{Social Inequality Mortality Index, (SIMI)}?
%Something catchy.).

\subsection{Ages 40--64}

%The distribution over time periods.
<<12getske4060>>=
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, women, cancer death."
scap1w <-  "Cumulative hazards, ages 40--64, women, cancer death."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, men, cancer death."
scap1m <-  "Cumulative hazards, ages 40--64, men, cancer death."
##
cap12 <- "Total hazard of dying before age 65 for a 40 year old person standardized by HISCLASS and decade, women (top) and men (bottom), cancer death."
scap12 <- "Total hazard by HISCLASS, ages 40--64, cancer death."
cap12log <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cancer death."
scap12log <- "Total hazard by HISCLASS, ages 40--64, last years, cancer death."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, women, cancer death."
scap2w <-  "Cumulative hazards, ages 65--89, women, cancer death."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, men, cancer death."
scap2m <-  "Cumulative hazards, ages 65--89, men, cancer death."
##
cap22 <- "Total hazard of dying before age 90 for a 65 year old person standardized by HISCLASS and decade, women (top) and men (bottom), cancer death."
scap22 <- "Total hazard standardized by HISCLASS, ages 65--89, cancer death."
cap22log <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cancer death."
scap22log <- "Total hazard by HISCLASS, ages 65--89, last years, cancer death."

##
cvb4064 <- cvbTab[as.numeric(as.character(cvbTab$age)) < 64.5, ]
cvb4064$age <- factor(cvb4064$age)
@

<<12hazcalw>>=
source("R/getPHglm2.R")
resw <- getPHglm2(cvb4064[cvb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(cvb4064[cvb4064$sex == "male", ],
            nper, labb)
@

%%We will use as a summary measure the value of the cumulative hazards at
%%age 65, given survival to age 40.

<<12cumsw, fig.cap = cap12, fig.scap = scap12, fig.height = 10>>=
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE, ##exclude = 1:2,
                    ylab = "TMR(65|40)",  main = "Women 40-64")
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE, ##exclude = 1:2,
                    ylab = "TMR(65|40)", main = "Men 40-64")
par(oldpar)
@

%%<<cumsm, fig.cap = "Total haz of deaths in ages 40--64 per individual by HISCLASS and decade, men.">>=
%%@

Figure~\ref{fig:12cumsw} shows relative mortality levels according to HISCLASS during the different studied periods. The social differences were small until the middle of the 20th century, for women as well as for men. For the periods after 1980, the rather small differences remain. There is an indication of the elite having lower cancer mortality for women sexes during the last periods, but these differences are small and we do not find any clear social gradient in the risks. Thus, for this age group cancer mortality is not something that distinguishes between classes.

%% HHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHHH

% \subsubsection*{The effect of income and education on cancer mortality, 1990--2005}
% 
% In the previous section, we concluded that there were no social inequality in cancer mortality according to hisclass. The question here is if the lacking socio-economic differentiation holds also when using education and income as categorising variables. The analysis for the period 1990--2005 is based on a separate analysis with income and education as explanatory variables.
% 
% <<12lissyoung>>=
% library(eha)
% library(xtable)
% ##load("data/loTab.rda")
% ##loTab$period <- factor(loTab$period)
% liss4064 <- loTab[loTab$age < 64.5, ]
% @
% 
% 
% The statistical significance of the covariates are shown in Table~\ref{tab:12liswom} for women and Table~\ref{tab:12lismen} for men. The variables of interest here are all significant.
% <<12liswom, results = 'asis'>>=
% fitw1 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss4064[liss4064$sex == "female", ])
% ##cat("Women:\n")
% drw <- drop1(fitw1, test = "Chisq")
% xtable(drw, caption = "Effect of covariates for women 40-64, 1990-2005, cancer death.", label = "tab:12liswom")
% ##round(summary(fitw1)$coef[-(1:13), ], 4)
% @
% 
% <<12lismen, results='asis'>>=
% fitm1 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss4064[liss4064$sex == "male", ])
% ##cat("\nMen:\n")
% drm <- drop1(fitm1, test = "Chisq")
% xtable(drm, caption = "Effect of covariates for men 40-64, 1990-2005, cancer death.", label = "tab:12lismen")
% ##round(summary(fitm1)$coef[-(1:13), ], 4)
% dcol <- c("blue", "darkgreen", "darkgreen", "red")
% @
% 
% The estimated effect sizes are shown in Figure \ref{fig:12liswomfig}.
% 
% <<12liswomfig, fig.cap = "Effect sizes, women 40-64, 1990-2005, cancer death.", fig.height = 4>>=
% source("R/plotWEduc.R")
% plotWEduc(fitw1)
% @
% 
% 
% And for men:
% 
% 
% <<12lismanfig, fig.cap = "Effect sizes, men 40-64, 1990-2005, cancer death.", fig.height = 4>>=
% #source("R/plotWEduc.R")
% plotWEduc(fitm1)
% @
% 
% 
% <<12plotliss1, fig.cap = "Cumulative hazards, ages 40--64, 2002--2013, cancer death.", fig.height = 4>>=
% #oldpar <- par(mfrow = c(1, 2))
% #plot(fitw1, fn = "cum", col = dcol, main = "Women")
% #plot(fitm1, fn = "cum", col = dcol, main = "Men")
% #par(oldpar)
% ##summary(fitw1)
% ##summary(fitm1)
% @
% 
% %See Table~\ref{tab:tablisswom1} the PH regression for women.
% <<12tablisswom1, results = "asis">>=
% #fitwo1 <- coxreg(Surv(enter, exit, event) ~ DISPINK + CIV + SUN2000NIVA,
% #                 data = liss4064[liss4064$sex == "female", ])
% #dr <- drop1(fitwo1, test = "Chisq")
% #ltx(fitwo1, dr = dr, caption = "PH regression, women aged 40--64, 2002--2013.",
% #    label = "tab:tablisswom1")
% @
% 
% %See Table~\ref{tab:tablissman1} the PH regression for men.
% <<12tablissman1, results = "asis">>=
% #fitma1 <- coxreg(Surv(enter, exit, event) ~ DISPINK + CIV + SUN2000NIVA,
% #                 data = liss4064[liss4064$sex == "male", ])
% #dr <- drop1(fitma1, test = "Chisq")
% #ltx(fitma1, dr = dr, caption = "PH regression, men aged 40--64, 2002--2013.",
%  #   label = "tab:tablissman1")
% @
% 
% When it comes to the association with education and income, we find the similar pattern as those for all-cause mortality and cardiovascular mortality. There is a strong effect for both men and women of having more education and income. These variables seem to catch up important aspect of socio-economic differences in health and mortality.

\subsection{Ages 65--89}

%The distribution over time periods.
<<12getske6089>>=
cvb6589 <- cvbTab[as.numeric(as.character(cvbTab$age)) >= 64.5, ]
cvb6589$age <- factor(cvb6589$age) # NOTE:: Now change to '6589'!
@

We finally turn to cancer mortality among the elderly. In accordance with the analyses for the age group 40--64, the model adjusts for {\tt civil status} and with separate analyses for women and men.

%\subsubsection{Women}

<<12hazcalwold>>=
resw <- getPHglm2(cvb6589[cvb6589$sex == "female", ],
              nper, labb)
resm <- getPHglm2(cvb6589[cvb6589$sex == "male", ],
                nper, labb)
@

We use the value of the cumulative hazards at
duration 25 (age 90) as a summary measure, starting from age 65.

The relation between different classes and cancer mortality fluctuates considerably both for women and men in this age group. For some periods it has to do with quite few events to analyse, particularly for the 1960s where HISCLASS has been difficult to determine for the elderly. Women of higher classes sometimes have favourable mortality, but unfavourable at other times. It seems however that the higher classes have reached lower cancer levels compared to the working class from 2001 onwards. Men from the elite had higher mortality until the 1970s, but a social gradient has appeared thereafter. Now the elite with assumed more resources have lower mortality than other groups. Elderly of workers origin on the other side have higher cancer mortality during the last decades, sometimes substantially higher than the others.The pattern is however not completely clear.

<<12cumswold, fig.cap = cap22, fig.scap = scap22, fig.height = 10>>=
##source("R/plotED.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "H(90|65)",  main = "Women 65-89")
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "H(90|65)", main = "Men 65-89")
par(oldpar)
@

% \subsubsection*{The effect of income and education on cancer death, 1990--2005}
% 
% Thus, we find indications of social differences in the later decades, particularly between the elite and the workers. The alternative analysis for the last period is based on a separate analysis with
% income and education as explanatory variables.
% 
% <<12lissyoung2>>=
% library(eha)
% library(xtable)
% ##load("data/loTab.rda")
% ##loTab$period <- factor(loTab$period)
% liss6589 <- loTab[loTab$age > 64.5, ]
% @
% 
% 
% The statistical significance of the covariates are shown in Table~\ref{tab:12liswom2} for women and Table~\ref{tab:12lismen2} for men.
% <<12liswom2, results = 'asis'>>=
% fitw2 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss6589[liss6589$sex == "female", ])
% ##cat("Women:\n")
% drw2 <- drop1(fitw2, test = "Chisq")
% xtable(drw2, caption = "Effect of covariates for women 65-89, 1990-2005, cancer death.", label = "tab:12liswom2")
% ##round(summary(fitw1)$coef[-(1:13), ], 4)
% @
% 
% <<12lismen2, results='asis'>>=
% fitm2 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = poisson, data = liss6589[liss6589$sex == "male", ])
% ##cat("\nMen:\n")
% drm2 <- drop1(fitm2, test = "Chisq")
% xtable(drm2, caption = "Effect of covariates for men 65-89, 1990-2005, cancer death.", label = "tab:12lismen2")
% ##round(summary(fitm1)$coef[-(1:13), ], 4)
% dcol <- c("blue", "darkgreen", "darkgreen", "red")
% @
% 
% The estimated effect sizes are shown in Figure \ref{fig:12liswomfig2}
% 
% <<12liswomfig2, fig.cap = "Effect sizes, women 65-89, 1990-2005, cancer death.", fig.height = 4>>=
% #source("R/plotWEduc.R")
% plotWEduc(fitw2)
% @
% 
% 
% And for men:
% 
% %% <<12lismen2, results='asis'>>=
% %% fitm2 <- glm(event ~ offset(log(exposure)) + as.factor(age) + urban + period + civst + income + educ, family = ##poisson, data = liss6589[liss6589$sex == "male", ])
% %% ##cat("\nMen:\n")
% %% drm2 <- drop1(fitm2, test = "Chisq")
% %% xtable(drm2, caption = "Effect of covariates for men 65-89, 1990-2005, cancer death.")
% %% round(summary(fitm1)$coef[-(1:13), ], 4)
% %% dcol <- c("blue", "darkgreen", "darkgreen", "red")
% %% @
% 
% <<12lismanfig2, fig.cap = "Effect sizes, men 65-89, 1990-2005, cancer death.", fig.height = 4>>=
% #source("R/plotWEduc.R")
% plotWEduc(fitm2)
% @
% 
% The social inequality in cancer mortality is confirmed by the approach of analysing the effect of education and income. We find the same consistent pattern as we have found for all-cause mortality and cardiovascular mortality for this period. Higher levels of education and more income is associated with lower cancer mortality. We can therefore conclude that both education and income are strongly decisive aspects of social differentiation when it comes to mortality during the last decades.

%%{\huge Här ska 'conclusion.tex' in !!!}

\section{Discussion}

The WHO Commission on Social Determinants of Health framework for the social determinants of health discusses how socio-economic position impacts on health. The influences are not direct but works through intermediate determinants related to exposure and vulnerability \citep{who10}. People are placed within socio-economic and political contexts, as well as value systems in the society. The socio-economic position imply different access to resources, power and prestige, all aspects of crucial importance. The framework takes a holistic approach to the social determinants, including not only the effect of pure economic and material resources. The importance of understanding people in their context has also been emphasized by for example \citet{kri05}. What she calls embodiment refers to the recognition that human beings are simultaneously social beings and biological organisms.

The authors of this paper have scrutinised different aspects of the social
pattern of mortality in northern Sweden for quite some time, focusing on
different regions, different age groups and partly different social
categories. Already the earliest findings, concerning the social inequality
in a rapidly growing town in an industrial district, surprised and at least
at first confused the author \citep{edv92}. The expected social gradient in
survival, meaning that people from higher social classes always had an
advantage, was in many cases, particularly among adult men, not found
despite the well-documented socio-economic differences in this
society. There were however large differences between men and women and for
different age groups. Further studies have confirmed these
findings \citep{edli11, edbr12}. These earlier studies have ended with an
assumption that distinct and consistent differences to the advantage of higher
classes developed somewhere during and probably quite early in the 20th
century. Until now we have however not been able to investigate the
continuous development on mortality differences among the adult and elderly
population due to no available data. This is the first time it has been
possible, and the results continue to surprise us. The consistent advantage
for higher social classes do not appear until quite late in the regions we
have analysed. The social pattern of female mortality are closer to what is
expected, but a social gradient in male adult mortality seems to be a
recent phenomenon. Furthermore, this transition and the class pattern of mortality differ depending on age group. Inequality in mortality appear clearer for the population at working age, while the differences are smaller among the elderly, a somewhat modified result from what we found in the first and preliminary analyses presented in Leuven this year \citep{segb16b}.

We have also added analyses of cause-specific mortality to the previous version. Mortality from cardiovascular diseases contributes to a large proportion of all deaths during the 20th century, and we are not surprised that it has a fairly similar pattern as for all-cause mortality. What is surprising is however the small class differences among the elderly. Cancer is sometimes considered less connected to social class, and we find mixed results in our analyses. There are however indications of a social gradient during the last decades in both studied age groups. Development of death causes are however problematic since the ways they have been diagnosed and the proportion with missing or vague information has changed over time.

The perhaps surprising results makes it reasonable to consider their
validity and possible weaknesses in the study. There are of course
uncertainties involved in all studies of this sort, something that should
not be ignored.  One obstacle concerns the quality of the available
information in the sources. Occupational titles in historical sources are
not always distinct enough to define the correct class. It can be difficult
to identify the skill level in the working class population because many
were often only designated as labourer. The size of the enterprise for free
entrepreneurs is unknown and it may be impossible to distinguish if they
are self-employed or have many employed. The class system and the position
of different classes can also be discussed. Studies covering long period
have the problem that occupations and classes change character - they do
not represent the same thing. Nonetheless, these ambiguities and possible
mis-classifications are not of that sort and magnitude that it would radically change
the basic patterns observed. The social classes identified during the
different periods represent distinct groups with very large differences in
resources, capacities and power. A worker would not end up in the highest
SOCPO level and large-scale entrepreneurs and higher officials would
certainly be assigned the right level.Available data also show that the
differences in income and wealth were large, often enormous.

In this paper we have analysed social inequality in health from the
perspective of social classes using HISCLASS. However, there are
alternative ways to analyse social inequality, and we cannot exclude that
such analyses gives other results or at least would modify the image
presented here. There are other alternatives then HISCLASS, representing
status or class dimensions. SOCPO is another system based on hierarchic
categories of social power, which we have used in a couple of previous
studies. In many modern studies categorisations based on income, wealth
or education are common for studies of inequality. It is possible that
such categorisations are better fit to distinguish between groups in
society. They represent different dimensions of inequality, and we believe
that the topic discussed here would gain from analyses of different aspects
of inequality. One problem is however that reliable data on income and
education are rarely available in historical population databases. We can
however be certain that the highest HISCLASSES level have more education
and the lowest the least. We also know from the 1930 census that the
economic differences were very large at that time and that occupations in
higher HISCLASSES tend to have higher incomes. The groups that may diverge
are those within agriculture and those having a small enterprise where we
can assume that the reported income figures underestimate their economic
standard.

As a test, we have however included analyses of the associations of
education and income to mortality for the period 1990--2005. For women as
well as for men, for both age groups and for all-cause mortality as well as
cause-specific mortality we find consistent differences with higher
educational level and more income leading to lower mortality. As already
mentioned, the class analysis proved that it in most cases are strong
differences during the last centuries, but the pattern is much clearer in
the alternative ways of categorisation. This does not mean that a class
analysis is worthless, rather that the different categorisations measure
different things. But it certainly point to the importance to also consider
other aspect of social inequality in the past. We do not know if education
and income are equally important in the past and that these measures would
provide us with a completely different picture, but this is something that
calls for further investigation.

The results of small socio-economic gradient in the past and its late development are however in accordance with other studies \citep{bevp11},
for example by \citet{bedr11} in a very different setting, a couple of
Scanian parishes. They found the first indications of an emerging social
gradient in the 1950s.The similarities makes us more comfortable about our
results.

\citet{who10} lists three different types of intermediary social
determinants of health, material circumstances, behaviours and biological
factors, and finally psycho-social factors. Most studies have focused on
the material circumstances. Having a better economy can give advantages in
many aspects of relevance for health and mortality, for example good living
conditions, access to food, medical care, better conditions at work et
cetera. Even though these advantages ought to have been at hand in our
studied regions, the final result was not better health for men in the most
privileged groups until the later part of the 20th century. We must
consider other intermediary determinants. Psycho-social factors have been
frequently discussed during the most recent decades, for example by
\citet{mar04}. He argues that being higher in the status ladder leads to
better health, regardless of where in the status ladder we look. This is
also one of the arguments \citet{wipi09} use as an explanation why equal
societies always perform better than unequal ones when it comes to health
and other social conditions. In the case studied here, we do not find any
indication of a status syndrome, since mortality levels was in the opposite
directions to the one we would expect from these theories.

This makes us consider the possible impact of behavioural factors. Our results of the gender aspect might be of interest in this case. When it comes to gender differences, the present study confirms what we have found on the Sundsvall region. There is a clear gendered pattern when it comes to social inequality in mortality. While high HISCLASS did not have any positive impact on survival for men in the early 20th century, we find  that higher social
positions among women led to longer lives. Within the same households, the effect of social
class seems to lead to different results. This opens up for behavioural
aspects based on different gendered expectations. Alcohol has been
suggested as part of the large gender differences in the 19th century
\citep{wil99, edv92}, and a more risk-taking life may be a component of the
male mortality.

Having a high social status could be expressed in attitudes and behaviours
that were harmful for health. Alcohol consumption decreased rapidly during
the period late 19th century to the 1950s, thus possibly making room for an
increased role of economic resources. But the old pattern in social health
inequalities remained. It is however possible that the relative differences
in alcohol consumption prevailed, although at a lower level. The question
is however if the lower levels of alcohol consumption had any substantial
negative impact on health. Swedes at that time had very low consumption
levels. Alternative explanations should also be scrutinise. There
may have been strong and heavy requirements of being successful for the
male \"elite. Many were involved in businesses with high risks. It would be
very damaging to lose the high position and they may have felt that they
lived under constant threat of economic failure. There are several examples among
the economic \"elite in the Sundsvall region that they had such problems
\citep{edli11}. A possible future study involving analyses of death causes
might throw some light on this issue.

Our study illustrates the importance of considering contextual factors and
the restrictions of making general statements. We need to look into all
aspects of social classes, not only the material ones but also what this
means for behaviour, norms, way of living. \citet{edli11} suggested that an
explanation for the larger social health inequalities in the present-day
world is that status and higher social position is now expressed through
health and being fit and physical active, while in many historical contexts
higher social status was expressed by low physical activity, obsessive
eating (being fat was impressive), drinking, smoking and sometimes also
taking risks.

Another question is if our results can be generalised. We have studied
environments in northern Sweden without any large metropolis. Perhaps
Sweden and/or this part of Sweden is a special case. The social differences
may have been smaller, the involvement of government both at the national
and local level may have been stronger, thus mitigating socio-economic
disadvantages. These are relevant questions, which hopefully can be
addressed in other and maybe comparative studies.

Finally, we argue that our results have implications for our understanding
on several hot research issues. A central question concerning the mortality
transition is how the enormous increase in survival could be
accomplished. This is still a hot debated question. \citet{mck76} argues that the
improvements up to the middle of the 20th century was first of all made
possible by improved nutrition. This seems to implicate that having more
resources would lead to better survival. If research shows that the most
advantaged groups did not have better health and longer lives, this do not
necessarily speak against McKeown's ideas but it makes the question more
complex. The results are also difficult to fit into \citet{liph95}
hypothesis that social conditions should be seen as a fundamental cause of
disease. Even if we would eliminate disadvantages such as bad living
conditions, lower standard of medical care and other proximate determinants
for the poor, the health disadvantage would remain. In the historical
context we have analysed here, this does not get any support. This further
illustrate the need for situating the analyses in their historical and
geographical contexts.

In recent years, the question about the relation between health and income
inequality been widely discussed. Even though our study do not address this
issue directly, our results are of interest for this question as well. The
development of the social inequality in mortality do not seem to fit that
well into the development of income and wealth inequalities. The early 20th
century, in Sweden as well as in other countries, was characterised by very
large economic differences, still this did in our regions not lead to any
consistent advantage for the highest social classes. After the Second World
War, economic differences diminished, and was at its lowest in Sweden at
about 1980, i.e. at a time when a more expected pattern of a social
hierarchy in survival had appeared. The period thereafter is characterised
by increasing economic differences, and this may coincide with increasing
health differences as well. The long-term relation between economic
inequality and health is worth further studies.

\section{Conclusion}

We started this paper by raising the issue of the development of social
inequalities. When studying this question in history we often get surprised
of its complexity. In a society characterised by large economic inequality,
we would expect that this would result in corresponding mortality
differences, i.e. higher social classes having an advantage. We find the
appearance of a consistent social pattern in mortality, to the advantage of
the higher classes, but surprisingly late in our history. It is only in the
recent decades that this pattern dominates. Improved survival characterised
however all HISCLASSes. We can however confirm that there are gender
differences in the social pattern. While there is no clear social gradient
among men, the results at least partly indicate advantages for the higher
social classes for women already early in the 20th century. Females had
more to gain from a high social position. Previous studies of the mortality
transition have made similar observations. Among adults, gender has often
been of more significant importance than social class. The reason for this
difference and for the late appearance of the modern social pattern among
men is still an enigma that require further studies, even though we have
suggested the possible importance of behavioural factors and changes in
ideals connected to different social classes. This should stimulate us to
further disentangle the pathways for how social position influence health
and mortality in different societies. Furthermore, we do not find any substantial differences or distinct pattern in the relation between class and the two major cause-of-death groups. Finally, we find stronger differences in the social pattern among the working-age population, while the associations were smaller or not so clear among the elderly. However, when analysing the association with education and income in the period 1990--2005 we find consistent patterns of higher income and higher education level being associated with lower mortality, indicating that these variables are better to catch socio-economic differences at least in the present-day society.

\section*{Acknowledgements}

This work was undertaken within the Centre for Demographic and Ageing
Research (CEDAR) at Umeå University and with support from the Swedish
Research Council (VR) (grant no. 421-2012-892) and the project Old age ... financed by the Swedish Research Council. We thank Demographic Data
Base for giving us access to data and Jonas Helgertz, Department of
Economic History, Lund University, for providing us with scripts for
creating HISCLASS from Swedish Census data.

\bibliography{mybib}

\end{document}
@