### 2017-06-07 Umea.Rnw
Revised slides

\documentclass[t]{beamer}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{natbib}

%\usepackage[figuresonly]{endfloat}

%\bibliographystyle{apalike}

\usepackage{color}

\newcommand{\emp}[1]{\textcolor{blue}{#1}}

\usetheme{Singapore}% best choice.
  \setbeamercovered{transparent}%

\newcommand{\R}{{\bf R}}
\newcommand{\code}[1]{{\tt #1}}

\title{Long--term perspectives of social inequality in mortality among adults and elderly in
  northern  Sweden 1851--2013\footnote{Session "Health and Mortality in Arctic Populations" at The Ninth International Congress on Arctic Social Sciences (ICASS IX) June 10, 2017, Umeå}
}

\date{\today}

\author{Sören Edvinsson and Göran Broström \\
Centre for Ageing and Demographic Research, Umeå University}

\begin{document}

\maketitle

<<setup, include = FALSE>>=
library(knitr)
opts_chunk$set(engine='R',dev='pdf',fig.path='figs/ume-',tidy=FALSE,fig.height=5.5,fig.width=8,strip.white=TRUE,cache=TRUE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
@

<<environ>>=
library(knitr)
library(eha)
library(xtable)
library(skum)
## Load remote data:
con <- url("http://capa.ddb.umu.se/data/vb0.rda")
##load("data/vb0.rda")
load(con)
close(con)
vb0$hisclass[is.na(vb0$hisclass)] <- 4
vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "middle", "worker", "NA"))
##
con <- url("http://capa.ddb.umu.se/data/vb.rda")
##load("data/vb.rda")
load(con)
close(con)
##
con <- url("http://capa.ddb.umu.se/data/lisaold.rda")
##load("data/lisaold.rda")
load(con)
close(con)
##
vb0Tab <- aggregate(vb0[, c("event", "exposure")],
                   vb0[, c("sex", "age", "period", "cohort",
                          "urban", "civst", "hisclass")],
                   FUN = sum)
vb0Tab$age <- factor(vb0Tab$age)

@

\begin{frame}{Aims}

To investigate the association between social class and mortality among adults and elderly from the 19th century to the present and how it is connected to socio-economic inequality in society.
This is analysed according to 
\begin{itemize}
\item gender
\item age group (working age vs retired)
\item cause of death (cardiovascular, cancer)
\item (education 1990--2013)
\item (income 1986-2005)

% \item Has \emp{inequality in mortality} between \emp{social classes}
%   increased in the adult and elderly
%   population?

%%\item Is the mortality transition initiated by a certain social class?

% \item Are there any \emp{gender differences} in the effect of social position?
% 
% \item Is social position \emp{equally important} among the \emp{retired} population as
%   for those in \emp{working age}?
% \item Is there a difference between classes in mortality from
%   \emp{cardiovascular} diseases and \emp{cancers} respectively?
% \item Are there large differences in survival depending on level of
%   \emp{education} and \emp{income} categories?

\end{itemize}

\end{frame}

\begin{frame}{Main results}

\begin{itemize}
\item A consistent mortality difference according to social class appears surprisingly late in our regions and has no clear connection to the development of income inequality.
\item However, there is a strong gender difference in the development of social inequality.
\item We argue that health-related behavior is an important aspect in historical as well as present-day mortality, and that such behaviours are based on strong gender and class expectations.
\end{itemize}

\end{frame}

\begin{frame}{Economic inequality}

\includegraphics[height=8cm]{myfigs/Top_incomes.jpg}

\end{frame}

\begin{frame}{Outline}

\begin{center}
\begin{enumerate}

\item \emp{Data} management
\item \emp{Modeling} considerations
\item \emp{Analysis}
\item \emp{Results}
\item \emp{Implications} of the results
\end{enumerate}

\end{center}

\end{frame}

\begin{frame}{Two regions in Västerbotten}

\begin{columns}

\begin{column}{0.5\textwidth}

\includegraphics[height=7cm]{myfigs/sweden.pdf}

\end{column}

\begin{column}{0.5\textwidth}

Remote area in northern Sweden, dominated by agriculture until well into the 20th century. 19th century communications were rudimentary.

Skellefteå:
\begin{itemize}

\item Small town, industrialization of the region during the 20th century.

\end{itemize}

Umeå:

\begin{itemize}

\item Center for administration, education, and military. University from 1965

\end{itemize}

% Rural areas:
% 
% \begin{itemize}
% 
% \item Very similar: Small farms dominating.
% 
% \end{itemize}

\end{column}
\end{columns}
%\includegraphics[height=7cm]{myfigs/sweden.pdf}

\end{frame}

\begin{frame}[fragile]{Data sources}

\begin{itemize}
\item \emp{Swedish church book} records (1851--1950).
  \begin{itemize}
\item digitized by the \emp{Demographic Data Base} at Umeå University,
\begin{itemize}
  % \item \emp{\sc Popum} and
   \item \emp{\sc Poplink}
\end{itemize}
\item  \url{http://www.cedar.umu.se/}
\end{itemize}
\item The \emp{Linnaeus} data base (1960--2013)
  \begin{itemize}
\item Census data 1960, 1965, \ldots, 1990
\item Population registers ({\sc Lisa}, 1986--2013)
  \begin{itemize}
    \item \emp{Income}
    \item \emp{Education}
  \end{itemize}
 \item Death information from \emp{National Board of Health and Welfare}
   \begin{itemize}
\item Death dates from 1 January 1961 to 31 December 2013.
  \item \emp{Causes of death} ({\sc Ulorsak}, ``main cause of death'')
   \end{itemize}
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}{The sampling frame}

<<lexis2>>=
plot(c(1851, 1851, 1951, 1951, 1851), c(40, 90, 90, 40, 40),
     type = "l", xlim = c(1851, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 90))
axis(1, at = c(1851, 1901, 1951, 1961, 1981, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1961, 1961, 2014, 2014, 1961), c(40, 90, 90, 40, 40),
      col = "blue", lwd = 2)
#
n.rec <- NROW(vb0[vb0$dataset == "poplink", ] )
n.pl <- with(vb0[vb0$dataset == "poplink", ], length(unique(id)))
n.d <-  with(vb0[vb0$dataset == "poplink", ], sum(event))
text(1900, 70, "PopLink", cex = 2, col = "red")
text(1899, 60, paste(n.rec, "records"))
text(1900, 55, paste(n.pl, "persons"))
text(1899.5, 50, paste(n.d, "deaths"))
#
n.rec <- NROW(vb0[vb0$dataset != "poplink", ] )
n.pl <- with(vb0[vb0$dataset != "poplink", ], length(unique(id)))
n.d <-  with(vb0[vb0$dataset != "poplink", ], sum(event))

text(1985, 70, "Linnaeus", cex = 2, col = "blue")
text(1984, 60, paste(n.rec, "records"))
text(1985.2, 55, paste(n.pl, "persons"))
text(1985.2, 50, paste(n.d, "deaths"))
@
\end{frame}

\begin{frame}{Variables}

\begin{description}

\item[sex]
\item[birth date] By quarter (15 Feb, May, Aug, Nov)
\item[death date] By day (\emp{Socialstyrelsen})
\item[cause of death] from \emp{Socialstyrelsen}
\item[parish] Urban/Rural
\item[Social class] elite, middle, worker. Household definition 1851--1950. Individual definition 1960--2013
\item[income] 1990--2005, grouped by quartiles.
\item[education] 1990--2005, three categories.
\item[period]  1851--1875, \ldots,  2008--2013 (10 periods).
%\item[cohort] 1761--1771, \ldots, 1971--1974 (22 birth cohorts).

\end{description}

\end{frame}

\begin{frame}{\sc Social class}

\begin{enumerate}
\item \emp{\tt elite}, HISCLASS 1 and 2. Higher managers and professionals.
\item \emp{\tt middle}, HISCLASS 3, 4, 5, 6, and 8. Lower managers, lower white collar, and farmers.
%%\item \emp{\tt lowWhiteC}, HISCLASS 4, 5 and 6. Lower white collar.
\item \emp{\tt worker}, HISCLASS 7, 9, 10, 11 and 12. Workers of different skills including farm workers.
\end{enumerate}

\end{frame}

\begin{frame}{No proportional hazards (wrt class)!}

\begin{itemize}
\item Fit each social class \emp{separately}, for each time period.
\item Use the \emp{cumulative hazard} at 65 (or 90) for comparison.
\end{itemize}

So, we fit separate PH models for each combination of

\begin{itemize}
  \item \emp{\tt sex}
  \item \emp{\tt period}: 1851--75, 1876--1900, 1901--10, etc.
  \item \emp{\tt class}
  \item \emp{\tt age group}: 40--64, 65--89.
  \item \emp{\tt cause of death}
\end{itemize}
\end{frame}

\begin{frame}{Women and men, age 40--64}

%  \begin{center}
%\includegraphics[height = 7.5cm]{figure/cumsw-1}
%  \end{center}

<<hazcalw>>=
source("R/getPHglm2.R")
##nper <- 4
vb4064 <- vb0Tab[as.numeric(as.character(vb0Tab$age)) < 64.5, ]
vb4064$age <- factor(vb4064$age)
labb <- levels(vb0Tab$period)
nper <- length(labb)
resw <- getPHglm2(vb4064[vb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb4064[vb4064$sex == "male", ],
            nper, labb)
@

<<cumsw>>=
source("R/plotED.glm.R")
scol <- c("blue", "darkgreen", "red", "black")
oldpar <- par(mfrow = c(2, 1), cex = 0.7, lwd = 1.5)
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, ylab = "TMR(65|40)",
                main = "Women 40-64", exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, ylab = "TMR(65|40)",
                main = "Men 40-64")##, exclude = 5:6)
par(oldpar)
@

\end{frame}

\begin{frame}{Relative differences, ages 40--64 and 65--89}

  %\begin{center}
\includegraphics[height=7.5cm,width=5.4cm]{figs/slide-cumswstand-1}\includegraphics[height=7.5cm,width=5.4cm]{figs/slide-cumswold-1}
   %\end{center}
\end{frame}

\begin{frame}[fragile]{Education 1990--2013}

<<educ>>=

educ <- lisaold[!(lisaold$period %in% c("1986-1987", "1988-1989")), ]
educ$period <- factor(educ$period)
educ <- educ[!is.na(educ$educ), ]
levels(educ$educ) <- c("low", "low", "middle", "middle", "middle", "high", "high")
educ$educ <- relevel(educ$educ, ref = "middle")
educ$civst <- relevel(educ$civst, ref = "married")
fitm <- coxph(Surv(enter, exit, event) ~ strata(educ) + civst + urban,
             data = educ[educ$sex == "male", ])
fitf <- coxph(Surv(enter, exit, event) ~ strata(educ) + civst + urban,
             data = educ[educ$sex == "female", ])
cols <- c("darkgreen", "red", "black")
@

<<ploteduc>>=
oldpar <- par(mfrow = c(1, 2))

plot(survfit(fitm, start.time = 40), mark.time = FALSE, main = "Men",
     col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
axis(1)
axis(2, at = c(0, 0.5, 1.0), las = 1)
abline(h = 0.5, lty = 3)
box()
legend("bottomleft", legend = levels(educ$educ), col = cols, lty = 1, lwd = 1.5)

plot(survfit(fitf, start.time = 40), mark.time = FALSE, main = "Women",
     col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
axis(1)
axis(2, at = c(0, 0.5, 1.0), las = 1)
abline(h = 0.5, lty = 3)
box()
legend("bottomleft", legend = levels(educ$educ), col = cols, lty = 1, lwd = 1.5)

par(oldpar)
@

\end{frame}

\begin{frame}{Income 1986--2005}

<<femeduc>>=
income <- lisaold[!is.na(lisaold$income), ]
income$period <- factor(income$period)
##income <- income[!is.na(income$income), ]
##levels(income$income) <- c("low", "low", "middle", "middle", "middle", "high", "high")
income$income <- relevel(income$income, ref = "low")
income$civst <- relevel(income$civst, ref = "married")
fitm <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
             data = income[income$sex == "male", ])
fitf <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
             data = income[income$sex == "female", ])
cols <- c("darkgreen", "red", "blue", "black")

@

<<plotincome>>=
oldpar <- par(mfrow = c(1, 2))

plot(survfit(fitm, start.time = 40), mark.time = FALSE, main = "Men",
     col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
axis(1)
axis(2, at = c(0, 0.5, 1.0), las = 1)
abline(h = 0.5, lty = 3)
box()
legend("bottomleft", legend = levels(income$income), col = cols, lty = 1, lwd = 1.5)

plot(survfit(fitf, start.time = 40), mark.time = FALSE, main = "Women",
     col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
axis(1)
axis(2, at = c(0, 0.5, 1.0), las = 1)
abline(h = 0.5, lty = 3)
box()
legend("bottomleft", legend = levels(income$income), col = cols, lty = 1, lwd = 1.5)

par(oldpar)
@

\end{frame}

%%% Stop here!

\begin{frame}{Summarizing the results}

%\scriptsize

\begin{itemize}
\item \emp{Improved} survival characterised all {\sc Hisclass}es.
\item \emp{Late} appearance of the “modern” pattern of social differences in mortality among men
\item Distinct \emp{gender} differences in the development.
\item The \emp{upper} class(es) gradually take over as the \emp{most healthy} group.
\item No \emp{qualitative} but partly \emp{quantitative} differences between those in working age compared to the retired population.
\item \emp{Educational} and \emp{income} differences during the last decades conform to what we found for class differences.
\item In the long run, no clear connection between income inequality and social differences in mortality.
\end{itemize}

\end{frame}

\begin{frame}{Some implications}

%\scriptsize

\begin{itemize}
\item The development of social health inequalities and the role of \emp{health-related behavior} in different phases of the transition (Antonovsky and Omran).
\item Inequality as a \emp{contextual} health risk (Wilkinson/Pickett).
\item The role of \emp{living standard} for mortality decline (McKeown).
\item \emp{Gender} and mortality (Watkins, Mckinnon, Janssens).
\item Social conditions as \emp{fundamental causes of death} (Link/Phelan).

\end{itemize}

\end{frame}

%\begin{frame}{Conclusion}

%\scriptsize

% \begin{itemize}
% \item \emp{Improved} survival characterised all {\sc Hisclass}es.
% \item There are \emp{gender differences} in the social pattern
% \item No difference in the social pattern in the two studied age groups.
% \item The \emp{upper} class(es) gradually take over as the \emp{most healthy} group.
% \item \emp{Educational} and \emp{income} income differences during the last decades conform with what we found for class differences.
% \end{itemize}

%\end{frame}

\end{document}