### 2017-07-29 article.Rmd
Revising results and references

---
title: "Is high social class always beneficial for survival? Long-term perspectives of social inequality in mortality among adults and elderly in Northern Sweden 1851--2013""
author: Sören Edvinsson and Göran Broström
date: "24 July 2017"
output:
  bookdown::pdf_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
link-citations: yes
bibliography: mybib.bib
biblio-style: apalike
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(engine='R',fig.path='figs/art-',tidy=FALSE,fig.height=3.5,fig.width=8,strip.white=TRUE,cache=TRUE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
library(knitr)
library(eha)
library(xtable)
library(skum)
```

```{r readalldata}
source("R/get.file.R")
vb0 <- get.file("vb0")

vb0$hisclass[is.na(vb0$hisclass)] <- 4
vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "middle", "worker", "NA"))
##

## We don't need "vb", do we?
## vb <- get.file("vb")

lisaold <- get.file("lisaold")

loTab <- get.file("loTab")

loTab$period <- factor(loTab$period)
##
vb0Tab <- aggregate(vb0[, c("event", "exposure")],
                   vb0[, c("sex", "age", "period", "cohort",
                          "urban", "civst", "hisclass")],
                   FUN = sum)
vb0Tab$age <- factor(vb0Tab$age)
```

# Introduction

In the present paper, we take a long-term perspective on the development of
social inequalities in mortality in the adult and elderly population. The areas under investigation are the Skellefteå and Umeå regions in the north of Sweden, and the time period is 1851--2013 for Skellefteå and 1901--2013 for Umeå. This issue is then analysed according to gender and age-group (working age vs retired) and major causes of death (cardiovascular diseases and cancers). Furthermore we put this in he context of how the inequality in mortality is associated
with the development of economic inequality in society.The results are then discussed in relation to our understanding of the mortality transition and the social determinants of health and mortality, and their implications on some of the most influential hypotheses and concepts in health research.

On the basis of these results, we argue that a consistent hierarchical social gradient in mortality appears surprisingly late, but that there were large gender differences in trends and levels. We suggest that more attention should be paid on health-related behaviour also in earlier phases of the mortality transition.

One of the central aspects of survival is social class and access to economic
and other resources. Even in present-day welfare societies, social position is a
strong determinant when it comes to health and mortality and the impact even
seems to be increasing [@kuetal04, @macketal16, @frlu07, @broba12, @strand10]. It has been suggested that
"... social conditions have been, are and will continue to be irreducible
determinants of health outcomes and therefore deserves their appellation as
'fundamental causes' of disease and death" [@liph95]. For a long time,
the general view has been that socio-economic health and mortality inequalities were high in
historical societies, probably higher than in modern societies. This is a
reasonable assumption since these societies were in most cases characterised by
very large differences between groups and classes. Knowing that access to
resources provides advantages in all aspects of life, the health advantage of higher
classes ought to be obvious. @ant67 and @smi83 suggest that social
inequality in mortality has passed through different historical phases. According
to them, differences were comparatively small during the
pre-transitional phase when space was a strong determinant for the spread of
disease. Differences then increased during the transitional phase when mortality
declined and wealthy groups used their economic and human resources to gain better health. Finally,
mortality differences decreased again resulting in marginal levels of inequality in modern low-mortality societies when instead
health-related behaviour became the decisive determinant for health and survival.
@omran82 comes to a similar conclusion states, in the third proposition in his theory of the
epidemiological transition, that even if the class differential in mortality was
maintained during the transition, the decline set in earlier and was faster
among privileged groups. These assumption if partly supported by data from England and Wales [@pamuk85].

Recent studies, investigating social inequality in health and mortality
with micro-data, has however questioned how generalizable the assumed
pattern is [@bevp11]. Instead, some studies suggest that the modern
pattern might be a fairly recent development [@bedr11], although the pattern differs
depending on age, gender or historical context. When it comes to the adult
and elderly population results from different studies have diverged, some
reporting substantial differences but many, particularly those based on
Swedish data, have found surprisingly small or even reversed socio-economic
differences during the 19th and early 20th century [@bedr11,
@edli11,@edbr12], sharply contrasting from the situation in present-day society.

It thus appears as if a more consistent social health divide appeared at
some point during the mortality transition, at least in Sweden. Solid
empirical evidence about the process is however lacking and studies
focusing on the issue are still few. There is a need for more reliable
studies from different geographical and historical settings in order to 
better understand the role of socio-economic conditions for health and survival over time. 
Furthermore, previous studies indicate substantial differences in the development according to age and gender.

Another aspect of the development of social inequality in mortality,
concerns its relation to economic inequality. @wipi09 argue that
income inequality has an independent effect on mortality, separate from the
direct effect of actual access to economic resources. They argue
that unequal societies perform less well when it comes to health than equal
ones. Wilkinson and Pickett find a positive association between inequality
and health (as well as other social conditions) at country level in the
present-day economically developed world. This has initiated a vital
scholarly debate and the topic has been extensively studied [@suka04,
@wado00]. How the association looked like in previous periods is
unknown. This is of particular interest since the levels of economic
inequality were very different from those of the recent decades. Even if it
not a necessary implication, it is reasonable to assume that poorer groups were
those most disadvantaged of large inequality in itself. During the last decades, the association between trends of inequalities in mortality and income respectively are weak or non-existant according to @hoetal16.

When it comes to Sweden both income and wealth distribution was strongly skewed from the 1870's to the early 20th century. Starting during the inter-war period, income and
wealth inequality continuously diminished to reach a low point at around
1980. Thereafter, economic inequality has increased substantially until the
present, although far from being as high as a century ago [@rowa08,
@rowa09]. The Swedish development resembles that of other European and North
American countries in its basic features [@rowa14].

# The Skellefteå and Umeå regions

```{r mapp,fig.cap = "The Skellefteå and Umeå regions in Sweden.", out.width="3in"}
include_graphics("myfigs/sweden.png")
```
The Skellefteå and Umeå regions (Figure \@ref(fig:mapp)) are part of the
county of Västerbotten in the north of Sweden along the coast of Gulf
of Bothnia. This was a remote part of the country, where communication with
the rest of Sweden was difficult until the late 19th century. The economy
was dominated by agriculture, making it vulnerable to harvest failures, and
several severe famines affected the regions during the 1800s. One living
long in memory was the famine of 1867–-1868 [@segb14]. During the long
winters, sea communication was hindered due to the Gulf of Bothnia being
frozen. The first ships during the year came to this area in May or even as
late as June in some cases [@fah56]. Towards the end of the 19th
century, the Swedish railway system reached this part of Sweden, thereby
facilitating contacts with the rest of Sweden, improving the economy and
mitigating the effects of harvest failures. The inland was particularly
remote and isolated. During the 19th century, our regions became
increasingly integrated in the same epidemiological pattern as the rest of
Sweden. Diseases reached to places that were spared in earlier times.

The Skellefteå region until the 1950 in our dataset consists of a selection
of parishes surrounding the town of Skellefteå, founded in 1845 but with a
very small population during the 19th century. The data from the period
after 1960 cover the Skellefteå and Malå municipalities, the same area as
for the earlier period but with the addition of two more parishes. The
majority of the 19th century population lived in rural villages and
hamlets, getting their livelihood from agricultural production. During the
20th century, industrialisation took place, partly by the opening of mines
in the in-lands. This also led to a population increase both in the town
and in the rural parts, and a much more diversified economy.

Mortality was fairly low in comparison with other parts of the country and
the fertility transition was late [@coa86]. The Skellefteå population
size as defined in our data sets was 16\,473 on January 1, 1850, 41\,352 on
January 1, 1900, 61\,938 on January 1, 1950 and 76\,723 at the end of the
century.

The Umeå region in this study consists of Umeå urban and rural parish until
1950, and from 1960 onwards of Umeå municipality when another three
parishes are included. This region has a somewhat different character from
that of Skellefteå. Umeå town had a small population but was substantially
larger than Skellefteå town during almost the whole studied period. It was
the administrative centre in the county of Västerbotten. Schools and
military regiments were placed here and the economy was more diversified
compared to Skellefteå. Agriculture dominated the rural part, but there
were also some foundries as well as industries for example in forestry and
small-scale production. The population size as defined in our data sets was
18\,970 on January 1, 1900, 32\,900 on January 1, 1950 and 103\,970 when
the century ended.

# Data and variables

The recent extension in time of data at the Demographic Data Base (DDB),
Umeå University (www.cedar.umu.se) by including the period between 1900 and
1950s makes it possible to investigate the long-term development of social
inequality in mortality, in the present case from 1851 to 2013. The
databases provide us with micro-data on a sufficiently large population and
with substantial social diversity for the Skellefteå and Umeå regions
[@wee16]. The data for the present study comes from two large
population databases at the DDB. The early period until 1950 is covered by
the database Poplink, the digitisation of historical parish registers for
the Skellefteå and Umeå region [@wee16]. Poplink is based on linked
parish records, allowing us to reconstruct life biographies on people as
long as they remained in the region. The records are linked within but not
between the regions. Data from Poplink are accessed for the period 1851--1950.

The other large data-set is extracted from the Linnaeus database
[@malm10]. Linnaeus is based on different linked national population
registers from 1960 to 2013 and is used within the ageing program at CEDAR, Umeå
University. The study period from 1960 to 2000 are constructed from
censuses every fifth year 1960--1990, with additional information on deaths
from National Board of Health and Welfare. For the period 2002--2013 we use
the information from the yearly population registers from Statistics Sweden
(LISA) together with death information from National Board of Health and
Welfare.

Information on individuals are anonymised and the two databases are not
linked. This stops us from following individuals present
in both databases throughout their lives, thus they are treated separately. This
also makes it impossible to add information on individuals in the Linnaeus
database from what we potentially could find in Poplink, for example family
background or occupations earlier in life.

```{r lexis2,fig.cap="The sampling frame.", fig.height = 5}
plot(c(1851, 1851, 1951, 1951, 1851), c(40, 90, 90, 40, 40),
     type = "l", xlim = c(1851, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 90))
axis(1, at = c(1851, 1901, 1951, 1961, 1981, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1961, 1961, 2014, 2014, 1961), c(40, 90, 90, 40, 40),
      col = "blue", lwd = 2)
#
pn.rec <- NROW(vb0[vb0$dataset == "poplink", ] )
pn.pl <- with(vb0[vb0$dataset == "poplink", ], length(unique(id)))
pn.d <-  with(vb0[vb0$dataset == "poplink", ], sum(event))
pn.d <- as.integer(pn.d)
text(1900, 70, "PopLink", cex = 2, col = "red")
text(1899, 60, paste(pn.rec, "records"))
text(1900, 55, paste(pn.pl, "persons"))
text(1899.5, 50, paste(pn.d, "deaths"))
#
n.rec <- NROW(vb0[vb0$dataset != "poplink", ] )
n.pl <- with(vb0[vb0$dataset != "poplink", ], length(unique(id)))
n.d <-  with(vb0[vb0$dataset != "poplink", ], sum(event))
n.d <- as.integer(n.d)

text(1985, 70, "Linnaeus", cex = 2, col = "blue")
text(1984, 60, paste(n.rec, "records"))
text(1985.2, 55, paste(n.pl, "persons"))
text(1985.2, 50, paste(n.d, "deaths"))
```

In the data-set analysed here, all individuals 40 years and older ever
being resident in either of the regions are included. The data file
contains the variables on social class (Hisclass) sex, urban/rural
residence, birth date, death date, first and last date of observation and
type of entrance/exit. Total number of persons are `r pn.pl` leading to
`r pn.d` deaths in Poplink and `r n.pl` persons leading 
to `r n.d` deaths in the Linnaeus database. See 
Figure \@ref(fig:lexis2).

## Presence periods

Differences in available information in the two datasets as well as within
the Linnaeus data makes it necessary to apply different approaches when it
comes to identification of presence periods within our
regions. The Poplink data provides us with exact dates or at least year of
start and exit of the presence of all individuals in our data, allowing us
to have full and continuous control of the de jure population. This is
however not the case with the Linnaeus database. For the period 1960--2000
we use presence in the censuses 1960--1990 together with the yearly
population registers 1986--2000 and information on deaths from the National
Board of Health and Welfare to identify the observation periods in our
analyses. There is however no full information on what happened in-between
the censuses, for example the exact date of exit for persons that moved
between two censuses. We have solved this by considering everyone as
resident in our regions during the complete five-year period between
censuses if the person is missing in the later census (the last period covering ten years). This overestimates
their presence in our regions, but should not bias the results when it
comes to socioeconomic differences. We furthermore miss those that both
moved in and moved out or vice versa between censuses. Exit dates are exact
if they refer to death, and in these cases we consider them as constantly
present from the last census to the death data even if the death took place
in another part of Sweden.

For the last period, 2002--2013, we use the yearly population registers
(LISA data), together with information on deaths from National Board of
Health and Welfare. This data allows for full control of the presence periods
in the studied regions.

## Social class

The choice of variable for analysing social differences vary between studies. A common alternative to a class analysis is to use income or education for
investigating socio-economic health differences. @goldk10 criticize
how many economists and epidemiologists operationalize inequality in their
studies, for example by not distinguishing between class and status
aspects. Class refers to relational aspects in society, while many other
variables focus on the distribution of material and/or immaterial
resources. He advocates studies of different aspects, which together
complement each other. Comparing different approaches are valuable in order
to understand why some are advantaged and others disadvantaged and what
sort of factors determines health differences.

In this paper we analyse mortality differences according to social class
based on the classification scheme Hisclass [@lema11]. The only
type of socio-economic information available both in the Linnaeus and Poplink database
is occupations. Thus, analysing according to social class is apart from
being a scientific choice, also a practical one. Using occupational data is
the only and obvious alternative for making a longitudinal study of
socio-economic mortality differences with consistent classification
schemes. Social position is treated as a time-dependent variable.

The availability to information on occupations varies over time and between
sources. Occupations in the version of Poplink we are using are coded according to the original DDB coding system that then has been adapted to Hisclass. The censuses 1960--1990 include occupational
codes based on the system used in Sweden at that time, NYK (Nordisk
Yrkesklassificering). We consider the observed social class in a census to be valid for the full period until the next census, corresponding to our treatment of presence periods. For the period 1991--2000 the social class in the census 1990 is considered to be kept
until 2001. This strategy is also used for cases of missing social
class. Furthermore, we consider the last social position before retirement
to be kept until death or last observation for those above 65. For the
Linnaeus data, we use a script recoding HISCO into Hisclass, provided
to us by Jonas Helgertz, Lund University.

The use of different data sets, forces us to apply different approaches
towards defining the social position. A problem is that
Poplink usually only provides the occupation for the head of household,
thus under-estimating female working-force participation as well as that of
adult children residing with their parents [@vik10]. In this study and
according to the research questions in the present paper, we have chosen to
categorise wives according to position of the head of household, usually
the husband/father, assuming that the family shares the same socio-economic
position. This strategy makes the problem of under-recording less
problematic.

Female labour is much better covered in the Linnaeus database. A
complication is however that it is  difficult to define households in the
same way as in Poplink. Extra-marital cohabitation, connected to the Second
Demographic Transition [@van87, @van03], became common and female
labour force participation developed as the norm in Sweden during this
period. The occupation of the husband as signifier of social class became,
if not obsolete, at least less relevant. All included persons are signified
by their individual occupation for this period. Still, the results for the
different periods are not completely comparable and makes the
interpretation of the development for women more uncertain. Notice however that there is a strong social homogeneity between partners in a marriage, indicating that the results would not differ in any decisive way if we use social class on household level also for the Linneaus data.

Hisclass, the classification system used here, is a "...
HISCO-based historical international social class scheme" [@lema11,
@lema02]. The different classes in Hisclass represent distinct
categories, based on whether the work is manual or non–manual, skill level,
supervision and sector. Implicitly it reflects large differences in what is
in the forefront of other categorisations, as for example access to
economic and other resources, status, social power et cetera. Longitudinal analyses of social classes for such a long period covered in this study involves several problems, related to changes in occupational structure and the meaning and contents of occupations and social positions. We have
merged the original 12 classes in Hisclass, based on position in the
labour market, ending up with the following three and an additional fourth
category with unknown class:

\begin{enumerate}
\item {\tt elite}, Hisclass 1 and 2. Higher managers, professionals
and higher white collar.
\item {\tt middle}, Hisclass 3, 4, 5, 6 and 8. Lower managers,
farmers and lower white collar.
\item {\tt worker}, Hisclass 7, 9, 10, 11 and 12. Workers including
farm workers.
\end{enumerate}

Figures \@ref(fig:hiscskumm) and \@ref(fig:hiscskumw) show the distribution
(per cent) of exposure time according to class for men and women
respectively. The category of workers is fairly stable over time. The
highest class has increased substantially, although from a very low level
in the 19th century. The increase of academics in connection to the
establishment of Umeå University explains much of this increase. What is
invisible in Figure \@ref(fig:hiscskumm) is the transfer within the middle
group from a majority of farmers during the 19th century to lower white
collar from the middle of the 20th century. Furthermore we observe an
increase during the 1960s and partly during later decades as well in those
with undefined class. When first observed in the earliest included censuses,
occupation is often missing. For men the proportion is somewhat higher in
the age group 65 and above, i.e. those that were retired, but still low
enough to be used in our analyses. Until 1950, there were however very few
with missing occupation. For men in working-active age, the proportion
missing does not cause any problems for our analyses.

```{r hiscskumm, fig.cap = "Exposure by hisclass, men (from the top: elite, middle, workers, missing) and time period.", fig.height = 4.5}
xx <- with(vb0Tab[vb0Tab$sex == "male", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
yy <- round(prop.table(xx, 1) * 100, 1)
barplot(t(yy)[4:1, ], col = c("red", "blue", "green", "black"), las = 2, ylab = "Per cent")
```

```{r hiscskumw, fig.cap = "Exposure by hisclass, women (from the top: elite, middle, workers, missing) and time period.", fig.height = 4.5}
xx <- with(vb0Tab[vb0Tab$sex == "female", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
yy <- round(prop.table(xx, 1) * 100, 1)
barplot(t(yy)[4:1, ], col = c("red", "blue", "green", "black"), las = 2, ylab = "Per cent")
```

The proportion with missing values among women is however a larger problem,
see Figure \@ref(fig:hiscskumw). Corresponding to what we found for men, the
proportion without social position is very low until the middle of the 20th
century. This is a reflection of how womens social position is defined for
that period. Missing social class is however a real problem during the
1960s and 1970s in particular. Around 80 \% lack information for the 1960s
and 50 \% for the 1970s. Since social class for this period is defined
according to own occupation, the consequence of many women being outside
the workforce, the proportion missing becomes very high. We have chosen not
to report results for women for these two decades. For the later decades,
the results are fairly complete.

## Marital status

Marital status is treated as a time-dependent variable in the data
file. The statuses used are unmarried, married and dissolved marriages
(without distinguishing between widowed and divorces). If no explicit
status is given in the sources and partner is missing, the status has been
set to unmarried.

Regarding the distribution of marital status over time, see
Figure \@ref(fig:civstskum). The figure indicates an under-registration of
those married and over-registration of unmarried during the
1960s.

```{r civstskum, fig.cap = "Exposure by civil status (from the top: never married, married and previously married) and time period", fig.height = 4.5}
xx <- with(vb0Tab, tapply(exposure, list(period = period, civst = civst), sum))
yy <- round(prop.table(xx, 1) * 100, 1)
barplot(t(yy)[3:1, ], col = c("red", "blue", "black"), las = 2, ylab = "Per cent")
```

## Cause of death

In the present study, we analyse mortality in the two most common disease
groups in the ages we are investigating---cardiovascular diseases and
cancers. National compulsory reporting of cause of death was introduced in
1911, thus our analysis of the two disease groups starts from that
year. Classification systems have however changed. Sweden did not join ICD
until 1952, but the historical causes of death in the Poplink database have
been coded according to the ICD10 system. In the present study, we have
identified the codes relevant for cardiovascular diseases and cancers in
the ICD 8, 9 and 10.

The distribution of causes of death is shown in
Figure \@ref(fig:causeskum). Notice the high proportion of other causes in
the first two periods. This category include cases with unknown or
unspecific causes which has implications for the cause-specific analyses
(see later discussion). The total dominance of our two disease groups is
apparent.

```{r causeskum, fig.cap = "Causes of death by time period (top: cancer, middle: cardiovascular, bottom: other).", fig.height = 4.5}
vb01 <- vb0[vb0$event == 1, ]
cancer <- tapply(vb01$cancer, vb01$period, sum)
heart <- tapply(vb01$heart, vb01$period, sum)
death <- tapply(vb01$period, vb01$period, length)
other <- death - cancer - heart
ttt <- rbind(cancer, heart, other)
ttt <- prop.table(ttt, 2) * 100
barplot(ttt[3:1, -(1:2)], col = 3:1, las = 2, ylab = "Per cent")
```

## Urban vs rural residence

In 19th and early 20th century Sweden (as in other countries), mortality
was much higher in urban environments [@hani94, @edv92]. The unsanitary and overcrowded
living conditions as well as the risks for importing infections made the
health risks higher in towns and cities. In the present paper, we control
for this aspect by distinguishing between urban and rural residence, based
on parish of residence.

## Periods for analysis

We have made separate analyses for the different periods. The century
1851--1950 is divided into four periods of 25 years each. The 19th century
periods represent a mainly pre-industrial society dominated by agriculture
and with only basic welfare provisions. During the years 1901--1950,
industrialisation started and the urban environments increased in
population size. The modern Swedish welfare state began to develop during this
time. For the period 1960 onwards, there are separate analyses on decades
except for the two last periods. Sweden had developed into a wealthy
welfare society with low income inequality,
particularly until the 1990s. A financial crisis struck Sweden in the early
1990s, which among other things increased unemployment and changed parts of
the welfare system towards being less generous than before. The start of the
last period (2008--2013) coincides with the international financial
crisis.

# Models

The proportional hazards model in a survival analysis context is used,
allowing for adjustment for marital status and rural/urban environment. Since
the central explanatory variable, social class (Hisclass), do not follow the
proportionality property, the analyses are stratified with respect to the
variable, and the main results are presented graphically.

Another practical problem is the size of that data set:
It contains no less than `r n.rec + pn.rec` records of `r n.pl + pn.pl` individuals, of
which `r n.d + pn.d` are observed to die.
This makes it very inconvenient to use the nowadays so popular semi-parametric
technique known as
\emph{Cox regression} [@cox72,@gb02]. Instead we use a parametric model
that allows for reduction of the data
to \emph{sufficient statistics} [@fisher22]. The idea is to keep all
relevant covariates \emph{categorical}, and then create a table with all
possible combinations of levels of the covariates. For each such
combination, the total amount of exposure and the number of deaths are
calculated.  We have two sexes, three marital statuses, ten periods, two
urban/rural groups, four social classes at age 40,  four social classes at age 60,
ten age groups, and 22 cohorts, in all 422\,400 possible
combinations. However, in this scheme there are many structural zeroes
(impossible combinations) and some random zeros, so the actual number of
combinations we end up with is 24\,326. The data set is thus reduced in
size by 99 per cent.

The parametric model we use for the baseline hazard function is piece-wise
constant on five-year intervals, so we arrive at the following proportional
hazards regression model:

\begin{equation}\label{eq:prop}
h(t; \lambda, \boldsymbol{\beta}, \textbf{x}) =
\lambda(t)e^{\boldsymbol{\beta} \mathbf{x}}, \; \; 40 < t \le 90,
\end{equation}
where

\begin{equation*}
\begin{split}
\mathbf{x} &= (x_1, \ldots, x_n) \text{\ is a vector of covariates,} \\
\boldsymbol{\beta} &= (\beta_1, \ldots, \beta_n) \text{\ is the corresponding vector of regression
parameters.} \\
\end{split}
\end{equation*}
and
\begin{equation*}
\lambda(t) = \alpha_i,\; 40 + 5(i - 1) < t \le 40 + 5i, \; i = 1, \ldots, 10.
\end{equation*}

Since all components of $\mathbf{x}$ are categorical covariates, the
estimation of \eqref{eq:prop} can be performed in two ways, first by the
original data set and a survival analysis routine handling a piecewise
constant hazard, or second, by the reduced data set (table) and
\emph{Poisson regression}. The second way is faster than the first by a
factor of size at least 100, often much more, depending on hardware, and
the point is that the two ways give exactly the same results. This is
because the table is a \emph{sufficient statistic} for \eqref{eq:prop} and
the original data set.

The analyses are performed in the **R** environment for statistical computing
and graphics [@cran], especially using the package **eha**
[@gb12,@eha].
The proportional hazards model allow us to scrutinise the differences and
the development in more detail. We divide the analysis into two age groups,
40--64 and 65--89 years of age, the first age group representing people that
were still mainly in the workforce, and the second those that mainly were
retired. The model controls for marital
status and whether the individual resides in the urban or rural parts of the regions. There
are no serious signs of deviation from proportionality assumption, except
regarding docial class. Each period are analysed separately, but we then combine results
in order to show the development of the social patterning of mortality
during the studied period.

The important explanatory variable, social class (**hisclass**), is included in the
proportional hazards models as a *stratification* variable. The main
reason for this choice is that the effect of social class (**hisclass**) on survival is
non-proportional, that is, it varies with age. With the reduced data set
and Poisson regression, this is simply achieved by introducing an
interaction term between **age** and social class (**hisclass**). Then, the main
results are presented graphically. Tables with results from the
different models are presented at the end of the article, but are not
commented on in detail. The comments on the results are primarily based on
the figures and focuses on the central variable in our study--- social class (**hisclass**). 
As can be seen from the tables, most variables in the models are
significant at 5 per cent level.

# The effect of social class, 1851--1950, 1961--2013

## All causes of death

```{r colorc}
## Colors for plotting strata:
scol <- c("blue", rep("darkgreen", 2), "red")
## Common for all groups:

labb <- levels(vb0Tab$period)
nper <- length(labb)
#source("R/getPHglm.R")
## End 'common'
```

The cumulative
hazards at the end of the time period (65 if 40--64, 90 if 65--89) are the
main target in our analysis: These numbers are used in the process of
graphically illustrating the development over time.

### Ages 40--64

```{r getske4060}
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, women."
scap1w <-  "Cumulative hazards, ages 40--64, women."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, men."
scap1m <-  "Cumulative hazards, ages 40--64, men."
##
cap0 <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom)."
scap0 <- "Total hazard by HISCLASS, ages 40--64."
cap12 <- "Total hazard of dying before age 65 for a 40 year old person standardised by HISCLASS and decade, women (top) and men (bottom)."
scap12 <- "Total hazard standardised by HISCLASS, ages 40--64."
cap12log <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades."
scap12log <- "Total hazard by HISCLASS, ages 40--64, last years."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, women."
scap2w <-  "Cumulative hazards, ages 65--89, women."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, men."
scap2m <-  "Cumulative hazards, ages 65--89, men."
##
cap22 <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom). Standardised."
scap22 <- "Total hazard by HISCLASS, ages 65--89."
cap22log <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades."
scap22log <- "Total hazard by HISCLASS, ages 65--89, last years."

##
vb4064 <- vb0Tab[as.numeric(as.character(vb0Tab$age)) < 64.5, ]
vb4064$age <- factor(vb4064$age)
##vb4064$hisclass <- vb4064$hisclass40
```

```{r hazcalw}
source("R/getPHglm2.R")
##nper <- 4
nper <- length(labb)
resw <- getPHglm2(vb4064[vb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb4064[vb4064$sex == "male", ],
            nper, labb)
```

```{r cumsw, fig.cap = cap0, fig.scap = scap0, fig.height = 8}
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, ylab = "TMR(65|40)",
                main = "Women 40-64", exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, ylab = "TMR(65|40)",
                main = "Men 40-64")##, exclude = 5:6)
par(oldpar)
```

From Table~\ref{tab:cumswtab} it is obvious that the statistical significance is strong.
The most interesting feature is the three-way interaction with **sex**, **period**, and social class (**hisclass**).

```{r cumswtab,results='asis',cap="Table"}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex * hisclass * period,
            data = vb4064, family = poisson)
dr <- drop1(fitt, test = "Chisq")
##knitr::kable(dr)
xtable(dr,label = "tab:cumswtab", caption = "Anova, all causes of death, 40-65.")
```

Figure \@ref(fig:cumsw) shows the summed hazards of dying in the age interval
40--64 during the different periods. Obviously, and not surprisingly, mortality
is decreasing over time for both sexes. Women have consistently better survival
than men. This is a well-known phenomenon in Swedish
demographic history. Male mortality has almost always been higher than
female in all age groups [@wil99, @suwi07]. We also observe a
substantial decline in mortality for both sexes, resulting in smaller
absolute differences between social classes. We must however distinguish between differences in absolute risks compared to relative risks [@hoetal16). The compression that is
visible for later decades is partly explained by the decreasing general
mortality and is not necessarily representing a homogenisation of relative risks.

By standardising the levels in each time period for each class, we 
illustrate the relative risks in Figure \@ref(fig:cumswstand). In the rest of 
the paper, we only present these
relative mortality levels for the different classes since they are the main focus in this study.

```{r cumswstand, fig.cap = cap12, fig.scap = scap12, fig.height = 8}
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(65|40)", main = "Women 40-64", exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(65|40)", main = "Men 40-64")##, exclude = 5:6)
par(oldpar)
```

Notice that we have not analysed the 1950s due to the fact that we only
have data for some parishes at that time. Furthermore, we have excluded the
results for women during the 1960s and 1970 out of the reasons discussed above. The
changing economy also have consequences for the analysis, particularly when
it comes to the agricultural sector. While farmers were common occupations
during the first half of the 20th century, these occupations has constantly
decreased in numbers thereafter. This means that the middle class has
somewhat different character during the last periods.

The most striking result of the analysis is the different patterns for men
and women. The social pattern of mortality among women is quite much in
accordance with previous analyses from the present research group on other
regions in northern Sweden and for other periods [@edli11,@edbr12]. The
highest social class usually have comparatively low mortality, while
working class women have high. There is however no substantial change in
the relative risks between classes. Only the group with unknown social
class is distinguished by much higher mortality from the 1980s
onwards. This can however have to do with how we have defined social
position for women before and after 1950. After 1950 it is based
on their own occupation. Being without occupation might include women that
are more frail both when it comes to health and social characteristics.

A different pattern appears among men, being the most surprising finding in
our analyses. The groups we would expect to have the lowest mortality, if
we assume that access to different resources and belonging to a high social class
determines survival, instead have by far the highest hazards of dying
during the nineteenth and much of the 20th century. Workers on the
other hand have comparatively low mortality until the 1960s, at the same
level as the middle class. The relative positions change radically towards
the end the 20th century. From the 1980s, there is a clear advantage when
it comes to survival for higher managers and professionals. The relative
differences are however more compressed than for the period until 1950. The
expected social gradient in mortality has developed. This is also what we
usually find in analyses of social health inequalities in present-day
Sweden as well as in most other countries. In the regions we are studying,
this is however a quite recent phenomenon. Corresponding to what we find
for women, those without known occupation have much higher mortality than
the others from the 1960s onwards. The assumption made for women of
unknowns social class is probably even more relevant for men. They probably
represent groups in society that are particularly vulnerable, having difficulties getting into the laour force.

A further observation can be made from Figure \@ref(fig:cumsw). The gender
difference in death hazards is very big for the élite group from
1851--1950, men having about three times higher risks than women in the
same class. In the other classes, the difference was much smaller. It was
men in high position that experienced the largest risks.

### Ages 65--89

```{r getske6089}
vb6589 <- vb0Tab[as.numeric(as.character(vb0Tab$age)) >= 64.5, ]
vb6589$age <- factor(vb6589$age) # NOTE:: Change to '6589'!
##vb6589$hisclass <- vb6589$hisclass60
```

We now turn to the elderly, i.e. those that in most cases have left the workforce.
Corresponding to the previous analysis, we use the value of the cumulative hazards at
duration 25 (age 90), starting from age 65 as a summary measure. In Table~\ref{tab:cumswoldtab} the results of the model is presented.

The general pattern and development of mortality (not shown here) do not
differ that much from the younger age group, of course representing a much
higher mortality level earlier in history. Mortality among women was, as
expected, lower than among men. We find a clear decline in mortality, but
the fastest decline did not set in definitely until sometimes after the
Second World War (results not shown here), i.e. later than for the age
group 40--64 years. This corresponds to the Swedish development where we
find a turning point at this time with better survival among elderly, to a
large extent caused by lower mortality in cardiovascular diseases.

Concerning the social pattern, the pattern is less clear compared to what
we found for the younger age group (see Figure \@ref(fig:cumswold). For
women the results are mixed. Once again, too much missing data on social
class leaves out the results for the 1960s and 1970s. We do not find the
same clear advantage for élite women during the period before 1950, but neither
do we find any large disadvantage for working class elderly women at that
time. In some periods, the social differences were small, and the élite
even had substantially higher mortality during the early 20th century. The
social differences continued to be small during the last decades of the
20th century. Only during the two last periods (after the year 2000) is it
possible to establish the expected social gradient in mortality.

For men we find indications of the same intriguing change among elderly as
for the adult male population 40--64 years. Mortality was in fact
substantially higher among élite men  but their disadvantage was most 
apparent during the 19th century. Differences are fairly small during the
first half of the 20th century. Not until the second half of the 20th
century is the survival much better for the highest social class while
workers in the age of retirement had high mortality in comparison.

```{r cumswold, fig.cap = cap22, fig.scap = scap22, fig.height = 8}
##source("R/plotED.R")
resw <- getPHglm2(vb6589[vb6589$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb6589[vb6589$sex == "male", ],
            nper, labb)

oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(90|65)", main = "Women 65-89", exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(90|65)", main = "Men 65-89")##, exclude = 5:6)
par(oldpar)
```

See Table~\ref{tab:cumswoldtab}.

```{r cumswoldtab,results='asis'}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex * hisclass * period,
            data = vb6589, family = poisson)
dr <- drop1(fitt, test = "Chisq")
##knitr::kable(dr)
xtable(dr,label = "tab:cumswoldtab", caption = "Anova, all causes of death, 65-89.")
```

## Cardiovascular mortality

In this and the following section, the focus is on analyses of mortality from the two
dominant causes of death groups among the adult and elderly population. We start
with cardiovascular diseases. Many studies have shown that socio-economic
conditions are strongly associated with mortality in this disease
group. Belonging to working class, having less education and low income
increases mortality risks. We would therefore expect to find a clear
social pattern in the present analysis.

A general observation is that the results on causes of death and
its development is difficult to interpret, partly because the number of
cases are fewer but also that the quality and coverage of reported death
causes differ in time. Thus, many deaths during the early 20th got quite vague diagnoses reported in the death registers. There was still a substantial part
of all deaths where cause was diagnosed as unknown or dying from old age,
despite the fact that reporting causes of death had become mandatory. For
the period 1901--1925 to 1926--1950, the proportion of deaths from
cardiovascular diseases increased from 25 to 43 per cent for the age group
40 and above, thus indicating an increase of these diseases. However, at
the same time unknown cause and death of old age decreased from 28 percent
to 18 per cent. It is quite reasonable to suspect that many cardiovascular
deaths were disguised in unspecific categories and that the increase is partly
explained by improvements in how deaths were diagnosed, something that also seems to have been the case in Scotland at this time [@reidetal15].

```{r causes41}
cvb <- vb0
cvb$event <- cvb$heart ## Can this be done better?
cvb <- cvb[!(cvb$period %in% c("1851-1875", "1876-1900")), ]
cvb$period <- factor(cvb$period)

# load("data/vb.rda")
# ##vb <- cal.window(vb, c(1961, 2001))
# ##x <- substr(vb$ULORSAK, 1, 2)
# vb$event <- vb$event & vb$heart
# vb$exposure <- with(vb, exit - enter)
# vb$cohort <- factor(vb$cohort)
# vb$period <- factor(vb$period)
cvbTab <- aggregate(cvb[, c("event", "exposure")],
                    cvb[, c("sex", "age", "period", "cohort",
                           "urban", "civst", "hisclass")],
                    FUN = sum)
 cvbTab$hisclass <- factor(cvbTab$hisclass)
 cvbTab$age <- factor(cvbTab$age)
```

```{r 41colorc}
## Colors for plotting strata:
scol <- c("blue", rep("darkgreen", 2), "red")
## Common for all groups:

labb <- levels(cvbTab$period)
nper <- length(labb)
##source("R/getPHglm.R")
## End 'common'
```

As before, models are fitted separately for sex and each ten-year time period, from  1901 to
2013. We stratify on social class (**hisclass**) and include the covariates **urban** (TRUE or FALSE) and
**civst** (marital status). The cumulative
hazards at the end of the age period (65 if 40--64, 90 if 65--89) are the
main target in our analysis: These numbers are used in the process of
graphically illustrating the development over time.

### Ages 40--64

```{r 41getske4060}
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, women, cardiovascular death."
scap1w <-  "Cumulative hazards, ages 40--64, women, cardiovascular death."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, men, cardiovascular death."
scap1m <-  "Cumulative hazards, ages 40--64, men, cardiovascular death."
##
cap12 <- "Total hazard of dying before age 65 for a 40 year old person standardised by HISCLASS and decade, women (top) and men (bottom), cardiovascular death."
scap12 <- "Total hazard by HISCLASS, ages 40--64, cardiovascular death."
cap12log <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cardiovascular death."
scap12log <- "Total hazard by HISCLASS, ages 40--64, last years, cardiovascular death."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, women, cardiovascular death."
scap2w <-  "Cumulative hazards, ages 65--89, women, cardiovascular death."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, men, cardiovascular death."
scap2m <-  "Cumulative hazards, ages 65--89, men, cardiovascular death."
##
cap22 <- "Total hazard of dying before age 90 for a 65 year old person standardised by HISCLASS and decade, women (top) and men (bottom), cardiovascular death."
scap22 <- "Total hazard standardised by HISCLASS, ages 65--89, cardiovascular death."
cap22log <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cardiovascular death."
scap22log <- "Total hazard by HISCLASS, ages 65--89, last years, cardiovascular death."

##
cvb4064 <- cvbTab[as.numeric(as.character(cvbTab$age)) < 64.5, ]
cvb4064$age <- factor(cvb4064$age)
```

Performing the same types of analyses as for all-cause mortality, we start
by looking at cumulative hazards by Hisclass at age 65, given
survival to age 40. The result of the model can be seen in Table \@ref(tab:41cumswtab).

```{r 41hazcalw}
source("R/getPHglm2.R")
resw <- getPHglm2(cvb4064[cvb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(cvb4064[cvb4064$sex == "male", ],
            nper, labb)
```

```{r 41cumsw, fig.cap = cap12, fig.scap = scap12, fig.height = 8}
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE, exclude = 3:4,
                    ylab = "TMR(65|40)",  main = "Women 40-64")
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE, ##exclude = 1:2,
                    ylab = "TMR(65|40)", main = "Men 40-64")
par(oldpar)
```

```{r 41cumswtab,results='asis'}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex * hisclass +
                sex * period + hisclass * period,
            data = cvb4064, family = poisson)
dr <- drop1(fitt, test = "Chisq")
knitr::kable(dr, caption = "Anova, cardiovascular deaths, 40-64.")
#print(xtable(dr,label = "tab:41cumswtab", caption = "Anova, cardiovascular deaths, 40-64."), type = "html")
```

See Table \@ref(tab:41cumswtab).

Figure \@ref(fig:41cumsw) shows mortality levels during different periods
from 1901 on-wards. Women from the highest social class had consistently
lower mortality from cardiovascular diseases compared to the others. We
also observe increasing social differences from the first half of the
century to the last decades. A new pattern has developed from 1980
onwards. Now working class women have substantially higher cardiovascular
mortality compared to others. We can however not exclude that this is an
artefact of different diagnostics for deaths in different groups. Working
class women may have had less precise diagnoses in the early 20th century.

For men, there is a different pattern, a pattern however that to a large
extent is consistent with what we observed concerning all-cause
mortality. The elite group had much higher cardiovascular diseases up until
the middle of the 20th century, but instead have the lowest mortality from
1980 onwards. Working class men had a fairly low cardiovascular mortality,
but in the same way as among women, the levels were fairly similar to the
one of the middle category consisting of white collar, lower managers and farmers. In the last decades and similarly as for women, working class men have the highest
mortality from this cause of death group. A social
hierarchy in cardiovascular mortality developed both for men
and women.

### Ages 65--89

```{r 41getske6089}
cvb6589 <- cvbTab[as.numeric(as.character(cvbTab$age)) >= 64.5, ]
cvb6589$age <- factor(cvb6589$age) # NOTE:: Now change to '6589'!
```

In this section we focus on cardiovascular mortality among those that
mostly have left the workforce, the age group 65--89 years. We use as a
summary measure the value of the cumulative hazards at
duration 25 (age 90), starting from age 65. The results from the model is presented in Table \@ref(tab:41cumswoldtab).

```{r 41hazcalwold}
resw <- getPHglm2(cvb6589[cvb6589$sex == "female", ],
              nper, labb)
resm <- getPHglm2(cvb6589[cvb6589$sex == "male", ],
                nper, labb)
```

Concerning the social pattern, we find somewhat different results for elderly women compared to the women in working-active age. Cardiovascular mortality among elderly
women do not present a consistent class pattern. Female cardiovascular
mortality was somewhat higher among the elite and the middle group than
among working class women until the last decades. Elderly from a working
class background were not disadvantaged for most of the 20th century. The
social class pattern has changed somewhat during the last periods becoming more
in line to what is expected. The differences are however rather small.

Even if results for elderly men point in different directions for some
periods, the general trend resembles that of working-active men. The
highest class had high cardiovascular mortality during the first half of
the 20th century, while the expected social pattern developed and became apparent only
during the most recent periods.

```{r 41cumswold, fig.cap = cap22, fig.scap = scap22, fig.height = 8}
##source("R/plotED.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "H(90|65)",  main = "Women 65-89", exclude = 3:4)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "H(90|65)", main = "Men 65-89")
par(oldpar)
```

```{r 41cumswoldtab,results='asis'}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex * period *hisclass,
                ##sex * period + hisclass * period,
            data = cvb6589, family = poisson)
dr <- drop1(fitt, test = "Chisq")
knitr::kable(dr, caption = "Anova, cardiovascular deaths, 65-90.")
##xtable(dr,label = "tab:41cumswoldtab", caption = "Anova, cardiovascular deaths, 65-90.")
```

See Table \@ref(tab:41cumswoldtab).

## Cancer mortality

While we assume clear social differences in cardiovascular diseases, this
is not necessarily an obvious assumption when it comes to
cancers. Different forms of cancers have different origin and
different pathways to disease and death. The social pattern of higher
mortality in classes with less power and accordingly less life chances are
considered to be less clear when it comes to cancers. In our analyses,
there is a risk that cancer mortality is under-estimated during the first
half of the 20th century, illustrated by the already mentioned large
proportion of unknown or vague causes at that time. We believe however that
cancers were better diagnosed than cardiovascular deaths. This is supported
by the stability in proportion of cancer deaths between the two first
periods, 12.5 per cent 1901--1925 and 13.5 per cent for the period
1926--1950.

```{r causes12}
cvb <- vb0
cvb$event <- cvb$cancer ## Can this be done better?
cvb <- cvb[!(cvb$period %in% c("1851-1875", "1876-1900")), ]
cvb$period <- factor(cvb$period)

# load("data/vb.rda")
# ##vb <- cal.window(vb, c(1961, 2001))
# ##x <- substr(vb$ULORSAK, 1, 2)
# vb$event <- vb$event & vb$heart
# vb$exposure <- with(vb, exit - enter)
# vb$cohort <- factor(vb$cohort)
# vb$period <- factor(vb$period)
cvbTab <- aggregate(cvb[, c("event", "exposure")],
                    cvb[, c("sex", "age", "period", "cohort",
                           "urban", "civst", "hisclass")],
                    FUN = sum)
 cvbTab$hisclass <- factor(cvbTab$hisclass)
 cvbTab$age <- factor(cvbTab$age)
```

```{r 12colorc}
## Colours for plotting strata:
scol <- c("blue", rep("darkgreen", 2), "red")
## Common for all groups:

labb <- levels(cvbTab$period)
nper <- length(labb)
source("R/getPHglm.R")
## End 'common'
```

We use similar types of analyses as in the earlier sections. The cumulative
hazards at the end of the period (65 if 40--64, 90 if 65--89) are the
main target in our analysis. These numbers are used in the process of
graphically illustrating the development over time.

### Ages 40--64

```{r 12getske4060}
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, women, cancer death."
scap1w <-  "Cumulative hazards, ages 40--64, women, cancer death."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, men, cancer death."
scap1m <-  "Cumulative hazards, ages 40--64, men, cancer death."
##
cap12 <- "Total hazard of dying before age 65 for a 40 year old person standardised by HISCLASS and decade, women (top) and men (bottom), cancer death."
scap12 <- "Total hazard by HISCLASS, ages 40--64, cancer death."
cap12log <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cancer death."
scap12log <- "Total hazard by HISCLASS, ages 40--64, last years, cancer death."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, women, cancer death."
scap2w <-  "Cumulative hazards, ages 65--89, women, cancer death."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, men, cancer death."
scap2m <-  "Cumulative hazards, ages 65--89, men, cancer death."
##
cap22 <- "Total hazard of dying before age 90 for a 65 year old person standardised by HISCLASS and decade, women (top) and men (bottom), cancer death."
scap22 <- "Total hazard standardised by HISCLASS, ages 65--89, cancer death."
cap22log <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cancer death."
scap22log <- "Total hazard by HISCLASS, ages 65--89, last years, cancer death."

##
cvb4064 <- cvbTab[as.numeric(as.character(cvbTab$age)) < 64.5, ]
cvb4064$age <- factor(cvb4064$age)
```

```{r 12hazcalw}
source("R/getPHglm2.R")
resw <- getPHglm2(cvb4064[cvb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(cvb4064[cvb4064$sex == "male", ],
            nper, labb)
```

```{r 12cumsw, fig.cap = cap12, fig.scap = scap12, fig.height = 8}
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE, exclude = 3:4,
                    ylab = "TMR(65|40)",  main = "Women 40-64")
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE, ##exclude = 1:2,
                    ylab = "TMR(65|40)", main = "Men 40-64")
par(oldpar)
```

```{r 12cumswtab,results='asis'}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex *hisclass +
                sex * period + hisclass * period,
            data = cvb4064, family = poisson)
dr <- drop1(fitt, test = "Chisq")
knitr::kable(dr, caption = "Anova, cancer deaths, 40-64.")
##xtable(dr,label = "tab:12cumswtab", caption = "Anova, cancer deaths, 40-64.")
```

For the results of the model, see Table \@ref(tab:12cumswtab).

Figure \@ref(fig:12cumsw) shows relative mortality levels according to
Hisclass during the different studied periods for the age group
40--64. Cancer mortality among the elite women was lower than that for the
other classes during the complete studied period, while working class women
had the highest levels during almost all periods. Consequently, we find a
more consistent advantage of higher class in cancers compared to
cardiovascular diseases for women, thus confirming what we have expected. We do however not find the same pattern
for men. It is difficult to identify any general features of 20th century
male cancer mortality when it comes to social differences.

### Ages 65--89

```{r 12getske6089}
cvb6589 <- cvbTab[as.numeric(as.character(cvbTab$age)) >= 64.5, ]
cvb6589$age <- factor(cvb6589$age) # NOTE:: Now change to '6589'!
```

In our analyses of class patterns in mortality, we finally turn to cancer
mortality among the elderly. The model is presented in Table \@ref(tab:12cumswoldtab).

```{r 12hazcalwold}
resw <- getPHglm2(cvb6589[cvb6589$sex == "female", ],
              nper, labb)
resm <- getPHglm2(cvb6589[cvb6589$sex == "male", ],
                nper, labb)
```

The relation between different classes and cancer mortality fluctuates
considerably both for women and men in this age group. For some periods it
has to do with quite few events to analyse. Women of the highest class have
favourable mortality during several periods, consistent with the advantages
we found for women in the younger age group, but on the other hand had high
levels at other times. The social differences are however quite small
during the most recent periods.

For men we find a similar lack of social pattern in cancer mortality as we
found for men in working-active ages. There is however a tendency during
the last decades of higher mortality among working class men.

```{r 12cumswold, fig.cap = cap22, fig.scap = scap22, fig.height = 8}
##source("R/plotED.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(90|65)",  main = "Women 65-89", exclude = 3:4)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(90|65)", main = "Men 65-89")
par(oldpar)
```

```{r 12cumswoldtab,results='asis'}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex +
                sex * period + hisclass * period,
            data = cvb6589, family = poisson)
dr <- drop1(fitt, test = "Chisq")
knitr::kable(dr, caption = "Anova, cancer deaths, 65-90.")
##xtable(dr,label = "tab:12cumswoldtab", caption = "Anova, cancer deaths, 65-90.")
```

See Table \@ref(tab:12cumswoldtab).

\section{Discussion}

The authors of this paper have scrutinised different aspects of the social
pattern of mortality in historical northern Sweden for some time, focusing
on different regions, different age groups and different time periods
[@edv92, @edv04, @edli11, @edbr12]. When it comes to infant and child
mortality pre---and early---transitional time, the expected socio-economic
differences was often but not always confirmed [@edv92, @edv04],
something that has been found in other studies as well [@bredro02,
@bube01, @mol17]. Studies based on microdata on the adult population in
history have been rare, mainly due to lack of adequate data. The results
from such studies show other patterns than that of infant and child
mortality. The expected social gradient in survival, meaning that people
from higher social classes always had an advantage, was in many cases,
particularly among adult men, not found despite the well-documented
socio-economic differences [@edli11, @edbr12]. There were however large
differences between men and women and for different age groups. These
earlier studies have suggested that distinct and consistent differences to
the advantage of higher classes developed somewhere during and probably
quite early in the 20th century [@bevp11]. Until now we have however
not been able to investigate the continuous development on mortality
differences among the adult and elderly population due to no available
data. This is the first time it has been possible, and the results continue
to surprise us. The consistent advantage for higher social classes do not
appear until quite late in the regions we have analysed. The social pattern
of female mortality are in line to what is expected, but a social gradient
in male adult mortality seems to be a recent phenomenon. Furthermore, this
transition and the class pattern of mortality differ depending on age
group. Inequality in mortality appear clearer for the population at working
age, while the differences are smaller among the elderly.

The results of small socio-economic gradient in the past and its late
development are however in accordance with other studies [@bevp11],
for example by @bedr11 in a very different setting, a couple of
Scanian parishes, although with somewhat different social categories. They
found the first indications of an emerging social gradient in the
1950s. Similar results have also been found in Estonia [@jaadla17].

When it comes to specific causes of death, the analysis shows that
cardiovascular diseases contributes to a large proportion of all deaths
during the 20th century, and we are not surprised that it has a fairly
similar pattern as for all-cause mortality. Cancer is sometimes considered
less connected to social class, and we find mixed results in our analyses.

The surprising results makes it reasonable to consider the validity and
possible weaknesses in the study. One obstacle concerns the quality of
available information. Occupational titles in historical sources are not
always distinct enough to define the correct class. It can be difficult to
identify the skill level in the working class population because many were
often only designated as labourer. The size of the enterprise for free
entrepreneurs is unknown and it may be impossible to distinguish if they
are self-employed or have employees. The class system and the position
of different classes can also be discussed. Studies covering long period
have the problem that occupations and classes change character - they do
not represent the same thing. Nevertheless, these ambiguities and possible
mis-classifications are not of that sort and magnitude that it would
radically change the basic patterns observed. The social classes identified
during the different periods represent distinct groups with very large
differences in resources, capacities and power. Available data, for example
from the 1930 census confirm that class differences in income and wealth
were large, often enormous.

In this paper we have analysed social inequality in health from the
perspective of social classes using Hisclass. However, there are
alternative ways to analyse social inequality, and we cannot exclude that
such analyses gives other results or at least would modify the image
presented here. In many modern studies categorisations based on income,
wealth or education are common for studies of inequality. They represent
different dimensions of inequality. One problem is however that reliable
data on income and education are rarely available in historical population
databases.  As a test, we have included analyses of the associations of
education and income to mortality for recent decades. For women as
well as for men, for both age groups and for all-cause mortality as well as
cause-specific mortality we find consistent differences with higher
educational level and more income leading to lower mortality, thus
basically confirming what we find in the class analysis. We do not know if
education and income are equally important in the past.

The WHO Commission on Social Determinants of Health framework for the
social determinants of health provides a model on how socio-economic
position influence health. Social class works through intermediate
determinants related to exposure and vulnerability [@who10]. People
are situated within socio-economic and political contexts, as well as
prevailing value systems in society. Socio-economic position imply
different access to resources, power and prestige, all being aspects of
crucial importance. The framework takes a holistic approach to the social
determinants, including not only the effect of pure economic and material
resources. The importance of understanding people in their context is also
emphasised by for example @kri05. What she calls embodiment refers
to the recognition that human beings are simultaneously social beings and
biological organisms. The WHO framework  [@who10] lists three
different types of intermediary social determinants of health; material
circumstances, behaviours, biological factors, and finally psycho-social
factors. Most studies focus on material circumstances. Having a good
economy gives advantages in many aspects of relevance for health and
mortality, for example good living conditions, access to food, medical care
and better working conditions. Even though these advantages were at hand in
our studied regions, the final result was not better health for men in the
most privileged groups until the late 20th century. We must therefore
consider other explanations. Psycho-social factors have been frequently
discussed during the most recent decades, for example by @mar04, who
argues that a higher status leads to better health, regardless of where in
the status ladder we make our observations. Psychosocial factors are also
suggested by @wipi09 as an explanation why equal societies always
perform better than unequal ones when it comes to health and other social
conditions. In the case studied here, we do not find any indication of a
status syndrome, since mortality levels was in the opposite directions to
the one we would expect from these theories.

This makes us consider the possible impact of behavioural factors. Our
results of the gender aspect might be of interest in this case. When it
comes to gender differences, the present study confirms what we previously
have found in the Sundsvall region [@edli11]. There is a clear
gendered pattern when it comes to social inequality in mortality.  Within
the same households, the effect of social class seems to lead to different
results, high social position being beneficial for women but not for
men. We might consider the role of behaviours based on different gendered
expectations. Having a high social status could be expressed in attitudes
and behaviours that were harmful for health (this is discussed in more
detail in @edli11. Alcohol has been suggested as part of the large
gender differences in the 19th century [@wil99, @edv92], and a more
risk-taking life may be a component of male mortality [@cour00]. There
are however some problems with attributing most of the male excess
mortality to heavy drinking. Alcohol consumption decreased substantially
during the period late 19th century to the 1950s, but the old pattern in
social health inequalities remained. Swedes at that time had very low
consumption levels. Alternative explanations should also be
considered. There were strong and heavy requirements of being successful
for the male \"elite. Many were involved in businesses with high risks and
they may have felt a constant threat of economic failure.

Our study illustrates the importance of considering contextual factors and
the restrictions of making general statements. We need to look into all
aspects of social classes, not only the material ones but also what this
means for behaviour, norms, way of living. @edli11 suggest that an
explanation for the larger social health inequalities in present-day
society is that high status is expressed through health, being fit and
physically active, while in many historical contexts higher social status
(for men) was expressed in low physical activity, obsessive eating,
drinking, smoking and sometimes also taking risks.  Different historical
contexts promote different ways to spend economic resources.This is also in
line with what @rasp06 and @jaadla17 suggest. If this
interpretation is correct, the hypothesis of @ant67 on an increasing
role of health related behaviour in the modern low-mortality régime,while
economic inequality determined the social mortality pattern during the
early phases of the mortality transition. Instead, health-related seems to
be at least equally strong earlier in history but having a completely
different expression.

One of the major findings in this study relate to gender. As several
researchers have pointed out, demography has often ignored power relations
and different roles within families [@watkins93, @mack95, @jans07]. By
investigating possible differences in mortality patterns according to sex,
we have identified gendered patterns within the family. Different patterns
between the sexes may be caused by discrimination on the female part in the
household. We know that the society was strongly segregated according to
gender and that women were discriminated and had less space for action. In
our case, we do not find this to have negative consequences for women,
indicating that the relation between sexes in the society was characterised
by distinct gendered expectations on behaviour.

Another question is if our results can be generalised to other historical
contexts. We have studied environments in northern Sweden without any large
metropolis. Perhaps Sweden and/or this part of Sweden is a special
case. The social differences may have been smaller, the involvement of
government both at the national
and local level may have been stronger, thus mitigating socio-economic
disadvantages. In comparison with many other countries, the level of
mortality has been fairly low. These are relevant questions, which
hopefully can be addressed in other and maybe comparative studies.

Finally, we argue that our results have implications for the understanding
on several other hot research topics. A central and hot debated question
concerning the mortality transition is how the enormous increase in
survival was accomplished. @mck76 argues that the improvements up to
the middle of the 20th century was primarily caused by improved
nutrition. This implicate that having more resources would lead to better
survival. If research shows that the most advantaged groups did not have
better health and longer lives, this do not necessarily speak against
McKeown's ideas but it makes the question more complex. The results are
also difficult to fit into @liph95 hypothesis that social conditions
should be seen as a fundamental cause of disease and not only expressions
of different disadvantages. In the historical context analysed here, this
idea is not supported.

In recent years, the question about the relation between health and income
inequality been widely discussed [@wipi09]. Even though our study do
not address this issue directly, our results are of interest for this
question as well. The
development of the social inequality in mortality do not seem to fit that
well into the development of income and wealth inequalities. The early 20th
century, in Sweden as well as in other countries, was characterised by very
large economic differences, still this did in our regions not lead to any
consistent advantage for the highest social classes. After the Second World
War, economic differences diminished, and was at its lowest in Sweden at
about 1980, i.e. at a time when a more expected pattern of a social
hierarchy in survival had appeared. The period thereafter is characterised
by increasing economic differences, and this may coincide with increasing
health differences as well. The long-term relation between economic
inequality and health is worth further studies.

# Acknowledgements

This work was undertaken within the research program *Ageing and
Living Conditions* at the Centre for Demographic and Ageing Research
(CEDAR) at Umeå University, financed by the Swedish Research Council (grant
No 349-2008-6592) and the project *Old age, health and inequality:
Mortality from a life course perspective in Sweden 1986--2009*, financed by
the Swedish Research Council (grant No 421-2012-892). We thank the
Demographic Data Base, Umeå University, for giving us access to data and
Jonas Helgertz, Department of Economic History, Lund University, for
providing us with scripts for creating Hisclass from Swedish Census
data.

# References