### 2017-09-05 article2.Rmd

---
title: Is high social class always beneficial for survival? A study of adults and
  elderly in Northern Sweden 1801--2013
author: "Sören Edvinsson and Göran Broström"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::word_document2:
    toc: yes
    toc_depth: 2
    toc_float: yes
bibliography: mybib.bib
documentclass: article
citation-package: natbib
biblio-style: apalike
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(engine='R',fig.path='figs/art2-',tidy=FALSE,fig.height=3.5,fig.width=8,strip.white=TRUE,cache=FALSE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
library(knitr)
library(eha)
library(xtable)
library(skum)
```

```{r readalldata}
source("R/get.file.R")
skel <- get.file("skel")
##
skume <- get.file("skume")
##
foli <- get.file("foli")
skel$urban <- FALSE
vb0Tab <- rbind(skel, skume[, names(skel)], foli[, names(skel)])
```

# Introduction

A major theme in demographic and epidemiological studies, is the seemingly persistent effect of social class on mortality. In the present study, we approach this issue by taking a long-term perspective on the development of social inequalities in mortality in the adult and elderly population. Our main focus is on the relative differences. The areas under investigation are the Skellefteå and Umeå regions in the north of Sweden for the period 1801--2013 for Skellefteå and 1901--2013 for Umeå. The main issue is then analysed according to gender and age-group (working age vs retired) and major causes of death (cardiovascular diseases and cancers). Furthermore we put this in the context of how the inequality in mortality is associated with the development of economic inequality in society. The results are discussed in relation to our understanding of the mortality transition and the social determinants of health and mortality, and their implications on some of the most influential hypotheses and concepts in health research.

On the basis of the results we present, we argue that high social class is not necessarily favourable for survival. Mortality risks in different contexts must be understood in the intersection between class and gender. We suggest that health-related behaviour was important not only in present-day societies, but were decisive also in earlier phases of the mortality transition. The results implicate that the association between social class and health is more complex than is assumed in many of the dominant theories in demography and epidemiology.

One of the central aspects of survival is social class and access to economic
and other resources. Even in present-day welfare societies, social position is a
strong determinant when it comes to health and mortality and the impact even
seems to be increasing [@kuetal04, @macketal16, @frlu07, @broba12, @strand10]. It has been suggested that
"... social conditions have been, are and will continue to be irreducible
determinants of health outcomes and therefore deserves their appellation as
'fundamental causes' of disease and death" [@liph95]. The persistence of social inequality in mortality to the disadvantage of the lower classes is one of the main assumptions in this theory [@liph98]. For a long time, the general view has been that socio-economic health and mortality inequalities were high in
historical societies, probably higher than in modern societies. This is a
reasonable assumption since these societies were in most cases characterised by
very large socioeconomic differences. Knowing that access to
resources provides advantages in all aspects of life, the health advantage of higher
classes ought to be obvious. @ant67 suggests that social
inequality in mortality has passed through different historical phases. According
to him, differences were comparatively small during the
pre-transitional phase when space was a strong determinant for the spread of
disease. Differences then increased during the transitional phase when mortality
declined and wealthy groups used their economic and human resources to gain better health. Finally,
mortality differences decreased again resulting in marginal levels of inequality in modern low-mortality societies when instead
health-related behaviour became the decisive determinant for health and survival.
@omran82 comes to a similar conclusion in his theory of the epidemiologic transition. He states in the third proposition in the theory, that even if the class differentials in mortality were
maintained during the transition, the decline set in earlier and was faster
among privileged groups.

Recent studies, investigating social inequality in health and mortality
with micro-data, have however questioned how general the assumed
pattern is [@bevp11]. Instead, some studies suggest that the modern
pattern might be a fairly recent development [@bedr11], although the pattern differs
depending on age, gender or historical context. When it comes to the adult
and elderly population results from different studies have diverged, some
reporting substantial differences but many, particularly those based on
Swedish data, have found surprisingly small or even reversed socio-economic
differences during the 19th and early 20th century [@bedr11,
@edli11, @edbr12], sharply contrasting from the situation in present-day society. It thus appears as if a more consistent social health divide in adult age appeared at
some point during the mortality transition, at least in Sweden. Solid
empirical evidence about the process is however lacking and studies
focusing on the issue are still few. There is a need for more reliable
studies from different geographical and historical settings in order to 
better understand the role of socio-economic conditions for health and survival over time.

Another aspect of the development of social inequality in mortality,
concerns its relation to economic inequality. @wipi09 argue that
income inequality has an independent effect on mortality, separate from the
direct effect of actual access to economic resources. They argue
that unequal societies (basically countries) perform less well when it comes to health (as well as other social conditions) than equal ones in the present-day economically developed world. This has initiated a vital scholarly debate and the topic has been extensively studied [@suka04,
@wado00]. During the last decades, the association between trends of inequalities in mortality and income respectively are weak or non-existant according to @hoetal16. How the association looked like in previous periods is unknown. This is of particular interest since the levels of economic
inequality were very different from those of the recent decades. Even if it
not a necessary implication, it is reasonable to assume that poorer groups were
those most disadvantaged of large inequalities.

When it comes to Sweden both income and wealth distribution was strongly skewed from the 1870's to the early 20th century. Starting during the inter-war period, income and
wealth inequality continuously diminished to reach a low point at around
1980. Thereafter, economic inequality has increased substantially until the
present, although far from being as high as a century ago [@rowa08,
@rowa09]. The Swedish development resembles that of other European and North
American countries in its basic features [@rowa14].

# The Skellefteå and Umeå regions

```{r mapp,fig.cap = "The Skellefteå and Umeå regions in Sweden.", out.width="3in"}
knitr::include_graphics("myfigs/sweden.png")
```
The Skellefteå and Umeå regions (Figure \@ref(fig:mapp)) are part of the
county of Västerbotten in the north of Sweden along the coast of Gulf
of Bothnia. This was a remote part of the country, where communication with
the rest of Sweden was difficult until the late 19th century. The economy
was dominated by agriculture, making it vulnerable to harvest failures, and
several severe famines frequently affected the regions during the 1800s, for example after the harvest failure in 1867 [@segb14]. During the long winters, sea communication was hindered due to the Gulf of Bothnia being frozen. The first ships during the year came to this area in May or even as
late as June in some cases [@fah56]. Towards the end of the 19th
century, the Swedish railway system reached this part of Sweden. Thereby
contacts with the rest of Sweden were facilitated, improving the economy and making it possible to mitigate the effects of harvest failures. The inland was particularly
remote and isolated. During the 19th century, our regions became
increasingly integrated in the same epidemiological pattern as the rest of
Sweden.

The Skellefteå region until the 1950 in our dataset consists of a selection
of parishes surrounding the town of Skellefteå, founded in 1845 but with a
very small population during the 19th century. The data from the period
after 1960 cover the Skellefteå and Malå municipalities, the same area as
for the earlier period but with the addition of two more parishes. The
majority of the 19th century population lived in rural villages and
hamlets, getting their livelihood from agricultural production. During the
20th century, industrialisation took place, partly by the opening of mines
in the inlands. This also led to a population increase both in the town
and in the rural parts, and a much more diversified economy.

Mortality was fairly low in comparison with other parts of the country and
the fertility transition was late [@coa86]. The Skellefteå population
size as defined in our data sets (all ages) was 6142 on January 1, 1801, 17355 on January 1, 1851, 43212 on
January 1, 1901, 62136 on December 31, 1950 and 76723 at the end of the
20th century.

The Umeå region consists of Umeå urban and rural parish until
1950, and from 1960 onwards of Umeå municipality when another three
parishes are included. This region has a somewhat different character from
that of Skellefteå. Umeå town had a small population but was substantially
larger than Skellefteå town during almost the whole studied period. It was
the administrative centre in the county of Västerbotten. Schools and
military regiments were placed here and the economy was more diversified
compared to Skellefteå. Agriculture dominated the rural part, but there
were also some foundries as well as industries for example in forestry and
small-scale production. The population size as defined in our data sets (all ages) was
19138 on January 1, 1901, 33393 on December 31, 1950 and 103970 when
the 20th century ended.

# Data and variables

The recent extension in time of data at the Demographic Data Base (DDB),
Umeå University (www.cedar.umu.se) by including the period between 1900 and
1950s makes it possible to investigate the long-term development of social
inequality in mortality in a part of Sweden. The data for the present study comes from two large
population databases at the DDB, whom provide us with micro-data on a sufficiently large population and
with substantial social diversity for the Skellefteå and Umeå regions
[@wee16]. The early period until 1950 is covered by
the database Poplink, the digitisation of historical parish registers for
the Skellefteå and Umeå region [@wee16]. Poplink is based on linked
parish records, allowing us to reconstruct life biographies on people as
long as they remained in the region. The records are linked within but not
between the regions. Data from Poplink are accessed for the period 1851--1950.

The other large data-set is extracted from the Linnaeus database
[@malm10]. Linnaeus is based on different linked national population
registers from 1960 to 2013 and is used within the ageing program at CEDAR, Umeå
University. The study period from 1960 to 2000 are constructed from
censuses every fifth year 1960--1990, with additional information on deaths
from National Board of Health and Welfare. For the period 2002--2013 we use
the information from the yearly population registers from Statistics Sweden
(LISA) together with death information from National Board of Health and
Welfare.

Information on individuals are anonymised and the two databases are not
linked. This stops us from following individuals present
in both databases throughout their lives, thus they are treated separately. This
also makes it impossible to add information on individuals in the Linnaeus
database from what we potentially could extract from Poplink, for example family
background or previous social class.

```{r lexis2,fig.cap="The sampling frame.", fig.height = 5}
plot(c(1801, 1801, 1951, 1951, 1801), c(40, 90, 90, 40, 40),
     type = "l", xlim = c(1800, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 90))
axis(1, at = c(1801, 1851, 1901, 1951, 1961, 1981, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1961, 1961, 2014, 2014, 1961), c(40, 90, 90, 40, 40),
      col = "blue", lwd = 2)
#
pn.exp <- round(sum(skel$exposure) + sum(skume$exposure))
pn.d <-  as.integer(sum(skel$event) + sum(skume$event))
#
text(1875, 70, "PopLink", cex = 2, col = "red")
##text(1899, 60, paste(pn.rec, "records"))
text(1875, 55, paste(pn.exp, "p-years"))
text(1874.5, 50, paste(pn.d, "deaths"))
#
n.exp <- round(sum(foli$exposure))
n.d <-  as.integer(sum(foli$event))

text(1985, 70, "Linnaeus", cex = 2, col = "blue")
##text(1984, 60, paste(n.rec, "records"))
text(1985.2, 55, paste(n.exp, "p-years"))
text(1985.2, 50, paste(n.d, "deaths"))
```

In the data-set analysed here, all individuals 40 years and older ever
being resident in either of the regions are included. The data file
contains the variables social class, sex, urban/rural
residence, birth date, death date, first and last date of observation and
type of entrance/exit. Total number of person years is `r as.integer(pn.exp)` leading to
`r pn.d` deaths in Poplink and `r as.integer(n.exp)` person years leading 
to `r n.d` deaths in the Linnaeus database. See 
Figure \@ref(fig:lexis2).

## Presence periods

Differences in available information in the two datasets as well as within
the Linnaeus data makes it necessary to apply different approaches when it
comes to identification of presence periods. The Poplink data provides us with exact dates or at least year of start and exit of the presence of all individuals in our data, allowing us
to have full and continuous control of the de jure population. This is
however not the case with the Linnaeus database. For the period 1960--2000
we use presence in the censuses 1960--1990 and information on deaths from the National
Board of Health and Welfare to identify the observation periods in our
analyses. There is however no full information on what happened in-between
the censuses, for example on the exact date of exit for persons having moved
between two censuses. We have solved this by considering everyone as
resident in our regions during the complete five-year period between
censuses if the person is missing in the later census (the last period covering ten years). This overestimates their presence, but should not bias the results when it
comes to socioeconomic differences. We furthermore miss those that both
moved in and moved out or vice versa between censuses. Exit dates are exact
if they refer to death, and in these cases we consider them as constantly
present from the last census to the death data even if the death took place
in another part of Sweden.

For the last period, 2002--2013, we use the yearly population registers
(LISA data), together with information on deaths from National Board of
Health and Welfare. This data allows for full control of the presence periods
in the studied regions.

## Social class

The choice of variable for analysing social differences vary between studies. A common alternative to a class analysis is to use income or education for
investigating socio-economic health differences. @goldk10 criticize
how many economists and epidemiologists operationalize inequality in their
studies, for example by not distinguishing between class and status
aspects. Class refers to relational aspects in society, while many other
variables focus on the distribution of material and/or immaterial
resources. He advocates studies of different aspects, which together
complement each other. Comparing different approaches are valuable in order
to understand why some are advantaged and others disadvantaged and what
sort of factors determines health differences.

In this paper we analyse mortality differences according to social class
based on the classification scheme Hisclass in modified form [@lema11]. The only
type of socio-economic information available both in the Linnaeus and Poplink database
is occupations. Thus, analysing according to social class is apart from
being a scientific choice, also a practical one. Using occupational data is
the only and obvious alternative for making a longitudinal study of
socio-economic mortality differences with consistent classification
schemes. Social position is treated as a time-dependent variable.

The availability to information on occupations varies over time and between
sources. Occupations in the version of Poplink we are using are coded according to the original DDB coding system that then has been adapted to Hisclass. The censuses 1960--1990 include occupational
codes based on the system used in Sweden at that time, NYK (Nordisk
Yrkesklassificering). We consider the observed social class in a census to be valid for the full period until the next census, corresponding to our definition of presence periods. For the period 1991--2000 the social class in the census 1990 is considered to be kept
until 2001. This strategy is also used for cases of missing social
class. Furthermore, we consider the last social position before retirement
to be kept until death or last observation for those above 65. For the
Linnaeus data, we use a script recoding HISCO into Hisclass, provided
to us by Jonas Helgertz, Lund University.

The use of different data sets, forces us to apply different approaches
towards defining the social position. A problem is that
Poplink usually only provides the occupation for the head of household,
thus under-estimating female working-force participation as well as that of
adult children residing with their parents [@vik10]. For the Poplink period (until 1950), we have chosen to categorise wives according to position of the head of household, usually
the husband/father, assuming that the family shares the same socio-economic
position. This strategy makes the problem of under-recording less
problematic.

Female labour is much better covered in the Linnaeus database. A
complication is however that it is  difficult to define households in the
same way as in Poplink. Extra-marital cohabitation, connected to the Second
Demographic Transition [@van87, @van03], became common and female
labour force participation developed as the norm in Sweden during this
period. The occupation of the husband as signifier of social class became,
if not obsolete, at least less relevant. All included persons are signified
by their individual occupation for this period. Still, the results for the
different periods are not completely comparable and makes the
interpretation of the development for women uncertain for some periods. Notice however that there is a strong social homogeneity between partners in a marriage, indicating that the results would not differ in any decisive way by using occupations on individual level instead of on household level for defining social class from the Linneaus data.

Hisclass, the classification system used here, is a "...
HISCO-based historical international social class scheme" [@lema11,
@lema02]. The different classes in Hisclass represent distinct
categories, based on whether the work is manual or non–manual, skill level,
supervision and sector. Implicitly it reflects large differences in what is
in the forefront of other categorisations, as for example access to
economic and other resources, status, social power et cetera. Longitudinal analyses of social classes for such a long period covered in this study involves several problems, related to changes in occupational structure and the meaning and contents of occupations and social positions. We have chosen to work with a broad definition of classes, merging the original 12 classes in Hisclass into the following three and an additional fourth
category with unknown class:

\begin{enumerate}
\item {\tt elite}, Hisclass 1 and 2. Higher managers and higher professionals.
\item {\tt middle}, Hisclass 3, 4, 5, 6 and 8. Lower managers,
farmers and lower professionals, clerical and sales personell.
\item {\tt worker}, Hisclass 7, 9, 10, 11 and 12. Workers including
farm workers.
\end{enumerate}

Figures \@ref(fig:hiscskumm) and \@ref(fig:hiscskumw) show the distribution
(per cent) of exposure time according to class for men and women
respectively. The category of workers is fairly stable over time. The
highest class has increased substantially, although from a very low level
in the 19th century. The increase of academics in connection to the
establishment of Umeå University explains much of this increase. What is
not apparent from Figure \@ref(fig:hiscskumm) is the transfer within the middle
group from a majority of farmers during the 19th century to lower white
collar from the middle of the 20th century. Furthermore we observe an
increase during the 1960s and partly during later decades as well in those
with undefined class. When first observed in the earliest included censuses,
occupation is often missing. For men the proportion is somewhat higher in
the age group 65 and above, i.e. those that were retired, but still low
enough to be used in our analyses. Until 1950, there were however very few
with missing occupation. For men in working-active age, the proportion
missing does not cause any problems for our analyses.

```{r hiscskumm, fig.cap = "Exposure by hisclass, men (from the top: elite, middle, workers, missing) and time period.", fig.height = 4.5}
xx <- with(skel[skel$sex == "male", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xy <- with(skume[skume$sex == "male", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xz <- with(foli[foli$sex == "male", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xxx <- rbind(xx, xy, xz)
yy <- round(prop.table(xxx, 1) * 100, 1)
barplot(t(yy)[4:1, ], col = c("red", "blue", "green", "black"), las = 2, ylab = "Per cent")
```

```{r hiscskumw, fig.cap = "Exposure by hisclass, women (from the top: elite, middle, workers, missing) and time period.", fig.height = 4.5}
xx <- with(skel[skel$sex == "female", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xy <- with(skume[skume$sex == "female", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xz <- with(foli[foli$sex == "female", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xxx <- rbind(xx, xy, xz)
yy <- round(prop.table(xxx, 1) * 100, 1)
barplot(t(yy)[4:1, ], col = c("red", "blue", "green", "black"), las = 2, ylab = "Per cent")
```

The proportion with missing values among women is however a larger problem,
see Figure \@ref(fig:hiscskumw). Corresponding to what we found for men, the
proportion without social position is very low until the middle of the 20th
century, reflecting the way womens social position is defined. Missing social class is however a real problem during particularly the 1960s and 1970s. Around 80 \% lack information for the 1960s and 50 \% for the 1970s. Since social class for this period is defined
according to own occupation, the consequence of many women being outside
the workforce, the proportion missing becomes very high. We have chosen not
to report results for women for these two decades. For the later decades,
the results are fairly complete.

## Marital status

Marital status is treated as a time-dependent variable in the data
file. The statuses used are unmarried, married and dissolved marriages
(without distinguishing between widowed and divorces). If no explicit
status is given in the sources and partner is missing, the status has been
set to unmarried.

Regarding the distribution of marital status over time, see
Figure \@ref(fig:civstskum). The figure indicates an under-registration of
those married and over-registration of unmarried during the
1960s.

```{r civstskum, fig.cap = "Exposure by civil status (from the top: never married, married and previously married) and time period", fig.height = 4.5}
xx <- with(skel, tapply(exposure, list(period = period, civst = civst), sum))
xy <- with(skume, tapply(exposure, list(period = period, civst = civst), sum))
xz <- with(foli, tapply(exposure, list(period = period, civst = civst), sum))
xxx <- rbind(xx, xy, xz)
yy <- round(prop.table(xxx, 1) * 100, 1)
barplot(t(yy)[3:1, ], col = c("red", "blue", "black"), las = 2, ylab = "Per cent")
```

## Cause of death

In the present study, we analyse mortality in the two most common disease
groups in the ages we are investigating; cardiovascular diseases and
cancers. National compulsory reporting of cause of death was introduced in
1911, but our analysis of the two disease groups starts from 1901 since deaths were fairly frequently reported even before 1911. Classification systems have however changed. Sweden did not join ICD
until 1952, but the historical causes of death in the Poplink database have
been coded according to an adapted form of the ICD10 system. In the present study, we identify the codes relevant for cardiovascular diseases and cancers in
the ICD 8, 9 and 10.

The distribution of causes of death is shown in
Figure \@ref(fig:causeskum). Notice the high proportion of other causes in
the first two periods. This category include cases with unknown or
unspecific causes which has implications for the cause-specific analyses
(see later discussion). The total dominance of our two disease groups is
apparent.

```{r causeskum, fig.cap = "Causes of death by time period (top: cancer, middle: cardiovascular, bottom: other).", fig.height = 4.5}
##vb01 <- vb0[vb0$event == 1, ]
cancer <- c(tapply(skume$cancer, skume$period, sum), tapply(foli$cancer, foli$period, sum))
heart <- c(tapply(skume$heart, skume$period, sum), tapply(foli$heart, foli$period, sum))
death <- c(tapply(skume$event, skume$period, sum), tapply(foli$event, foli$period, sum))
other <- death - cancer - heart
ttt <- rbind(cancer, heart, other)
ttt <- prop.table(ttt, 2) * 100
barplot(ttt[3:1, ], col = 3:1, las = 2, ylab = "Per cent")
```

## Urban vs rural residence

In 19th and early 20th century Sweden (as in other countries), mortality
was much higher in urban environments [@hani94, @edv92]. The unsanitary and overcrowded
living conditions as well as the risks for importing infections made the
health risks higher in towns and cities. In the present paper, we control
for this aspect by distinguishing between urban and rural residence for the periods after 1900, based
on parish of residence.

## Periods for analysis

We have made separate analyses for the different periods. The century
1851--1950 is divided into four periods of 25 years each. The 19th century
periods represent a mainly pre-industrial society dominated by agriculture
and with only basic welfare provisions. During the years 1901--1950,
industrialisation started and the urban environments increased in
population size. The modern Swedish welfare state began to develop. For the 
period 1960 onwards, there are separate analyses on decades except for the two 
last periods. Sweden had developed into a wealthy
welfare society with low income inequality,
particularly until the 1990s. A financial crisis struck Sweden in the early
1990s, which among other things increased unemployment and changed parts of
the welfare system towards being less generous. The start of the
last period (2008--2013) coincides with the international financial
crisis. During the last decades income inequality has increased considerably.

# Models

The proportional hazards model in a survival analysis context is used,
allowing for adjustment for marital status and rural/urban environment. Since
the central explanatory variable, social class (Hisclass), do not follow the
proportionality property, the analyses are stratified with respect to the
variable, and the main results are presented graphically.

Another practical problem is the size of that data set:
It contains no less than `r as.integer(round(n.exp + pn.exp))` person years and
 `r n.d + pn.d` deaths.
This makes it very inconvenient to use the nowadays so popular semi-parametric
technique known as
\emph{Cox regression} [@cox72, @gb02]. Instead we use a parametric model
that allows for reduction of the data
to \emph{sufficient statistics} [@fisher22]. The idea is to keep all
relevant covariates \emph{categorical}, and then create a table with all
possible combinations of levels of the covariates. For each such
combination, the total amount of exposure and the number of deaths are
calculated.  We have two sexes, three marital statuses, ten periods, two
urban/rural groups, four social classes at age 40,  four social classes at age 60,
ten age groups, and 22 cohorts, in all 422\,400 possible
combinations. However, in this scheme there are many structural zeroes
(impossible combinations) and some random zeros, so the actual number of
combinations we end up with is 24\,326. The data set is thus reduced in
size by 99 per cent.

The parametric model we use for the baseline hazard function is piece-wise
constant on five-year intervals, so we arrive at the following proportional
hazards regression model:

\begin{equation}\label{eq:prop}
h(t; \lambda, \boldsymbol{\beta}, \textbf{x}) =
\lambda(t)e^{\boldsymbol{\beta} \mathbf{x}}, \; \; 40 < t \le 90,
\end{equation}
where

\begin{equation*}
\begin{split}
\mathbf{x} &= (x_1, \ldots, x_n) \text{\ is a vector of covariates,} \\
\boldsymbol{\beta} &= (\beta_1, \ldots, \beta_n) \text{\ is the corresponding vector of regression
parameters.} \\
\end{split}
\end{equation*}
and
\begin{equation*}
\lambda(t) = \alpha_i,\; 40 + 5(i - 1) < t \le 40 + 5i, \; i = 1, \ldots, 10.
\end{equation*}

Since all components of $\mathbf{x}$ are categorical covariates, the
estimation of \@ref(eq:prop) can be performed in two ways, first by the
original data set and a survival analysis routine handling a piecewise
constant hazard, or second, by the reduced data set (table) and
\emph{Poisson regression}. The second way is faster than the first by a
factor of size at least 100, often much more, depending on hardware, and
the point is that the two ways give exactly the same results. This is
because the table is a \emph{sufficient statistic} for \@ref(eq:prop) and
the original data set.

The analyses are performed in the **R** environment for statistical computing
and graphics [@cran], especially using the package **eha**
[@gb12;@eha]. The proportional hazards model allows us to scrutinise the differences and
the development in more detail. We divide the analysis into two age groups,
40--64 and 65--89 years of age, the first age group representing people that
were still mainly in the workforce, and the second those that mainly were
retired. The model controls for marital status and whether the individual resides 
in the urban or rural parts of the regions. Each period are analysed separately, 
but we then combine results
in order to show the development of the social patterning of mortality
during the studied period.

The important explanatory variable, social class (**hisclass**), is included in the
proportional hazards models as a *stratification* variable. The main
reason for this choice is that the effect of social class (**hisclass**) on survival is
non-proportional, that is, it varies with age. With the reduced data set
and Poisson regression, this is simply achieved by introducing an
interaction term between **age** and social class (**hisclass**). Then, the main
results are presented graphically.  The comments on the results are based on
the figures and focuses on the central variable in our study--- social class (**hisclass**).

# The effect of social class, 1801--1950, 1961--2013

## All causes of death

```{r colorc}
## Colors for plotting strata:
scol <- c("blue", rep("darkgreen", 2), "red")
## Common for all groups:

labb <- levels(vb0Tab$period)
nper <- length(labb)
#source("R/getPHglm.R")
## End 'common'
```

The cumulative
hazards at the end of the time period (65 if 40--64, 90 if 65--89) are the
main target in our analysis: These numbers are used in the process of
graphically illustrating the development over time.

### Ages 40--64

```{r getske4060}
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, women."
scap1w <-  "Cumulative hazards, ages 40--64, women."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, men."
scap1m <-  "Cumulative hazards, ages 40--64, men."
##
cap0 <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom)."
scap0 <- "Total hazard by HISCLASS, ages 40--64."
cap12 <- "Total hazard of dying before age 65 for a 40 year old person standardised by HISCLASS and decade, women (top) and men (bottom)."
scap12 <- "Total hazard standardised by HISCLASS, ages 40--64."
cap12log <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades."
scap12log <- "Total hazard by HISCLASS, ages 40--64, last years."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, women."
scap2w <-  "Cumulative hazards, ages 65--89, women."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, men."
scap2m <-  "Cumulative hazards, ages 65--89, men."
##
cap22 <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom). Standardised."
scap22 <- "Total hazard by HISCLASS, ages 65--89."
cap22log <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades."
scap22log <- "Total hazard by HISCLASS, ages 65--89, last years."

##
vb4064 <- vb0Tab[as.numeric(as.character(vb0Tab$age)) < 64.5, ]
vb4064$age <- factor(vb4064$age)
##vb4064$hisclass <- vb4064$hisclass40
```

```{r hazcalw}
source("R/getPHglm2.R")
##nper <- 4
nper <- length(labb)
resw <- getPHglm2(vb4064[vb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb4064[vb4064$sex == "male", ],
            nper, labb)
```

```{r cumsw, fig.cap = cap0, fig.scap = scap0, fig.height = 8}
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, ylab = "TMR(65|40)",
                main = "Women 40-64")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, ylab = "TMR(65|40)",
                main = "Men 40-64")##, exclude = 5:6)
par(oldpar)
```

From Table \@ref(tab:cumswtab) it is obvious that the statistical significance is strong.
The most interesting feature is the three-way interaction with **sex**, **period**, and social class (**hisclass**).

```{r cumswtab,cap="Table",results="asis"}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex * hisclass * period,
            data = vb4064, family = poisson)
dr <- drop1(fitt, test = "Chisq")
##knitr::kable(dr)
print(xtable(dr,label = "tab:cumswtab", caption = "Anova, all causes of death, ages 40--65."), 
      type = "latex", comment = FALSE)
```

Figure \@ref(fig:cumsw) shows the summed hazards of dying in the age interval
40--64 during the different periods. Obviously, and not surprisingly, mortality
is decreasing over time for both sexes. Women have consistently better survival
than men. This is a well-known phenomenon in Swedish
demographic history. Male mortality has almost always been higher than
female in all age groups [@wil99, @suwi07]. We also observe a
substantial decline in mortality for both sexes, resulting in smaller
absolute differences between social classes. We must however distinguish between differences in absolute risks compared to relative risks [@hoetal16). The compression that is
visible for later decades is partly explained by the decreasing general
mortality and is not necessarily representing a homogenisation of relative risks.

By standardising the levels in each time period for each class, we 
illustrate the relative risks in Figure \@ref(fig:cumswstand). In the rest of 
the paper, we only present these
relative mortality levels for the different classes since they are the main focus in this study.

```{r cumswstand, fig.cap = cap12, fig.scap = scap12, fig.height = 8}
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(65|40)", main = "Women 40-64")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(65|40)", main = "Men 40-64")##, exclude = 5:6)
par(oldpar)
```

Notice that we have not analysed the 1950s due to the fact that we only
have data for some parishes at that time. Furthermore, the
results for women during the 1960s and 1970 are highly misleading for the reasons discussed above. The
changing economy also have consequences for the analysis, particularly when
it comes to the agricultural sector. While farmers were common occupations
up until the first half of the 20th century, these occupations has constantly
decreased in numbers thereafter. This means that the middle class has
somewhat different character during the last periods.

The most striking result of the analysis is the different patterns for men
and women. The social pattern of mortality among women is quite much in
accordance with previous analyses from the present research group on other
regions in northern Sweden and for other periods [@edli11,@edbr12]. The
highest social class usually have comparatively low mortality, while
working class women have high, but about the same level as the middle category. There is however no substantial change in the relative risks between classes. Only the group with unknown social
class is distinguished by much higher mortality from the 1980s
onwards. This can however have to do with how we have defined social
position for women before and after 1950. After 1950 it is based
on their own occupation. Being without occupation probably include women that
are more frail when it comes to health or are vulnerable due to social characteristics.

A completely different pattern appears among men, being the most surprising finding in
our analyses. The groups we would expect to have the lowest mortality, if
we assume that access to important resources and belonging to a high social class
determines survival, instead have by far the highest hazards of dying
during the nineteenth and much of the 20th century. Workers on the
other hand have comparatively low mortality until the 1960s, at the same
level as the middle class. The relative positions change radically towards
the end the 20th century. From the 1980s, there is a clear advantage when
it comes to survival for higher managers and professionals. The relative
differences are however more compressed than for the period until 1950. The
expected social gradient in mortality has developed. This is also what we
usually find in analyses of social health inequalities in present-day
Sweden as well as in most other countries. In the regions we are studying,
this is however a quite recent phenomenon. Corresponding to what we find
for women, those without known occupation have much higher mortality than
the others from the 1960s onwards. The assumption made for women of
unknowns social class is probably even more relevant for men. They probably
represent groups in society that are particularly vulnerable, having difficulties getting into the labour force.

A further observation can be made from Figure \@ref(fig:cumsw). The gender
difference in death hazards is extremely large for the elite group from
1851--1950, men having about three times higher risks than women in the
same class (and in the same households). In the other classes, the difference was much smaller. It was
men in high position that experienced the largest risks.

### Ages 65--89

```{r getske6089}
vb6589 <- vb0Tab[as.numeric(as.character(vb0Tab$age)) >= 64.5, ]
vb6589$age <- factor(vb6589$age) # NOTE:: Change to '6589'!
##vb6589$hisclass <- vb6589$hisclass60
```

We now turn to the elderly, i.e. those that in most cases have left the workforce.
Corresponding to the previous analysis, we use the value of the cumulative hazards at
duration 25 (age 90), starting from age 65 as a summary measure.

The general pattern and development of mortality (not shown here) do not
differ that much from the younger age group, of course representing a much
higher mortality level earlier in history. Mortality among women was, as
expected, lower than among men. We find a clear decline in mortality, but
the fastest decline did not set in definitely until sometimes after the
Second World War, i.e. later than for the age
group 40--64 years. This corresponds to the Swedish development where we
find a turning point at this time with better survival among elderly, to a
large extent caused by lower mortality in cardiovascular diseases.

Concerning the social pattern, the pattern is less clear compared to what
we find for the younger age group (see Figure \@ref(fig:cumswold)). For
women the results are mixed. Once again, too much missing data on social
class force us to leave out the results for the 1960s and 1970s. We do not find the
same clear advantage for elite women during the period before 1950, but neither
do we find any large disadvantage for working class elderly women at that
time. In some periods, the social differences were small, and the elite
even had substantially higher mortality during the early 20th century. The
social differences continued to be small until the last decades of the
20th century. Only during the two last periods (after the year 2000) is it
possible to establish the expected social gradient in mortality.

For men we find indications of the same intriguing change among elderly as
for the adult male population 40--64 years. Mortality was in fact
substantially higher among elite men but their disadvantage was mainly 
apparent during the 19th century. Differences are fairly small during the
first half of the 20th century. Not until the second half of the 20th
century is the survival much better for the highest social class while
workers in the age of retirement had high mortality in comparison.

```{r cumswold, fig.cap = cap22, fig.scap = scap22, fig.height = 8}
##source("R/plotED.R")
resw <- getPHglm2(vb6589[vb6589$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb6589[vb6589$sex == "male", ],
            nper, labb)

oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(90|65)", main = "Women 65-89")##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(90|65)", main = "Men 65-89")##, exclude = 5:6)
par(oldpar)
```

```{r cumswoldtab,results='asis',eval=FALSE}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex * hisclass * period,
            data = vb6589, family = poisson)
dr <- drop1(fitt, test = "Chisq")
##knitr::kable(dr)
xtable(dr,label = "tab:cumswoldtab", caption = "Anova, all causes of death, 65-89.")
```

## Cardiovascular mortality

In this and the following section, the focus is on analyses of mortality from the two
dominant groups of death causes in the adult and elderly population. We start
with cardiovascular diseases. Many studies have shown that socio-economic
conditions are strongly associated with mortality in this disease
group. Belonging to working class, having less education and low income
increases mortality risks. We would therefore expect to find a clear
social pattern in the present analysis.

A general observation is that the results on causes of death and
its development is difficult to interpret, partly because the number of
cases are fewer but also that the quality and coverage of reported death
causes differ in time. Thus, many deaths during the early 20th got quite vague diagnoses reported in the death registers. There was still a substantial part
of all deaths where cause was diagnosed as unknown or dying from old age,
despite the fact that reporting causes of death had become mandatory. For
the period 1901--1925 to 1926--1950, the proportion of deaths from
cardiovascular diseases increased from 25 to 43 per cent for the age group
40 and above, thus indicating an increase of these diseases. However, at
the same time unknown cause and death of old age decreased from 28 percent
to 18 per cent. It is quite reasonable to suspect that many cardiovascular
deaths were disguised in unspecific categories and that the increase is partly
explained by improvements in how deaths were diagnosed, something that also seems to have been the case in Scotland at this time [@reidetal15].

```{r causes41}
cvbTab <- rbind(skume[, names(foli)], foli)
cvbTab$event <- cvbTab$heart
```

```{r 41colorc}
## Colors for plotting strata:
scol <- c("blue", rep("darkgreen", 2), "red")
## Common for all groups:

labb <- levels(cvbTab$period)
nper <- length(labb)
##source("R/getPHglm.R")
## End 'common'
```

As before, models are fitted separately for sex and each ten-year time period, from  1901 to
2013. We stratify on social class (**hisclass**) and include the covariates **urban** (TRUE or FALSE) and
**civst** (marital status). The cumulative
hazards at the end of the age period (65 if 40--64, 90 if 65--89) are the
main target in our analysis. These numbers are used in the process of
graphically illustrating the development over time.

### Ages 40--64

```{r 41getske4060}
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, women, cardiovascular death."
scap1w <-  "Cumulative hazards, ages 40--64, women, cardiovascular death."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, men, cardiovascular death."
scap1m <-  "Cumulative hazards, ages 40--64, men, cardiovascular death."
##
cap12 <- "Total hazard of dying before age 65 for a 40 year old person standardised by HISCLASS and decade, women (top) and men (bottom), cardiovascular death."
scap12 <- "Total hazard by HISCLASS, ages 40--64, cardiovascular death."
cap12log <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cardiovascular death."
scap12log <- "Total hazard by HISCLASS, ages 40--64, last years, cardiovascular death."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, women, cardiovascular death."
scap2w <-  "Cumulative hazards, ages 65--89, women, cardiovascular death."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, men, cardiovascular death."
scap2m <-  "Cumulative hazards, ages 65--89, men, cardiovascular death."
##
cap22 <- "Total hazard of dying before age 90 for a 65 year old person standardised by HISCLASS and decade, women (top) and men (bottom), cardiovascular death."
scap22 <- "Total hazard standardised by HISCLASS, ages 65--89, cardiovascular death."
cap22log <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cardiovascular death."
scap22log <- "Total hazard by HISCLASS, ages 65--89, last years, cardiovascular death."

##
cvb4064 <- cvbTab[as.numeric(as.character(cvbTab$age)) < 64.5, ]
cvb4064$age <- factor(cvb4064$age)
```

Performing the same types of analyses as for all-cause mortality, we start
by looking at cumulative hazards by Hisclass at age 65, given
survival to age 40. The result of the model can be seen in Table \@ref(tab:41cumswtab).

```{r 41hazcalw}
source("R/getPHglm2.R")
resw <- getPHglm2(cvb4064[cvb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(cvb4064[cvb4064$sex == "male", ],
            nper, labb)
```

```{r 41cumsw, fig.cap = cap12, fig.scap = scap12, fig.height = 8}
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE, ##exclude = 3:4,
                    ylab = "TMR(65|40)",  main = "Women 40-64")
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE, ##exclude = 1:2,
                    ylab = "TMR(65|40)", main = "Men 40-64")
par(oldpar)
```

```{r 41cumswtab,results='asis',eval = FALSE}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex * hisclass +
                sex * period + hisclass * period,
            data = cvb4064, family = poisson)
dr <- drop1(fitt, test = "Chisq")
knitr::kable(dr, caption = "Anova, cardiovascular deaths, 40-64.")
#print(xtable(dr,label = "tab:41cumswtab", caption = "Anova, cardiovascular deaths, 40-64."), type = "html")
```

Figure \@ref(fig:41cumsw) shows mortality levels during different periods
from 1901 on-wards. Women from the highest social class had consistently
lower mortality from cardiovascular diseases compared to the others. We
also observe increasing social differences from the first half of the
century to the last decades. A new pattern has developed from 1980
onwards. Now working class women have substantially higher cardiovascular
mortality compared to others. We can however not exclude that this is an
artefact of different diagnostics for deaths in different groups. Working
class women may have had less precise diagnoses in the early 20th century.

For men, there is a different pattern, a pattern however that to a large
extent is consistent with what we observed concerning all-cause
mortality. The elite group had much higher cardiovascular mortality up until
the middle of the 20th century, but instead have the lowest mortality from
1980 onwards. Working class men had a fairly low cardiovascular mortality,
but in the same way as among women, the levels were fairly similar to the
one of the middle category consisting of white collar, lower managers and farmers. In the last decades and similarly as for women, working class men have the highest
mortality from this cause of death group. A social
hierarchy in cardiovascular mortality developed both for men
and women.

### Ages 65--89

```{r 41getske6089}
cvb6589 <- cvbTab[as.numeric(as.character(cvbTab$age)) >= 64.5, ]
cvb6589$age <- factor(cvb6589$age) # NOTE:: Now change to '6589'!
```

In this section we focus on cardiovascular mortality among those that
mostly have left the workforce, the age group 65--89 years. We use as a
summary measure the value of the cumulative hazards at
duration 25 (age 90), starting from age 65. The results from the model is presented in Table \@ref(tab:41cumswoldtab).

```{r 41hazcalwold}
resw <- getPHglm2(cvb6589[cvb6589$sex == "female", ],
              nper, labb)
resm <- getPHglm2(cvb6589[cvb6589$sex == "male", ],
                nper, labb)
```

Concerning the social pattern, we find somewhat different results for elderly women compared to the women in working-active age. Cardiovascular mortality among elderly
women do not present a consistent class pattern. Female cardiovascular
mortality was somewhat higher among the elite and the middle group than
among working class women until the last decades. Elderly from a working
class background were not disadvantaged for most of the 20th century. The
social class pattern has changed somewhat during the last periods becoming more
in line to what is expected. The differences are however rather small.

Even if results for elderly men point in different directions for some
periods, the general trend resembles that of working-active men. The
highest class had high cardiovascular mortality during the first half of
the 20th century, while the expected social pattern developed and became apparent only
during the most recent periods.

```{r 41cumswold, fig.cap = cap22, fig.scap = scap22, fig.height = 8}
##source("R/plotED.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "H(90|65)",  main = "Women 65-89")##, exclude = 3:4)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "H(90|65)", main = "Men 65-89")
par(oldpar)
```

```{r 41cumswoldtab,results='asis',eval=FALSE}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex * period *hisclass,
                ##sex * period + hisclass * period,
            data = cvb6589, family = poisson)
dr <- drop1(fitt, test = "Chisq")
knitr::kable(dr, caption = "Anova, cardiovascular deaths, 65-90.")
##xtable(dr,label = "tab:41cumswoldtab", caption = "Anova, cardiovascular deaths, 65-90.")
```

## Cancer mortality

While we assume clear social differences in cardiovascular diseases, this
is not necessarily an obvious assumption when it comes to
cancer. Different forms of cancers have different origin and
different pathways to disease and death. The social pattern of higher
mortality in classes with less power and accordingly less life chances are
considered to be less clear when it comes to cancers. In our analyses,
there is a risk that cancer mortality is under-estimated during the first
half of the 20th century, illustrated by the already mentioned large
proportion of unknown or vague causes at that time. We believe however that
cancers were better diagnosed than cardiovascular deaths. This is supported
by the stability in proportion of cancer deaths between the two first
periods, 12.5 per cent 1901--1925 and 13.5 per cent for the period
1926--1950.

```{r causes12}
cvbTab <- rbind(skume[, names(foli)], foli)
cvbTab$event <- cvbTab$cancer
```

```{r 12colorc}
## Colours for plotting strata:
scol <- c("blue", rep("darkgreen", 2), "red")
## Common for all groups:

labb <- levels(cvbTab$period)
nper <- length(labb)
source("R/getPHglm.R")
## End 'common'
```

We use similar types of analyses as in the earlier sections. The cumulative
hazards at the end of the period (65 if 40--64, 90 if 65--89) are the
main target in our analysis. These numbers are used in the process of
graphically illustrating the development over time.

### Ages 40--64

```{r 12getske4060}
cap1w <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, women, cancer death."
scap1w <-  "Cumulative hazards, ages 40--64, women, cancer death."
cap1m <-  "Cumulative hazards for HISCLASS by time period, ages 40--64, men, cancer death."
scap1m <-  "Cumulative hazards, ages 40--64, men, cancer death."
##
cap12 <- "Total hazard of dying before age 65 for a 40 year old person standardised by HISCLASS and decade, women (top) and men (bottom), cancer death."
scap12 <- "Total hazard by HISCLASS, ages 40--64, cancer death."
cap12log <- "Total hazard of dying before age 65 for a 40 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cancer death."
scap12log <- "Total hazard by HISCLASS, ages 40--64, last years, cancer death."

cap2w <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, women, cancer death."
scap2w <-  "Cumulative hazards, ages 65--89, women, cancer death."
cap2m <-  "Cumulative hazards for HISCLASS by time period, ages 65--89, men, cancer death."
scap2m <-  "Cumulative hazards, ages 65--89, men, cancer death."
##
cap22 <- "Total hazard of dying before age 90 for a 65 year old person standardised by HISCLASS and decade, women (top) and men (bottom), cancer death."
scap22 <- "Total hazard standardised by HISCLASS, ages 65--89, cancer death."
cap22log <- "Total hazard of dying before age 90 for a 65 year old person by HISCLASS and decade, women (top) and men (bottom). The last decades, cancer death."
scap22log <- "Total hazard by HISCLASS, ages 65--89, last years, cancer death."

##
cvb4064 <- cvbTab[as.numeric(as.character(cvbTab$age)) < 64.5, ]
cvb4064$age <- factor(cvb4064$age)
```

```{r 12hazcalw}
source("R/getPHglm2.R")
resw <- getPHglm2(cvb4064[cvb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(cvb4064[cvb4064$sex == "male", ],
            nper, labb)
```

```{r 12cumsw, fig.cap = cap12, fig.scap = scap12, fig.height = 8}
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE, ##exclude = 3:4,
                    ylab = "TMR(65|40)",  main = "Women 40-64")
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE, ##exclude = 1:2,
                    ylab = "TMR(65|40)", main = "Men 40-64")
par(oldpar)
```

```{r 12cumswtab,results='asis',eval=FALSE}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex *hisclass +
                sex * period + hisclass * period,
            data = cvb4064, family = poisson)
dr <- drop1(fitt, test = "Chisq")
knitr::kable(dr, caption = "Anova, cancer deaths, 40-64.")
##xtable(dr,label = "tab:12cumswtab", caption = "Anova, cancer deaths, 40-64.")
```

Figure \@ref(fig:12cumsw) shows relative mortality levels according to
Hisclass during the different studied periods for the age group
40--64. Cancer mortality among elite women was lower than that of the
other classes during the complete studied period, while working class women
had the highest levels during almost all periods. Consequently, we find a
more consistent advantage of higher class in cancers compared to
cardiovascular diseases for women, consistent with the idea of a social hierarchy in health. What is less expected, is that this we find a stronger association between social class and survival in cancer than cardiovascular diseases. We do however not find the same pattern
for men. It is difficult to identify any general trends and features of 20th century
male cancer mortality when it comes to social differences.

### Ages 65--89

```{r 12getske6089}
cvb6589 <- cvbTab[as.numeric(as.character(cvbTab$age)) >= 64.5, ]
cvb6589$age <- factor(cvb6589$age) # NOTE:: Now change to '6589'!
```

In our analyses of class patterns in mortality, we finally turn to cancer
mortality among the elderly. The model is presented in Table \@ref(tab:12cumswoldtab).

```{r 12hazcalwold}
resw <- getPHglm2(cvb6589[cvb6589$sex == "female", ],
              nper, labb)
resm <- getPHglm2(cvb6589[cvb6589$sex == "male", ],
                nper, labb)
```

The relation between different classes and cancer mortality fluctuates
considerably both for women and men in this age group. For some periods it
has to do with quite few events to analyse. Women of the highest class have
favourable mortality levels during several periods, consistent with the advantages
we found for women in the younger age group, but on the other hand had high
levels at other times. The social differences are however quite small
during the most recent periods.

For men we find a similar lack of social pattern in cancer mortality as we
found for men in working-active ages. There is however a tendency during
the last decades of higher mortality among working class men.

```{r 12cumswold, fig.cap = cap22, fig.scap = scap22, fig.height = 8}
##source("R/plotED.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(90|65)",  main = "Women 65-89")##, exclude = 3:4)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, relative = TRUE,
                    ylab = "TMR(90|65)", main = "Men 65-89")
par(oldpar)
```

```{r 12cumswoldtab,results='asis',eval=FALSE}
fitt <- glm(event ~ offset(log(exposure)) + urban + civst + age + sex +
                sex * period + hisclass * period,
            data = cvb6589, family = poisson)
dr <- drop1(fitt, test = "Chisq")
knitr::kable(dr, caption = "Anova, cancer deaths, 65-90.")
##xtable(dr,label = "tab:12cumswoldtab", caption = "Anova, cancer deaths, 65-90.")
```

\section{Discussion}

The authors of this paper have scrutinised different aspects of the social
pattern of mortality in historical northern Sweden for some time, focusing
on different regions, age groups and time periods [@edv92, @edv04, @edli11, @edbr12]. When it comes to infant and child mortality during pre- and early-transitional time, the expected socio-economic
differences was often but not always confirmed [@edv92, @edv04],
something that has been found in other Swedish studies as well [@bredro02,
@bube01, @mol17].

Studies based on microdata on the adult population in
history have been rare, mainly due to lack of adequate data. The results
from such studies show other patterns than that for infant and child
mortality. The expected social gradient in survival, meaning that people
from higher social classes always had an advantage, was in many cases,
particularly among adult men, not found despite the well-documented
socio-economic differences in society [@edli11, @edbr12]. The results presented here show that the consistent advantage for higher social classes do not appear until quite late in the investigated regions.

The results showing a small or even reversed socio-economic gradient in the past and its late development are however in accordance with other studies [@bevp11], for example by @bedr11 in a different setting, a couple of Scanian parishes in southern Sweden, although with somewhat different social categories. They find the first indications of an emerging social gradient in the 1950s. Similar results have also been found in Estonia [@jaadla17]. A study of mortality and social class in Sweden 1960--1968 found that self-employed men and non-manual male workers in fact had higher mortality than manual workers in total and cardiovascular mortality [@vano89]. These findings are however contrary towards what usually is assumed and what other studies have shown. Those studies usually referred to when discussing the historical large social inequalities in health are mainly based on data from England and Wales [@ant67, @wowi95 @pamuk85]. We know less about how general these results are, but apparently the Swedish context has been different from the English and Welsh.

The surprising results make it reasonable to consider the validity and possible weaknesses in the study. Some of the limitations have been addressed in the presentation of the variables: The fact that social class was defined according to household until 1950 but individually defined from 1960 (affecting the social class of women), different definitions of presence during the same two periods and some differences in parishes included at different periods. These differences are however not of the magnitude that the main conclusions are affected. Most women would end up in the same category as their husbands and the migrations between census dates or census date and death were not frequent.

One obstacle concerns the quality of available information. Occupational titles in historical sources are not always distinct enough to define the correct class. Studies covering long periods face the problem that occupations and classes change character --- they do not represent the same thing. We have chosen to work with a few broad groups. Nevertheless, these ambiguities and possible misclassifications should not radically change the basic patterns observed. The social classes identified during the different periods represent distinct groups when it comes to resources, capacities and power.

In this paper we have analysed social inequality in health from the
perspective of social classes. However, there are
alternative ways to analyse social inequality, and we cannot exclude that
such analyses give other results or at least would modify the conclusions suggested here. In many modern studies categorisations based on income, wealth or education are common for studies of inequality. They represent different dimensions of inequality. It would be interesting to compare the results from this study with alternative measures, but reliable
data on income and education are unfortunately rarely available in historical population
databases.

How does the results correspond to research on other health measures? In a study of southern Sweden, @oberg14 finds large social differences in body height, a measure that is assumed to sum up health during the life span. This does not contradict what we find here. Height is mainly determined by conditions in early years, and are not reflecting inequalities in adulthood. Earlier studies in Sweden and in the areas studied here [@bredro02, @bube01, @mol17, @edv04] indicate that social disadvantage of being in lower social position were strong in childhood.

Our study seems to be in conflict with many of the dominating theories in present—day health research. In an overview of explanations for the persistence of health inequalities (primarily focusing on high-income countries with extensive welfare arrangements), @mack12 identifies a couple of theories, emphasizing different mechanisms and circumstances such as psycho-social pathways, material conditions, cultural capital et cetera. Most of these explanations are however based on the assumption that higher social classes are always advantaged when it comes to health and mortality. One of these leading theories is the hypothesis of @liph95 that social conditions are fundamental causes of disease and not only expressions of different disadvantages connected to the social position. Others emphasize psycho-social factors, for example @mar04, who argues that a higher status leads to better health regardless of where in the status ladder individuals are situated. Psychosocial factors are also suggested by @wipi09 as an explanation why equal societies always perform better than unequal ones when it comes to health and other social conditions. In the long-term perspective provided here, neither of these theories fits the observation in earlier history. Their validity are mainly apparent for the most recent decades. To be fair, however, our study do not explicitly test for the possible impact of different material or psycho-social mechanisms. These factors could still be valid and influence the social pattern of mortality. What our study establishes is that there there is no necessary eternal social gradient favouring higher social classes. The lacking social hierarchy may be caused by other mechanisms having stronger impact on the pattern. We might have to consider the possibility of socially countervailing mechanisms [@lufr05].

The historical perspective shows the complexity of how health and mortality is socially constructed and illustrates the need to contextualize our analyses. The dominant theories are not necessarily wrong but they may need to be further developed and refined. The WHO Commission on Social Determinants of Health framework provides a holistic model on how socio-economic position influences health. Social class works through intermediate determinants (material circumstances, behaviours, biological factors and psycho-social factors) related to exposure and vulnerability [@who10]. People are situated within socio-economic and political contexts, as well as prevailing societal value systems. The importance of understanding people in their context is also emphasized by @kri05. What she calls embodiment refers to the recognition that human beings are simultaneously social beings and biological organisms. Having a good economy gives advantages in many aspects of relevance for health and mortality, for example good living conditions, access to food, medical care and better working conditions, but even though these advantages were at hand in our studied regions, it did not result in better health for men in the most privileged groups until the late 20th century. It is possible that they had a stressful life, for example entrepreneurs facing the risk of losing an advantagous position. We must however consider other and potentially more important explanations as well.

This makes us consider the possible impact of behavioural factors, something that could represent mechanisms related to stressful lives. Our results of the gender aspect is important in this respect. As several researchers have pointed out, demographic research has often ignored power relations and different roles within families [@watkins93, @mack95, @jans07]. The gender difference in mortality may be caused by discrimination on the female part in the household. We know that historical Sweden was strongly gender segregated, discriminating women and giving them less space for action. In our study as well as in a previous study [@edli11], we do not find this to have negative health consequences for women. Within the same households, the effect of social class leads to different results, high social position being beneficial for women but not for men. Instead, we might consider the role of behaviours based on distinct gendered and class expectations. Having a high social status is sometimes expressed in attitudes, norms and way of living potentially harmful for health. @edli11 suggest that an explanation for the larger social health inequalities in present-day society is that high status is expressed through health, being fit and physically active, while in many historical contexts higher social status (for men) was expressed in low physical activity, obsessive eating, drinking, smoking and sometimes also taking risks.  Alcohol has been suggested as part of the large gender differences in the 19th century [@wil99, @edv92], and a more risk-taking life may be an important component of male mortality [@cour00]. Different historical contexts promote different ways to spend economic resources (compare @rasp06 and @jaadla17.

In the present-day high-income world including Sweden, smoking and drinking contributes to a substantial part of the socio-economic health differences [@mapemy14, @syla16]. Our study suggests a strong impact also in historical Sweden. If this interpretation is correct, the hypothesis of @ant67 on an increasing role of health related behaviour in the modern low-mortality régime, while economic inequality determined the social mortality pattern during the early phases of the mortality transition can be questioned. Instead, health-related can be equally strong earlier in history but expressed in a completely other way. To corroborate this claim, we would need to have evidence on the social differentiation of consumption patterns and behaviours in history. Unfortunately, such data are rarely available before the 1960’s. At best, we have circumstantial and anecdotal evidence. Such evidence can however support the idea of less health-beneficial ways of living among elite men (@edli11 ).

Overweight, obesity and physical inactivity were probably more common in higher social classes. @vano89 find a higher total as well as cardiovascular mortality among self-employed men and non-manual male workers, something they suggest can be explained the higher tobacco consumption among men in these groups. The social pattern among women was however reversed, something that is in line with the present study. There is no solid information on smoking in different social classes before the 1960, but there are international studies indicating higher prevalence and consumption of tobacco in higher classes during the early phases of the tobacco epidemic, illustrated for example by the fact that smoking initially was more prevalent among physicians. There was also an association between higher incomes and consumption [@loco94].

The same lack of statistical evidence characterizes knowledge about alcohol consumption in different social classes before 1960s, but with some exceptions. Access to alcohol was much easier for the wealthy during the period of strong alcohol restriction 1915-1955 and the purchase of alcoholic beverages was substantially higher among those with highest incomes. There was also an over-representation of alcohol-related deaths in higher social classes in the 1960 [@noro98]. @noro98 suggest that during the last decades of the 20th century, there has been an equalization and even reversal of alcohol consumption between social strata, with higher levels among manual labourers but also with a more skewed consumption distribution, indicating more high-consumers in this category. A similar reversal has been observed in England and Wales during the 20th century when it comes to cirrhosis mortality, maybe reflecting a change in consumption patterns. Mortality from this disease was substantially higher in the managerial and professional classes in 1921, being inconsistent in 1961 and reversed in 1991 with highest mortality in the lowest class [@cropr10]. If these indications of changes in health-related behaviour are significant enough to explain our findings here can be discussed. We believe that they certainly are at least an important part of the story.

\section{Conclusion}
In this article we set our task to analyse social differences in mortality from a long-term perspective. Until recently it has not been possible to investigate the development on mortality differences among adult and elderly populations. Still, the prevailing idea in many parts of the scholarly society is that strong socio-economic differences in health and mortality have always existed and were possibly even stronger in historical societies due to the large class differences. The present study is unique in its time perspective by investigating the development from the 19th century until the present. The results probably surprise most readers, and partly surprised us as well. The consistent advantage of belonging to the elite have not always existed and the modern pattern of social divide in health appeared much later than expected in the regions we have studied. Apparently high social class is not always beneficial for survival.

There were however large differences in the pattern between men and women as well as between different age groups. Class as well as gender must be considered. The social pattern of female mortality are in line with the expected, but a social gradient in male adult mortality is a recent phenomenon. Furthermore, inequality in mortality appears clearer for the working—age population, while the differences are smaller among the elderly. Concerning specific causes of death, the analysis shows that cardiovascular diseases contributes to a large proportion of all deaths during the 20th century, and we are not surprised that it has a fairly similar pattern as for all-cause mortality. Cancer is often considered less connected to social class, and we find mixed results in our analyses.

To explain the observed differences is difficult and we can only provide tentative hypotheses. We have discussed the possible role of behavioural influences on mortality in different historical context, suggesting that such aspects are not only important in the modern epidemiological regime but might have been decisive in earlier phases of the epidemiologic transition as well. The results indicate that we should consider the role of gender and class expectations of behaviour. Unfortunately, the evidence for behavioural differences in history are usually circumstantial and anecdotal, at least when it comes to social class. Studies of these aspects and their role for health and mortality in different times and places require further development.

An intriguing aspect of this study is the lack of association between level of economic inequality and social differences in mortality. Even though our study do not directly address the hypothesis raised by [@wipi09], our results are of interest for this question as well. The early 20th century, in Sweden as well as in other countries, was characterised by very large economic differences, still this did in our regions not lead to any consistent advantage for the highest social class. After the Second World War, economic differences diminished, and was at its lowest in Sweden at about 1980, i.e. at a time when a more expected pattern of a social hierarchy in survival appeared. The period thereafter is characterized by increasing economic differences, coinciding with increasing health differences as well. The long-term relation between economic inequality and health require further studies.

An important question is if our results can be generalised to other historical contexts. We have studied environments in northern Sweden without any large metropolis. Perhaps Sweden and/or this part of Sweden is a special case. The involvement of government both at the national and local level may have been stronger, thus mitigating socio-economic disadvantages. The level of income inequality was certainly lower in our regions than in other parts of the country but it was still substantially higher in 1930 than towards the end of the 20th century. In comparison with many other countries, the level of mortality has been fairly low. This is a relevant question, which hopefully can be addressed in other studies, maybe with a comparative approach.

# Acknowledgements

This work was undertaken within the research program *Ageing and
Living Conditions* at the Centre for Demographic and Ageing Research
(CEDAR) at Umeå University, financed by the Swedish Research Council (grant
No 349-2008-6592) and the project *Old age, health and inequality:
Mortality from a life course perspective in Sweden 1986--2009*, financed by
the Swedish Research Council (grant No 421-2012-892). We thank the
Demographic Data Base, Umeå University, for giving us access to data and
Jonas Helgertz, Department of Economic History, Lund University, for
providing us with scripts for creating Hisclass from Swedish Census
data.

# References