### 2017-10-05 cape.Rnw
first commit

\documentclass[t]{beamer}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{natbib}

%\usepackage[figuresonly]{endfloat}

%\bibliographystyle{apalike}

\usepackage{color}

\newcommand{\emp}[1]{\textcolor{blue}{#1}}

\usetheme{Singapore}% best choice.
  \setbeamercovered{transparent}%

\newcommand{\R}{{\bf R}}
\newcommand{\code}[1]{{\tt #1}}

\title{Increasing the mortality gap? Social inequality in mortality among 
elderly and adults in Northern  Sweden 1851--2013\footnote{Session "Health, 
morbidity, and mortality" at the Nordic Demographic Symposium 14--16 June 2017, 
Turku, Finland}
}

\date{\today}

\author{Sören Edvinsson and Göran Broström \\
Centre for Ageing and Demographic Research \\
Umeå University, Umeå, Sweden}

\begin{document}

\maketitle

<<setup, include = FALSE>>=
library(knitr)
opts_chunk$set(engine='R',dev='pdf',fig.path='figs/ume-',tidy=FALSE,fig.height=5.5,fig.width=8,strip.white=TRUE,cache=TRUE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
@

<<environ>>=
library(knitr)
library(eha)
library(xtable)
library(skum)
@

<<readalldata>>=
source("R/get.file.R")
skel <- get.file("skel")
##
skume <- get.file("skume")
##
foli <- get.file("foli")
skel$urban <- FALSE
vb0Tab <- rbind(skel, skume[, names(skel)], foli[, names(skel)])
vb0Tab$age <- factor(vb0Tab$age)
@

\begin{frame}{Aims}

To investigate the association between social class and mortality among adults and elderly from the 19th century to the present and how it is connected to socio-economic inequality in society.
This is analysed according to 
\begin{itemize}
\item gender
\item age group (working age vs retired)
\item *cause of death (cardiovascular, cancer)
\item *education 1990--2013)
\item *income 1986-2005)

% \item Has \emp{inequality in mortality} between \emp{social classes}
%   increased in the adult and elderly
%   population?

%%\item Is the mortality transition initiated by a certain social class?

% \item Are there any \emp{gender differences} in the effect of social position?
% 
% \item Is social position \emp{equally important} among the \emp{retired} population as
%   for those in \emp{working age}?
% \item Is there a difference between classes in mortality from
%   \emp{cardiovascular} diseases and \emp{cancers} respectively?
% \item Are there large differences in survival depending on level of
%   \emp{education} and \emp{income} categories?

\end{itemize}

\end{frame}

\begin{frame}{Main results}

\begin{itemize}
\item A consistent mortality difference according to social class appears surprisingly late in our regions and has no clear connection to the development of income inequality.
\item However, there is a strong gender difference in the development of social inequality.
\item We argue that health-related behavior is an important aspect in historical as well as present-day mortality, and that such behaviours are based on strong gender and class expectations.
\end{itemize}

\end{frame}

\begin{frame}{Economic inequality (figure from Roine and Waldenström 2008)}

\includegraphics[height=8cm]{myfigs/Top_incomes.jpg}

\end{frame}

% \begin{frame}{Outline}
% 
% \begin{center}
% \begin{enumerate}
% 
% \item \emp{Data} management
% \item \emp{Modeling} considerations
% \item \emp{Analysis}
% \item \emp{Results}
% \item \emp{Implications} of the results
% \end{enumerate}
% 
% \end{center}
% 
% \end{frame}

\begin{frame}{Two regions in Västerbotten}

\begin{columns}

\begin{column}{0.5\textwidth}

\includegraphics[height=7cm]{myfigs/sweden.pdf}

\end{column}

\begin{column}{0.5\textwidth}

Remote area in northern Sweden, dominated by agriculture until well into the 20th century.

Skellefteå:
\begin{itemize}

\item Small town with rural hinterland, industrialization of the region during the 20th century.

\end{itemize}

Umeå:

\begin{itemize}

\item Center for administration, education, and military. University from 1965

\end{itemize}

% Rural areas:
% 
% \begin{itemize}
% 
% \item Very similar: Small farms dominating.
% 
% \end{itemize}

\end{column}
\end{columns}
%\includegraphics[height=7cm]{myfigs/sweden.pdf}

\end{frame}

\begin{frame}[fragile]{Data sources}

\begin{itemize}
\item \emp{Swedish church book} records (1851--1950).
  \begin{itemize}
\item digitized by the \emp{Demographic Data Base} at Umeå University,
\begin{itemize}
  % \item \emp{\sc Popum} and
   \item \emp{\sc Poplink}
\end{itemize}
\item  \url{http://www.cedar.umu.se/}
\end{itemize}
\item The \emp{Linnaeus} data base (1960--2013)
  \begin{itemize}
\item Census data 1960, 1965, \ldots, 1990
\item Population registers ({\sc Lisa}, 1986--2013)
  \begin{itemize}
    \item \emp{Income}
    \item \emp{Education}
  \end{itemize}
 \item Death information from \emp{National Board of Health and Welfare}
   \begin{itemize}
\item Death dates from 1 January 1961 to 31 December 2013.
  \item \emp{Causes of death} ({\sc Ulorsak}, ``main cause of death'')
   \end{itemize}
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}{The sampling frame}
{Two separate datasets}

<<lexis2>>=
plot(c(1801, 1801, 1951, 1951, 1801), c(40, 90, 90, 40, 40),
     type = "l", xlim = c(1800, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 90))
axis(1, at = c(1801, 1851, 1901, 1951, 1976, 1991, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1976, 1976, 2014, 2014, 1976), c(40, 90, 90, 40, 40),
      col = "blue", lwd = 2)
#
pn.exp <- round(sum(skel$exposure) + sum(skume$exposure))
pn.d <-  as.integer(sum(skel$event) + sum(skume$event))
#
text(1875, 70, "PopLink", cex = 2, col = "red")
##text(1899, 60, paste(pn.rec, "records"))
text(1875, 55, paste(pn.exp, "p-years"))
text(1874.5, 50, paste(pn.d, "deaths"))
#
n.exp <- round(sum(foli$exposure))
n.d <-  as.integer(sum(foli$event))

text(1995, 70, "Linnaeus", cex = 1.5, col = "blue")
##text(1984, 60, paste(n.rec, "records"))
text(1995.2, 55, paste(n.exp, "p-years"), cex = 0.8)
text(1995.2, 50, paste(n.d, "deaths"))
@
\end{frame}

\begin{frame}{Variables}

\begin{description}

\item[Sex]
\item[Marital status]
\item[Birth date] By quarter (15 Feb, May, Aug, Nov)
\item[Death date] By day (\emp{Socialstyrelsen}, \emp{Poplink})
\item[*Death cause] from \emp{Socialstyrelsen} and \emp{Poplink}. Adapted to ICD10
\item[Parish] Urban/Rural
\item[Social class] Elite, middle, worker. Household definition 1851--1950. Individual definition 1960--2013
\item[*Income] 1986--2005, grouped by quartiles.
\item[*Education] 1990--2013, three categories.
\item[Period]  1851--1875, \ldots,  2008--2013 (10 periods). No results presented for women 1961-70, 1971-80 due to missing information on class.
\item[Age group] Separate analyses for working active (40-65) vs retired (65-89) population
%\item[cohort] 1761--1771, \ldots, 1971--1974 (22 birth cohorts).

\end{description}

\end{frame}

\begin{frame}{\sc Social class}

\begin{enumerate}
\item \emp{\tt elite}, HISCLASS 1 and 2. Higher managers and professionals.
\item \emp{\tt middle}, HISCLASS 3, 4, 5, 6, and 8. Lower managers, lower white collar, and farmers.
%%\item \emp{\tt lowWhiteC}, HISCLASS 4, 5 and 6. Lower white collar.
\item \emp{\tt worker}, HISCLASS 7, 9, 10, 11 and 12. Workers of different skills including farm workers.
\end{enumerate}

\end{frame}

\begin{frame}{No proportional hazards (wrt class)!}

\begin{itemize}
\item Fit each social class \emp{separately}, for each time period.
\item Use the \emp{cumulative hazard} at 65 (or 90) for comparison.
\end{itemize}

So, we fit separate PH models for each combination of

\begin{itemize}
  \item \emp{\tt sex}
  \item \emp{\tt period}: 1851--75, 1876--1900, 1901--10, etc.
  \item \emp{\tt class}
  \item \emp{\tt age group}: 40--64, 65--89.
  \item \emp{\tt cause of death}
\end{itemize}
\end{frame}

\begin{frame}{Women and men, age 40--64}

%  \begin{center}
%\includegraphics[height = 7.5cm]{figure/cumsw-1}
%  \end{center}

<<hazcalw>>=
source("R/getPHglm2.R")
##nper <- 4
vb4064 <- vb0Tab[as.numeric(as.character(vb0Tab$age)) < 64.5, ]
vb4064$age <- factor(vb4064$age)
labb <- levels(vb0Tab$period)
nper <- length(labb)
resw <- getPHglm2(vb4064[vb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb4064[vb4064$sex == "male", ],
            nper, labb)
@

<<cumsw>>=
source("R/plotED.glm.R")
scol <- c("blue", "darkgreen", "red", "black")
oldpar <- par(mfrow = c(2, 1), cex = 0.7, lwd = 1.5)
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, ylab = "TMR(65|40)",
                main = "Women 40-64", exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, ylab = "TMR(65|40)",
                main = "Men 40-64")##, exclude = 5:6)
par(oldpar)
@

\end{frame}

\begin{frame}{Relative differences, ages 40--64 and 65--89}

  %\begin{center}
\includegraphics[height=7.5cm,width=5.4cm]{figs/art2-cumswstand-1}\includegraphics[height=7.5cm,width=5.4cm]{figs/art2-cumswold-1}
   %\end{center}
\end{frame}

% \begin{frame}[fragile]{Education 1990--2013}
% 
% <<educ>>=
% 
% educ <- lisaold[!(lisaold$period %in% c("1986-1987", "1988-1989")), ]
% educ$period <- factor(educ$period)
% educ <- educ[!is.na(educ$educ), ]
% levels(educ$educ) <- c("low", "low", "middle", "middle", "middle", "high", "high")
% educ$educ <- relevel(educ$educ, ref = "middle")
% educ$civst <- relevel(educ$civst, ref = "married")
% fitm <- coxph(Surv(enter, exit, event) ~ strata(educ) + civst + urban,
%              data = educ[educ$sex == "male", ])
% fitf <- coxph(Surv(enter, exit, event) ~ strata(educ) + civst + urban,
%              data = educ[educ$sex == "female", ])
% cols <- c("darkgreen", "red", "black")
% @
% 
% <<ploteduc>>=
% oldpar <- par(mfrow = c(1, 2))
% 
% plot(survfit(fitm, start.time = 40), mark.time = FALSE, main = "Men",
%      col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
% axis(1)
% axis(2, at = c(0, 0.5, 1.0), las = 1)
% abline(h = 0.5, lty = 3)
% box()
% legend("bottomleft", legend = levels(educ$educ), col = cols, lty = 1, lwd = 1.5)
% 
% plot(survfit(fitf, start.time = 40), mark.time = FALSE, main = "Women",
%      col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
% axis(1)
% axis(2, at = c(0, 0.5, 1.0), las = 1)
% abline(h = 0.5, lty = 3)
% box()
% legend("bottomleft", legend = levels(educ$educ), col = cols, lty = 1, lwd = 1.5)
% 
% par(oldpar)
% @
% 
% \end{frame}

% \begin{frame}{Income 1986--2005}
% 
% <<femeduc>>=
% income <- lisaold[!is.na(lisaold$income), ]
% income$period <- factor(income$period)
% ##income <- income[!is.na(income$income), ]
% ##levels(income$income) <- c("low", "low", "middle", "middle", "middle", "high", "high")
% income$income <- relevel(income$income, ref = "low")
% income$civst <- relevel(income$civst, ref = "married")
% fitm <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
%              data = income[income$sex == "male", ])
% fitf <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
%              data = income[income$sex == "female", ])
% cols <- c("darkgreen", "red", "blue", "black")
% 
% @
% 
% <<plotincome>>=
% oldpar <- par(mfrow = c(1, 2))
% 
% plot(survfit(fitm, start.time = 40), mark.time = FALSE, main = "Men",
%      col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
% axis(1)
% axis(2, at = c(0, 0.5, 1.0), las = 1)
% abline(h = 0.5, lty = 3)
% box()
% legend("bottomleft", legend = levels(income$income), col = cols, lty = 1, lwd = 1.5)
% 
% plot(survfit(fitf, start.time = 40), mark.time = FALSE, main = "Women",
%      col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
% axis(1)
% axis(2, at = c(0, 0.5, 1.0), las = 1)
% abline(h = 0.5, lty = 3)
% box()
% legend("bottomleft", legend = levels(income$income), col = cols, lty = 1, lwd = 1.5)
% 
% par(oldpar)
% @
% 
% \end{frame}

%%% Stop here!

\begin{frame}{Summarizing the results}

%\scriptsize

\begin{itemize}
\item \emp{Improved} survival characterised all social classes.
\item \emp{Late} appearance of the “modern” pattern of social differences in mortality among men
\item Distinct \emp{gender} differences in the development.
\item The \emp{upper} class(es) gradually take over as the \emp{most healthy} group.
% \item No \emp{qualitative} but partly \emp{quantitative} differences between those in working age compared to the retired population.
% \item \emp{Educational} and \emp{income} differences during the last decades conform to what we found for class differences.
\item In the long run, no clear connection in the development of income inequality and social differences in mortality.
\item This suggests an important role of gender and class related expectations on health behaviour
\end{itemize}

\end{frame}

\begin{frame}{Some implications}

%\scriptsize

\begin{itemize}
\item The \emp{development} of social health inequalities and the role of \emp{health-related behavior} in different phases of the transition (Antonovsky and Omran).
\item Inequality as a \emp{contextual} health risk (Wilkinson/Pickett).
\item \emp{Status syndrome} (Marmot)
\item The role of \emp{living standard} for mortality decline (McKeown).
\item \emp{Gender} and mortality (Watkins, Mckinnon, Janssens).
\item Social conditions as \emp{fundamental causes of death} (Link/Phelan).

\end{itemize}

\end{frame}

%\begin{frame}{Conclusion}

%\scriptsize

% \begin{itemize}
% \item \emp{Improved} survival characterised all {\sc Hisclass}es.
% \item There are \emp{gender differences} in the social pattern
% \item No difference in the social pattern in the two studied age groups.
% \item The \emp{upper} class(es) gradually take over as the \emp{most healthy} group.
% \item \emp{Educational} and \emp{income} income differences during the last decades conform with what we found for class differences.
% \end{itemize}

%\end{frame}

\end{document}