### 2017-10-24 cape.Rnw
Adding and correcting slides

\documentclass[t]{beamer}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{natbib}

%\usepackage[figuresonly]{endfloat}

%\bibliographystyle{apalike}

\usepackage{color}

\newcommand{\emp}[1]{\textcolor{blue}{#1}}

\usetheme{Singapore}% best choice.
  \setbeamercovered{transparent}%

\newcommand{\R}{{\bf R}}
\newcommand{\code}[1]{{\tt #1}}

\title{Life course and long-term perspectives of social inequality in mortality among 
elderly and adults in Northern  Sweden 1801--2013\footnote{Session "Social inequalities in health at older ages: life course perspectives" at the XXVIII International Population Conference 29 October--4 November 2017, Cape Town, South Africa}
}

\date{\today}

\author{Sören Edvinsson and Göran Broström \\
Centre for Demographic and Ageing Research \\
Umeå University, Umeå, Sweden}

\begin{document}

\maketitle

<<setup, include = FALSE>>=
library(knitr)
opts_chunk$set(engine='R',dev='pdf',fig.path='figs/ume-',tidy=FALSE,fig.height=5.5,fig.width=8,strip.white=TRUE,cache=FALSE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
@

<<environ>>=n
library(knitr)
library(eha)
library(xtable)
library(skum)
@

<<readalldata>>=
source("R/get.file.R")
skel <- get.file("skel")
##
skume <- get.file("skume")
##
foli <- get.file("foli")
skel$urban <- FALSE
vb0Tab <- rbind(skel, skume[, names(skel)], foli[, names(skel)])
vb0Tab$age <- factor(vb0Tab$age)
vb0Tab$civst <- relevel(vb0Tab$civst, ref = "married")
@

\begin{frame}{Aims}

To investigate the association between social class and mortality among adults and elderly from the 19th century to the present and how it is connected to socio-economic inequality in society.
This is analysed according to 
\begin{itemize}
\item gender
\item age group (working age vs retired)
\item *cause of death (cardiovascular, cancer)
%\item *education 1990--2013)
%\item *income 1986-2005)

% \item Has \emp{inequality in mortality} between \emp{social classes}
%   increased in the adult and elderly
%   population?

%%\item Is the mortality transition initiated by a certain social class?

% \item Are there any \emp{gender differences} in the effect of social position?
% 
% \item Is social position \emp{equally important} among the \emp{retired} population as
%   for those in \emp{working age}?
% \item Is there a difference between classes in mortality from
%   \emp{cardiovascular} diseases and \emp{cancers} respectively?
% \item Are there large differences in survival depending on level of
%   \emp{education} and \emp{income} categories?

\end{itemize}

\end{frame}

\begin{frame}[fragile]{Two time perspectives}

\begin{itemize}
\item Development over \emp{time}.
  \begin{itemize}
\item Explicit or implicit assumption of an \epm{eternal} social gradient in history in many theories (Omran, Link and Phelan, Marmot, McKeown), some implying even larger in history.
\item Antonovsky – divergence followed by convergence
\end{itemize}
\item Development over the \emp/life course}
  \begin{itemize}
\item Convergence, divergence or maintenance.
\item Antonovsky – large differences in adulthood, convergence in old age
\end{itemize}

\end{frame}

\begin{frame}{Main results}

\begin{itemize}
\item A consistent mortality difference according to social class appears surprisingly late in our regions and has no clear connection to the development of income inequality.
\item The association with social class is stronger in the working-age population compared with the elderly, retired population.
\item However, there is a strong gender difference in the development of social inequality.
\item We argue that health-related behavior is an important aspect in historical as well as present-day mortality, and that such behaviours are based on strong gender and class expectations.
\end{itemize}

\end{frame}

\begin{frame}{Economic inequality (figure from Roine and Waldenström 2008)}

\includegraphics[height=8cm]{myfigs/Top_incomes.jpg}

\end{frame}

% \begin{frame}{Outline}
% 
% \begin{center}
% \begin{enumerate}
% 
% \item \emp{Data} management
% \item \emp{Modeling} considerations
% \item \emp{Analysis}
% \item \emp{Results}
% \item \emp{Implications} of the results
% \end{enumerate}
% 
% \end{center}
% 
% \end{frame}

\begin{frame}{Two regions in Västerbotten}

\begin{columns}

\begin{column}{0.5\textwidth}

\includegraphics[height=7cm]{myfigs/sweden.pdf}

\end{column}

\begin{column}{0.5\textwidth}

Remote area in northern Sweden, dominated by agriculture until well into the 20th century.

Skellefteå:
\begin{itemize}

\item Small town with rural hinterland, industrialization of the region during the 20th century.

\end{itemize}

Umeå:

\begin{itemize}

\item Center for administration, education, and military. University from 1965

\end{itemize}

% Rural areas:
% 
% \begin{itemize}
% 
% \item Very similar: Small farms dominating.
% 
% \end{itemize}

\end{column}
\end{columns}
%\includegraphics[height=7cm]{myfigs/sweden.pdf}

\end{frame}

\begin{frame}[fragile]{Data sources}

\begin{itemize}
\item \emp{Swedish church book} records (1801--1950).
  \begin{itemize}
\item digitized by the \emp{Demographic Data Base} at Umeå University,
\begin{itemize}
  % \item \emp{\sc Popum} and
   \item \emp{\sc Poplink}
\end{itemize}
\item  \url{http://www.cedar.umu.se/}
\end{itemize}
\item The \emp{Linnaeus} data base (1976--2013)
  \begin{itemize}
\item Census data 1960, 1965, \ldots, 1990
\item Population registers ({\sc Lisa}, 1986--2013)
%  \begin{itemize}
%    \item \emp{Income}
%    \item \emp{Education}
%  \end{itemize}
 \item Death information from \emp{National Board of Health and Welfare}
   \begin{itemize}
\item Death dates from 1 January 1976 to 31 December 2013.
  \item \emp{Causes of death} ({\sc Ulorsak}, ``main cause of death'')
   \end{itemize}
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}{The sampling frame}
{Two separate datasets}

<<lexis2>>=
plot(c(1801, 1801, 1951, 1951, 1801), c(40, 90, 90, 40, 40),
     type = "l", xlim = c(1800, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 90))
axis(1, at = c(1801, 1851, 1901, 1951, 1976, 1991, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1976, 1976, 2014, 2014, 1976), c(40, 90, 90, 40, 40),
      col = "blue", lwd = 2)
#
pn.exp <- round(sum(skel$exposure) + sum(skume$exposure))
pn.d <-  as.integer(sum(skel$event) + sum(skume$event))
#
text(1875, 70, "PopLink", cex = 2, col = "red")
##text(1899, 60, paste(pn.rec, "records"))
text(1875, 55, paste(pn.exp, "p-years"))
text(1874.5, 50, paste(pn.d, "deaths"))
#
n.exp <- round(sum(foli$exposure))
n.d <-  as.integer(sum(foli$event))

text(1995, 70, "Linnaeus", cex = 1.5, col = "blue")
##text(1984, 60, paste(n.rec, "records"))
text(1995.2, 55, paste(n.exp, "p-years"), cex = 0.8)
text(1995.2, 50, paste(n.d, "deaths"))
@
\end{frame}

\begin{frame}{Variables}

\begin{description}

\item[Sex]
\item[Marital status]
\item[Birth date] By quarter (15 Feb, May, Aug, Nov)
\item[Death date] By day (\emp{Socialstyrelsen}, \emp{Poplink})
\item[*Death cause] from \emp{Socialstyrelsen} and \emp{Poplink}. Adapted to ICD10
\item[Parish] Urban/Rural
\item[Social class] Elite, middle, worker. Household definition 1851--1950. Individual definition 1976--2013
%\item[*Income] 1986--2005, grouped by quartiles.
%\item[*Education] 1990--2013, three categories.
\item[Period]  1851--1875, \ldots,  2008--2013 (13 periods). 
\item[Age group] Separate analyses for working active (40-65) vs retired (65-89) population
%\item[cohort] 1761--1771, \ldots, 1971--1974 (22 birth cohorts).

\end{description}

\end{frame}

\begin{frame}{\sc Social class}

\begin{enumerate}
\item \emp{\tt elite}, HISCLASS 1 and 2. Higher managers and professionals.
\item \emp{\tt middle}, HISCLASS 3, 4, 5, 6, and 8. Lower managers, lower white collar, and farmers.
%%\item \emp{\tt lowWhiteC}, HISCLASS 4, 5 and 6. Lower white collar.
\item \emp{\tt worker}, HISCLASS 7, 9, 10, 11 and 12. Workers of different skills including farm workers.
\end{enumerate}

\end{frame}

\begin{frame}{No proportional hazards (wrt class)!}

\begin{itemize}
\item Fit each social class \emp{separately}, for each time period.
\item Use the \emp{cumulative hazard} at 65 (or 90) for comparison.
\end{itemize}

So, we fit separate PH models for each combination of

\begin{itemize}
  \item \emp{\tt sex}
  \item \emp{\tt period}: 1851--75, 1876--1900, 1901--10, etc.
  \item \emp{\tt class}
  \item \emp{\tt age group}: 40--64, 65--89.
  \item \emp{\tt cause of death}
\end{itemize}
\end{frame}

\begin{frame}{Probabilities of dying, ages 40--64, 1801-2013}

%  \begin{center}
%\includegraphics[height = 7.5cm]{figure/cumsw-1}
%  \end{center}

<<hazcalw>>=
source("R/getPHglm2.R")
##nper <- 4
vb4064 <- vb0Tab[as.numeric(as.character(vb0Tab$age)) < 64.5, ]
vb4064$age <- factor(vb4064$age)
labb <- levels(vb0Tab$period)
nper <- length(labb)
resw <- getPHglm2(vb4064[vb4064$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb4064[vb4064$sex == "male", ],
            nper, labb)
@

<<cumsw>>=
source("R/plotED.glm.R")
scol <- c("blue", "darkgreen", "darkgreen", "red")
oldpar <- par(mfrow = c(2, 1), cex = 0.7, lwd = 1.5)
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, ylab = "P(dead at 65 | alive at 40)",
                main = "Women", ylim = c(0, 0.6))## exclude = 5:6
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, ylab = "P(dead at 65 | alive at 40)",
                main = "Men", ylim = c(0, 0.6))##, exclude = 5:6)
par(oldpar)
@

\end{frame}

\begin{frame}{Relative differences in probabilities of dying, ages 40--64 and 65--89}

  %\begin{center}
\includegraphics[height=7.5cm,width=5.4cm]{figs/art2-cumswstand-1}\includegraphics[height=7.5cm,width=5.4cm]{figs/art2-cumswold-1}
   %\end{center}
\end{frame}

% \begin{frame}{Relative differences, ages 40--64}
% 
%   %\begin{center}
% \includegraphics[height=7.5cm,width=10.8cm]{figs/art2-cumswstand-1}
%    %\end{center}
% \end{frame}
% 
% \begin{frame}{Relative differences, ages 65--89}
% 
%   %\begin{center}
% \includegraphics[height=7.5cm,width=10.8cm]{figs/art2-cumswold-1}
%    %\end{center}
% \end{frame}
%

% \begin{frame}[fragile]{Education 1990--2013}
% 
% <<educ>>=
% 
% educ <- lisaold[!(lisaold$period %in% c("1986-1987", "1988-1989")), ]
% educ$period <- factor(educ$period)
% educ <- educ[!is.na(educ$educ), ]
% levels(educ$educ) <- c("low", "low", "middle", "middle", "middle", "high", "high")
% educ$educ <- relevel(educ$educ, ref = "middle")
% educ$civst <- relevel(educ$civst, ref = "married")
% fitm <- coxph(Surv(enter, exit, event) ~ strata(educ) + civst + urban,
%              data = educ[educ$sex == "male", ])
% fitf <- coxph(Surv(enter, exit, event) ~ strata(educ) + civst + urban,
%              data = educ[educ$sex == "female", ])
% cols <- c("darkgreen", "red", "black")
% @
% 
% <<ploteduc>>=
% oldpar <- par(mfrow = c(1, 2))
% 
% plot(survfit(fitm, start.time = 40), mark.time = FALSE, main = "Men",
%      col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
% axis(1)
% axis(2, at = c(0, 0.5, 1.0), las = 1)
% abline(h = 0.5, lty = 3)
% box()
% legend("bottomleft", legend = levels(educ$educ), col = cols, lty = 1, lwd = 1.5)
% 
% plot(survfit(fitf, start.time = 40), mark.time = FALSE, main = "Women",
%      col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
% axis(1)
% axis(2, at = c(0, 0.5, 1.0), las = 1)
% abline(h = 0.5, lty = 3)
% box()
% legend("bottomleft", legend = levels(educ$educ), col = cols, lty = 1, lwd = 1.5)
% 
% par(oldpar)
% @
% 
% \end{frame}

% \begin{frame}{Income 1986--2005}
% 
% <<femeduc>>=
% income <- lisaold[!is.na(lisaold$income), ]
% income$period <- factor(income$period)
% ##income <- income[!is.na(income$income), ]
% ##levels(income$income) <- c("low", "low", "middle", "middle", "middle", "high", "high")
% income$income <- relevel(income$income, ref = "low")
% income$civst <- relevel(income$civst, ref = "married")
% fitm <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
%              data = income[income$sex == "male", ])
% fitf <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
%              data = income[income$sex == "female", ])
% cols <- c("darkgreen", "red", "blue", "black")
% 
% @
% 
% <<plotincome>>=
% oldpar <- par(mfrow = c(1, 2))
% 
% plot(survfit(fitm, start.time = 40), mark.time = FALSE, main = "Men",
%      col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
% axis(1)
% axis(2, at = c(0, 0.5, 1.0), las = 1)
% abline(h = 0.5, lty = 3)
% box()
% legend("bottomleft", legend = levels(income$income), col = cols, lty = 1, lwd = 1.5)
% 
% plot(survfit(fitf, start.time = 40), mark.time = FALSE, main = "Women",
%      col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
% axis(1)
% axis(2, at = c(0, 0.5, 1.0), las = 1)
% abline(h = 0.5, lty = 3)
% box()
% legend("bottomleft", legend = levels(income$income), col = cols, lty = 1, lwd = 1.5)
% 
% par(oldpar)
% @
% 
% \end{frame}

%%% Stop here!

\begin{frame}{Summarizing the results}

%\scriptsize

\begin{itemize}
\item \emp{Improved} survival characterised all social classes.
\item \emp{Late} appearance of the “modern” pattern of social differences in mortality among men
\item Distinct \emp{gender} differences in the development.
\item \emp{Large} differences between those in working age compared to the retired population, indications of \emp{status convergence}.
% \item \emp{Educational} and \emp{income} differences during the last decades conform to what we found for class differences.
\item In the long run, no clear connection in the development of income inequality and social differences in mortality.
\item This suggests an important role of gender and class related expectations on health behaviour
\end{itemize}

\end{frame}

\begin{frame}{Some implications}

%\scriptsize

\begin{itemize}
\item The \emp{development} of social health inequalities and the role of \emp{health-related behavior} in different phases of the transition (Antonovsky and Omran).
\item Inequality as a \emp{contextual} health risk (Wilkinson/Pickett).
\item \emp{Status syndrome} (Marmot)
\item The role of \emp{living standard} for mortality decline (McKeown).
\item \emp{Gender} and mortality (Watkins, Mckinnon, Janssens).
\item Social conditions as \emp{fundamental causes of death} (Link/Phelan).

\end{itemize}

\end{frame}

\begin{frame}

\begin{center}

\Huge

Thank you for your attention!

\end{center}

\end{frame}

%\begin{frame}{Conclusion}

%\scriptsize

% \begin{itemize}
% \item \emp{Improved} survival characterised all {\sc Hisclass}es.
% \item There are \emp{gender differences} in the social pattern
% \item No difference in the social pattern in the two studied age groups.
% \item The \emp{upper} class(es) gradually take over as the \emp{most healthy} group.
% \item \emp{Educational} and \emp{income} income differences during the last decades conform with what we found for class differences.
% \end{itemize}

%\end{frame}

\begin{frame}{Probabilities of dying, ages 40--64, 1976--2013}

<<cumswlate>>=
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1), cex = 0.7, lwd = 1.5)
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, ylab = "P(dead at 65 | alive at 40)",
                main = "Women", exclude = 1:6, ylim = c(0, 0.3))
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, ylab = "P(dead at 65 | alive at 40)",
                main = "Men", exclude = 1:6, ylim = c(0, 0.3))
par(oldpar)
@

\end{frame}

\begin{frame}{Probabilities of dying, ages 65--89, 1801-2013}

%  \begin{center}
%\includegraphics[height = 7.5cm]{figure/cumsw-1}
%  \end{center}

<<hazcalw2>>=
source("R/getPHglm2.R")
##nper <- 4
vb6589 <- vb0Tab[as.numeric(as.character(vb0Tab$age)) > 64.5, ]
vb6589$age <- factor(vb6589$age)
labb <- levels(vb0Tab$period)
nper <- length(labb)
resw <- getPHglm2(vb6589[vb6589$sex == "female", ],
            nper, labb)
resm <- getPHglm2(vb6589[vb6589$sex == "male", ],
            nper, labb)
@

<<cumsw2>>=
source("R/plotED.glm.R")
scol <- c("blue", "darkgreen", "darkgreen", "red")
oldpar <- par(mfrow = c(2, 1), cex = 0.7, lwd = 1.5)
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, ylab = "P(dead at 90 | alive at 65)",
                main = "Women", ylim = c(0.4, 1.0))## exclude = 5:6
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, ylab = "P(dead at 90 | alive at 65)",
                main = "Men", ylim = c(0.4, 1.0))##, exclude = 5:6)
par(oldpar)
@

\end{frame}

\begin{frame}{Probabilities of dying, ages 65--89, 1976--2013}

<<cumswlate2>>=
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1), cex = 0.7, lwd = 1.5)
woutp <- plotED.glm(resw, col = scol, logScale = FALSE, ylab = "P(dead at 90 | alive at 65)",
                main = "Women", exclude = 1:6, ylim = c(0.4, 1.0))
moutp <- plotED.glm(resm, col = scol, logScale = FALSE, ylab = "P(dead at 90 | alive at 65)",
                main = "Men", exclude = 1:6, ylim = c(0.4, 1.0))
par(oldpar)
@

\end{frame}

\begin{frame}{The sampling frame}

<<lexis2>>=
@

\end{frame}

\begin{frame}{Exposure by social class, women}

\begin{center}
\includegraphics[height = 5cm]{figs/art2-hiscskumw-1}

\small From the top: Elite, middle, workers, missing.
\end{center}
\end{frame}

\begin{frame}{Exposure by social class, men}

\begin{center}
\includegraphics[height = 5cm]{figs/art2-hiscskumm-1}

\small From the top: Elite, middle, workers, missing.
\end{center}
\end{frame}

\begin{frame}{Exposure by marital status}

\begin{center}
\includegraphics[height = 5cm]{figs/art2-civstskum-1.pdf}

\small From the top: Never married, married, previously married.
\end{center}

\end{frame}

\begin{frame}[fragile]{Cohort mortality, men 40--65}

<<cohortapproach>>=
vv <- vb0Tab[vb0Tab$cohort %in% seq(1751, 1951, by = 10) &
               vb0Tab$age %in% seq(40, 60, by = 5), ]
vv$age <- factor(vv$age)
vv$cohort <- factor(vv$cohort)
cohs <- levels(vv$cohort)
vv$age <- factor(vv$age)
n <- length(cohs)
coe <- matrix(0, nrow = 4, ncol = n)
for (i in 1:n){
   fit <- glm(event ~ offset(log(exposure)) + hisclass + age + urban + civst, 
              data = vv[vv$cohort == cohs[i] & vv$sex == "male", ],
              family = poisson)
   x <- coefficients(fit)[1:4]
   x[2:4] <- x[1] + x[2:4]
   coe[, i] <-  exp(x) / sum(exp(x))
}
x <- as.numeric(as.character(cohs))
par(lwd = 2)
plot(x[-n], coe[1, -n], type = "b", col = scol[1], ylim = c(0, 0.7), xlab = "Birth cohort", ylab = "Standardized rel. risks")
lines(x[-n], coe[2, -n], type = "b", col = scol[2], pch = 2, lty = 2)
lines(x[-n], coe[3, -n], type = "b", col = scol[3], pch = 3, lty = 3)
lines(x[-n], coe[4, -n], type = "b", col = scol[4], pch = 4, lty = 4)
legend("topleft", legend = c("elite", "middle", "worker", "NA"), pch = 1:4, col = scol, lty = 1:4)
@

\end{frame}

\begin{frame}{Crude statistics}

<<crude>>=
dat <- with(vb0Tab, aggregate(list(event = event, exposure = exposure), 
                              by = list(period = period, sex = sex), 
                              FUN = sum))
males <- dat[dat$sex == "male", c("period", "event", "exposure")]
females <- dat[dat$sex == "female", c("period", "event", "exposure")]
oldpar <- par(mfrow = c(1, 2))
x <- c(1812.5, 1837.5, 1862.5, 1888.5, 1912.5, 1937.5, 1987.5, 1982.5, 1987.5, 
       1992.5, 1997.5, 2004.5, 2010.5)
plot(x, males$event, type = "b", col = "blue", main = "Number of deaths by period",
     ylim = c(0, 7500), ylab = "No. of deaths", xlab = "Time period", lty = 1, axes = FALSE)
axis(1, at = c(1801, 1826, 1851, 1876, 1901, 1926, 1951, 1976, 2014), las = 2)
axis(2, las = 1)
box()
lines(x, females$event, type = "b", col = "red", lty = 2)
legend("bottomright", legend = c("Males", "Females"), lty = c(1, 2), col = c("blue", "red"))
plot(x, males$event / males$exposure * 1000, type = "b", col = "blue", 
     main = "Crude death rates, ages 40-89", ylim = c(0, 30), xlim = c(1800, 2014),
     xlab = "Time period", ylab = "Deaths per 1000", axes = FALSE)
axis(1, at = c(1801, 1826, 1851, 1876, 1901, 1926, 1951, 1976, 2014), las = 2)
axis(2, las = 1)
box()
lines(x, females$event / females$exposure * 1000, type = "b", col = "red")
legend("bottomright", legend = c("Males", "Females"), lty = c(1, 2), col = c("blue", "red"))
@

\end{frame}
\end{document}