### 2017-10-24 apc.Rmd
first commit, beginning of apc model

---
title: "apc.Rnw"
author: "Göran Broström"
date: "October 24, 2017"
output: html_document
---

## Read data

```{r setup2, include=FALSE}
library(knitr)
opts_chunk$set(engine='R',fig.path='figs/art3-',tidy=FALSE,fig.height=3.5,fig.width=8,strip.white=TRUE,cache=FALSE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
library(knitr)
library(eha)
library(xtable)
library(skum)
```

```{r readalldata}
source("R/get.file.R")
skel <- get.file("skel")
##
skume <- get.file("skume")
##
foli <- get.file("foli")
skel$urban <- FALSE
vb0Tab <- rbind(skel, skume[, names(skel)], foli[, names(skel)])
vb0Tab$civst <- relevel(vb0Tab$civst, ref = "married")
```

## Age-period-cohort data

```{r datai}
library(Epi)
dat <- vb0Tab[vb0Tab$sex == "male", c("age", "period", "cohort", "event", "exposure")]
dat$age <- as.numeric(dat$age)
dat$A <- dat$age + 2.5
dat$Per <- as.numeric(dat$period)
dat$P <- 25 * dat$Per + 1787.5
dat$P[dat$period == "1976-1980"] <- 1977.5
dat$P[dat$period == "1981-1985"] <- 1982.5
dat$P[dat$period == "1986-1990"] <- 1987.5
dat$P[dat$period == "1991-1995"] <- 1992.5
dat$P[dat$period == "1996-2001"] <- 1999
dat$P[dat$period == "2002-2007"] <- 2005
dat$P[dat$period == "2008-2013"] <- 2011
dat$D <-dat$event
dat$Y <- dat$exposure
summary(dat)
```

## APC model fit

```{r}
ex.H <- apc.fit(dat, npar = 7, model = "ns", dr.extr = "Holford",  parm = "ACP", scale = 100)
ex.W <- apc.fit(dat, npar = 7, model = "ns", dr.extr = "weighted", parm = "ACP", scale = 100)
ex.S <- apc.fit(dat, npar = 7, model = "ns", parm = "AC-P", scale = 100)
## Show the estimated drifts
ex.H[["Drift"]]
ex.W[["Drift"]]
ex.S[["Drift"]]
```

Then we plot the results

```{r}
# Plot the effects
fp <- apc.plot( ex.H )
apc.lines( ex.W, frame.par=fp, col="red" )
apc.lines( ex.S, frame.par=fp, col="blue" )
```