### 2018-01-20 ols.Rmd
revision

---
title: "Sampling deaths and OLS"
author: "Göran Broström"
date: "January 19, 2018"
output:
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::pdf_document2:
    citation_package: natbib
    number_sections: yes
    toc: no
    toc_depth: 2
    keep_tex: true

---

# Introduction

We check the effect of assuming normally distributed remaining life after age 60.

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(engine='R',fig.path='figs/ols-',tidy=FALSE,fig.height=4,fig.width=8,strip.white=TRUE,cache=FALSE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
library(knitr)
library(eha)
library(xtable)
library(skum)
options(show.signif.stars = FALSE)
```

# Data

Data come from Skellefteå and Umeå. The selection is all persons born between January 1, 1841 and December 31, 1850 and present in 
the data at age 60.
They are followed from age 60 to death or right censoring (at most until December 31, 1950).

```{r readalldata}
##source("R/get.file.R")
##skume <- get.file("skume")
## Not needed(?):
##vb0Tab <- skume
##vb0Tab$civst <- relevel(vb0Tab$civst, ref = "married")
##vb0Tab <- aggregate(vb0Tab[, c("event", "exposure")],
##                    vb0Tab[, c("sex", "age", "period", "cohort", "civst", "hisclass", "urban")],
##                    FUN = sum)
## End: Not needed(?)
ols <- obs[, c("id", "sex", "birthdate", "region", "civst", "sluttyp", "enter", "exit")]
ols$event <- ols$sluttyp == 2
ols$sluttyp <- NULL
ols <- ols[ols$birthdate >= 1841 & ols$birthdate < 1851, ]
ols <- age.window(ols, c(60, 100))
ids <- ols$id[ols$enter < 60.01]
ols <- ols[ols$id %in% ids, ]
indx <- tapply(ols$id, ols$id)
out <- tapply(ols$exit, ols$id, max)
ols$out <- out[indx]
ols <- ols[order(ols$id, -ols$exit, -ols$event),]
##sum(ols$event)
ols <- ols[!duplicated(ols$id), ]
##sum(ols$event)
ols$exit <- ols$out
ols$out <- NULL
summary(ols)
```

We want to observe time since age 60 (exact!), so we subtract 60 from the ages `enter` and `exit`.

```{r}
ols$enter <- ols$enter - 60
ols$exit <- ols$exit - 60
zero <- ols$enter >= ols$exit
ols$enter[zero] <- ols$exit[zero] - 0.001 # Fix zero-length intervals (no big deal)
```

In the following, we are interested in the *sex difference* in expected remaining life after 60 (*e60*) (just as an illustration). The categorical variable `region` is included as an example of a confounder.

# Standard survival analyses

## Ordinary *Cox regression*

We first try to get a picture of what really is happening in mortality relative to `sex` and `region` (the figure is truncated at duration 30, i.e., age 90).

```{r standards}
fit <- coxph(Surv(enter, exit, event) ~ strata(region, sex), data = ols)
plot(survfit(fit), fun = "cumhaz", col = 1:4, lty = 1:4, xlab = "Years after 60.", ylab = "Cumulative hazards", xlim = c(0, 30))
legend("topleft", legend = levels(strata(ols$region, ols$sex)), col = 1:4, lty = 1:4)
abline(h = 0)
```

It seems as if males have the same mortality experience in Umeå and Skellefteå, while women from Umeå are better off than those from Skellefteå. Of course, women are much better off than men, reardless of region. So, there is a tendency of *interaction* between the effects on mortality of `sex` and `region`.

We cannot easily calculate *e60* from this figure (or table above), but the estimated median remaining life at 60 (*m60*) is easy to illustrate graphically. For that porpose we need to plot the estimated *survival functions*.
```{r }
plot(survfit(fit), col = 1:4, lty = 1:4, xlab = "Years after 60.", 
     ylab = "Remaining fraction", xlim = c(0, 30), axes = FALSE)
axis(1, at = c(0, 9, 13, 20, 30))
axis(2, at = c(0, 0.5, 1), las = 1)
box()
legend("bottomleft", legend = levels(strata(ols$region, ols$sex)), col = 1:4, lty = 1:4)
abline(h = 0)
abline(h = 0.5, lty = 2)
```

We can also make an exact calculation, which results in

```{r exactki}
x <- survfit(fit)
x
```

## AFT regression

Then an *accelerated failure time (AFT)* model ("linear regression on the log scale, with censoring").

```{r afts}
fit <- aftreg(Surv(enter, exit, event) ~ sex + region, data = ols, dist = "lognormal", param = "lifeExp")
meanlog <- fit$coefficients["log(scale)"]
sdlog <- 1 / exp(fit$coefficients["log(shape)"])
fit$baselineMean <- exp(meanlog + sdlog^2 / 2)
summary(fit)
```

The *e60* for men in Skellefteå is 9.69, and for Skellefteå women it is $1.329 \times 9.69 = 12.88$, a difference of 3.19 years.

The lognormal baseline plot looks like this:

```{r plotlns}
plot(fit, fn = "haz", xlab = "Years after 60.")
```

Not very convincing: The mortality is *decreasing* after age 70 (forced by the lognormal assumption)!

We try the *Gompertz* distribution instead:

```{r aftgomps}
fit.g <- aftreg(Surv(enter, exit, event) ~ sex + region, data = ols, dist = "gompertz", param = "lifeExp")

summary(fit.g)
```

The *e60* for men in Skellefteå is now 11.2, and for Skellefteå women it is $1.075 \times 11.2 = 12.04$, a difference of only 0.84 years.

The Gompertz baseline plot looks like this:

```{r plotgomps}
plot(fit.g, fn = "haz", xlab = "Years after 60.")
```

A lot more reasonable.

# Sampling deaths

We repeat the analyses with the data subset consisting of all those who were observerd to die in the parishes.

```{r smallols}
ids <- ols$id[ols$event == 1]
sols <- ols[ols$id %in% ids, ]
```

## Ordinary *Cox regression*

```{r standard}
fit <- coxph(Surv(enter, exit, event) ~ sex + region, data = sols)
survfit(fit)
```

## AFT regression

Then an *accelerated failure time (AFT)* model ("linear regression on the log scale, with censoring").

```{r aft}
fit <- aftreg(Surv(enter, exit, event) ~ sex + region, data = sols, dist = "lognormal", param = "lifeExp")
meanlog <- fit$coefficients["log(scale)"]
sdlog <- 1 / exp(fit$coefficients["log(shape)"])
fit$baselineMean <- exp(meanlog + sdlog^2 / 2)
summary(fit)
```

The lognormal baseline plot looks like this:

```{r plotln}
plot(fit, fn = "haz", xlab = "Years after 60.")
```

Not very convincing: The mortality is *decreasing* after age 70 (forced by the lognormal assumption)!

We try the *Gompertz* distribution instead:

```{r aftgomp}
fit.g <- aftreg(Surv(enter, exit, event) ~ sex + region, data = sols, dist = "gompertz", param = "lifeExp")

summary(fit.g)
```

The Gompertz baseline plot looks like this:

```{r plotgomp}
plot(fit.g, fn = "haz", xlab = "Years after 60.")
```

# OLS regression

Finally, let us perform ordinary OLS regression.

```{r ordOLS}
ools <- sols[sols$event == 1, ]
summary(fit <- lm(exit ~ sex + region, data = ools)) 
```