### 2018-01-21 ols.Rmd
add three references

---
title: "On mortality of the dead and OLS"
author: "Göran Broström"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::pdf_document2:
    citation_package: natbib
    number_sections: yes
    toc: no
    toc_depth: 2
    keep_tex: true
bibliography: mybib.bib
---

# Introduction

We ask if studying mortality by sampling the dead is a good idea and whether it is reasonable to assume normally distributed remaining life after age 60. We answer *No* and *Maybe*.

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(engine='R',fig.path='figs/ols-',tidy=FALSE,fig.height=4,fig.width=8,strip.white=TRUE,cache=FALSE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
library(knitr)
library(eha)
library(xtable)
library(skum)
options(show.signif.stars = FALSE)
```

# Data

Data come from Skellefteå and Umeå. The selection is all persons born between January 1, 1841 and December 31, 1850 and present in 
the data at age 60.
They are followed from age 60 to death or right censoring (at most until December 31, 1950). The first six rows of the data frame:

```{r readalldata}
##source("R/get.file.R")
##skume <- get.file("skume")
## Not needed(?):
##vb0Tab <- skume
##vb0Tab$civst <- relevel(vb0Tab$civst, ref = "married")
##vb0Tab <- aggregate(vb0Tab[, c("event", "exposure")],
##                    vb0Tab[, c("sex", "age", "period", "cohort", "civst", "hisclass", "urban")],
##                    FUN = sum)
## End: Not needed(?)
ols <- obs[, c("id", "sex", "birthdate", "region", "civst", "sluttyp", "enter", "exit")]
ols$event <- ols$sluttyp == 2
ols$sluttyp <- NULL
ols <- ols[ols$birthdate >= 1841 & ols$birthdate < 1851, ]
ols <- age.window(ols, c(60, 100))
ids <- ols$id[ols$enter < 60.01]
ols <- ols[ols$id %in% ids, ]
indx <- tapply(ols$id, ols$id)
out <- tapply(ols$exit, ols$id, max)
ols$out <- out[indx]
ols <- ols[order(ols$id, -ols$exit, -ols$event),]
##sum(ols$event)
ols <- ols[!duplicated(ols$id), ]
##sum(ols$event)
ols$exit <- ols$out
ols$out <- NULL
head(ols)
```

We want to observe time since age 60 (exact!), so we subtract 60 from the ages `enter` and `exit`.

```{r}
ols$enter <- ols$enter - 60
ols$exit <- ols$exit - 60
zero <- ols$enter >= ols$exit
ols$enter[zero] <- ols$exit[zero] - 0.001 # Fix zero-length intervals (no big deal)
##ols$event[ols$exit > 30] <- 1
```

In the following, we are interested in the *sex difference* in expected remaining life after 60 (*e60*) (just as an illustration). The categorical variable `region` is included as an example of a confounder.

# OLS regression

OLS regression in survival analysis is a special case of the AFT regression model, see the next section.

## Full data

*OLS* means assuming *normally distributed* remaining life after 60, and we can use standard linear regression, given that we are willing to assume an *accelerated failure time (AFT)* model for our data. In the case of right censoring, we need to fall back on maximum likelihood for the estimation. The function *survreg* in the **R** [@cran] package **survival** facilitates this.

```{r survrcens}
fit <- survreg(Surv(exit, event) ~ sex + region, data = ols, dist = "gaussian")
x <- summary(fit)
x
```

So the *e60* for men from Skellefteå is `r round(x$coef[1], 2)` years and for women it is `r round(x$coef[2], 2)` more years. The difference is statistically significant.

## Sampling the dead

We repeat the analyses with the data subset consisting of all those who were observerd to die in the parishes. This implies that 13 percent censored individuals are thrown away.

```{r smallols, echo = FALSE}
ids <- ols$id[ols$event == 1]
sols <- ols[ols$id %in% ids, ]
```

```{r survrnocens}
fit2 <- survreg(Surv(exit, event) ~ sex + region, data = sols, dist = "gaussian")
x2 <- summary(fit2)
x2
```

So the *e60* for men from Skellefteå is now only `r round(x2$coef[1], 2)` years and for women it is `r round(x2$coef[2], 2)` more years. The difference is statistically significant, 
and larger than in the case with full data.

## Conclusion

Sampling of deaths leads to overestimation of mortality (underestimation of *e60*): This is a trivial (well-known) consequence. More seriously, it distorts the balance between the effects of categories (`sex`, `region`) on mortality in unpredictable ways.

# AFT regression

The OLS model is an example of what is called  an *accelerated failure time (AFT)* model in survival analysis, a competitor to the *proportional hazards (PH)* model. 
A common assumption about old age mortality is that a *Gompertz* distribution fits well.

So we go back to the full data situation and fit a Gompertz AFT model. The *survreg* function does not include the Gompertz distribution, but the *aftreg* function in the package **eha** [@eha] does.

```{r aft, echo = FALSE}
fit.g <- aftreg(Surv(exit, event) ~ sex + region, data = ols, dist = "gompertz", param = "lifeExp")
summary(fit.g)
em <- fit.g$baselineMean
ef <- fit.g$baselineMean * exp(fit.g$coef[1])
```
So the *e60* for Skellefteå men is `r round(em, 2)` years, and for the women we get 
`r round(ef, 2)` years, a difference of `r round(ef - em, 2)` years, that is, *men* are better off. This difference is *not* statistically significant.

## Is the assumption of normally distributed residual life reasonable?

Yes, as long as condition on survival past the mode of the survival distribution, according to @kann96. So what is the mode in the life distribution in our data?

```{r mode, echo = FALSE, cache = TRUE}
oj <- obs
oj$event <- oj$sluttyp == 2
oj <- age.window(oj, c(0, 100))
as <- oj$enter >= oj$exit
oj$enter[as] <- oj$exit[as] - 0.0001
ff <- phreg(Surv(enter, exit, event) ~ 1, data = oj, dist = "pch", cuts = 1:99)
par(las = 1)
plot(ff, fn = "den", xlab = "Age", main = "", col = "blue")
```

Well, it is closer to 80 than to 60, so the normal distribution assumption may be in doubt. We can compare the baseline (men in Skellefteå) distributions from 
the OLS fit and nonparametric estimation.

```{r plotln}
fit.cox <- coxreg(Surv(exit, event) ~ 1, data = ols[ols$sex == "male" & ols$region == "ske", ])
par(las = 1)
plot(fit.cox, fn = "cum", xlab = "Years after 60.", main = "Baseline Cumulative hazards  functions", ylim = c(0, 5))
x <- seq(0, 37, length = 1000)
y <- -log(pnorm(x, mean = 16.474, sd = 8.96, lower.tail = FALSE))
lines(x, y, col = "red")
legend("topleft", legend = c("Nonparametric", "OLS regression"), lty = c(1, 1), col = c("black", "red"))
abline(h = 0)
```

Looks quite ok, even though the OLS curve heavily overestimates the ages above 90. No simple parametric model will catch the heavy slow-down of the mortality increase in the high ages.

# PH regression

Maybe the PH model fits better than the AFT? Compare max log likelihoods.

```{r phbetter}
fit.ph <- phreg(Surv(exit, event) ~ sex + region, data = ols, dist = "gompertz")
fit.ph
```

No, the difference is very small: `r round(fit.ph$loglik[2] - fit.g$loglik[2], 3)` in favor of the PH model. Not much to argue about.

# A look at data

These are the mortality patterns for the four groups we are contemplating. Worth noting is that male mortality above age 90 is so low (compared to women).

```{r standards}
fit <- coxph(Surv(enter, exit, event) ~ strata(region, sex), data = ols)
par(las = 1)
plot(survfit(fit), fun = "cumhaz", col = 1:4, lty = 1:4, xlab = "Years after 60.", ylab = "Cum. hazards", axes = FALSE)
axis(1, at = c(0, 9, 13, 20, 30, 40))
axis(2, at = c(0, log(2), 3, 5), labels = c("0", "log(2)", "3", "5"))
box()
legend("topleft", legend = levels(strata(ols$region, ols$sex)), col = 1:4, lty = 1:4)
abline(h = 0)
abline(h = log(2), lty = 2)
```

It seems as if males have the same mortality experience in Umeå and Skellefteå, while women from Umeå are better off than those from Skellefteå. Of course, women are much better off than men (before age 90!), reardless of region. So, there is a tendency of *interaction* between the effects on mortality of `sex` and `region`.

We cannot easily calculate *e60* from this figure (or table above), but the estimated *median remaining life at 60 (m60)* is easy to illustrate: It is the years-after-60 at which the horizontal line at $\log(2)$ intersects the curves: 9 years for men and 13 for women. Maybe easier to see in the plot of the estimated *survival functions* below.
```{r }
plot(survfit(fit), col = 1:4, lty = 1:4, xlab = "Years after 60.", 
     ylab = "Remaining fraction", axes = FALSE)
axis(1, at = c(0, 9, 13, 20, 30))
axis(2, at = c(0, 0.5, 1), las = 1)
box()
legend("topright", legend = levels(strata(ols$region, ols$sex)), col = 1:4, lty = 1:4)
abline(h = 0)
abline(h = 0.5, lty = 2)
```

We can also make an exact calculation, which results in

```{r exactki}
x <- survfit(fit)
x
```

# Conclusion

We can live with the assumption of normally distributed residual life after age 60, but we must be aware of the fact that applying OLS implies rather strong distributional assumptions. It is a misconception to believe that least squares methods are "distribution free". Sampling of the dead, on the other hand, is problematic and should be avoided without measures that are specifically designed to take care of that situation (what ever they are).

# References