### 2018-01-30 slides3.Rnw
first commit

\documentclass[t]{beamer}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{natbib}

%\usepackage[figuresonly]{endfloat}

%\bibliographystyle{apalike}

\usepackage{color}

\newcommand{\emp}[1]{\textcolor{blue}{#1}}

\usetheme{Singapore}% best choice.
  \setbeamercovered{transparent}%

\newcommand{\R}{{\bf R}}
\newcommand{\code}[1]{{\tt #1}}

\title{Social inequality in mortality among adults and elderly in
  northern  Sweden 1851--2013\footnote{Seminar April 27, CEDAR, Umeå}
}

\date{April 27, 2017}

\author{Göran Broström and Sören Edvinsson}

<<setup, include = FALSE>>=
library(knitr)
opts_chunk$set(engine='R',dev='pdf',fig.path='figs/slide-',tidy=FALSE,fig.height=5.5,fig.width=8,strip.white=TRUE,cache=TRUE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
@

<<getgoing>>=
library(knitr)
library(eha)
library(xtable)
library(skum)
## Load remote data:
con <- url("http://capa.ddb.umu.se/data/vb0.rda")
##load("data/vb0.rda")
load(con)
close(con)
vb0$hisclass[is.na(vb0$hisclass)] <- 4
vb0$hisclass <- factor(vb0$hisclass, labels = c("elite", "middle", "worker", "NA"))
##
con <- url("http://capa.ddb.umu.se/data/vb.rda")
##load("data/vb.rda")
load(con)
close(con)
##
con <- url("http://capa.ddb.umu.se/data/lisaold.rda")
##load("data/lisaold.rda")
load(con)
close(con)
##
vb0Tab <- aggregate(vb0[, c("event", "exposure")],
                   vb0[, c("sex", "age", "period", "cohort",
                          "urban", "civst", "hisclass")],
                   FUN = sum)
vb0Tab$age <- factor(vb0Tab$age)

@

\begin{document}

\maketitle

\begin{frame}{Questions}

\begin{enumerate}
\item Has \emp{inequality in mortality} between \emp{social classes}
  increased in the adult and elderly
  population?

%%\item Is the mortality transition initiated by a certain social class?

\item Are there any \emp{gender differences} in the effect of social position?

\item Is social position \emp{equally important} among the \emp{retired} population as
  for those in \emp{working age}?
\item Is there a difference between classes in mortality from
  \emp{cardiovascular} diseases and \emp{cancers} respectively?
\item Are there large differences in survival depending on level of
  \emp{education} and \emp{income} categories?

\end{enumerate}

\end{frame}

\begin{frame}{Outline}

\begin{center}
\begin{enumerate}

\item \emp{Data} management
\item \emp{Modeling} considerations
\item \emp{Analysis}
\item \emp{Results}
\item \emp{Alternative} Results
\end{enumerate}

\end{center}

\end{frame}

\begin{frame}{Here!}

\begin{columns}

\begin{column}{0.5\textwidth}

%\includegraphics[height=7cm]{myfigs/skum17.pdf}
\includegraphics[height=7cm]{myfigs/skum.pdf}
\end{column}

\begin{column}{0.5\textwidth}

Skellefteå:

\begin{itemize}

\item Small town, heavy industry in the nineteenth century.

\end{itemize}

Umeå:

\begin{itemize}

\item Not so small town, older. Center for administration, education, and military.

\end{itemize}

Rural areas:

\begin{itemize}

\item Very similar: Small farms dominating.

\end{itemize}

\end{column}
\end{columns}
%\includegraphics[height=7cm]{myfigs/sweden.pdf}

\end{frame}

\begin{frame}[fragile]{Data sources}

\begin{itemize}
\item \emp{Swedish church book} records (1851--1950).
  \begin{itemize}
\item digitized by the \emp{Demographic Data Base} at Umeå University,
\begin{itemize}
  % \item \emp{\sc Popum} and
   \item \emp{\sc Poplink}
\end{itemize}
\item  \url{http://www.cedar.umu.se/}
\end{itemize}
\item The \emp{Linnaeus} data base (1960--2013)
  \begin{itemize}
\item Census data 1960, 1965, \ldots, 1990
\item Population registers ({\sc Lisa}, 1986--2013)
  \begin{itemize}
    \item \emp{Income}
    \item \emp{Education}
  \end{itemize}
 \item Death information from \emp{National Board of Health and Welfare}
   \begin{itemize}
\item Death dates from 1 January 1961 to 31 December 2013.
  \item \emp{Causes of death} ({\sc Ulorsak}, ``main cause of death'')
   \end{itemize}
\end{itemize}sampling fra
\end{itemize}

\end{frame}

\begin{frame}[fragile]{The sampling frame}

<<lexis2,echo = FALSE>>=
plot(c(1851, 1851, 1951, 1951, 1851), c(40, 90, 90, 40, 40),
     type = "l", xlim = c(1851, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 90))
axis(1, at = c(1851, 1901, 1951, 1961, 1981, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1961, 1961, 2014, 2014, 1961), c(40, 90, 90, 40, 40),
      col = "blue", lwd = 2)
#
n.rec <- NROW(vb0[vb0$dataset == "poplink", ] )
n.pl <- with(vb0[vb0$dataset == "poplink", ], length(unique(id)))
n.d <-  with(vb0[vb0$dataset == "poplink", ], sum(event))
text(1900, 70, "PopLink", cex = 2, col = "red")
text(1899, 60, paste(n.rec, "records"))
text(1900, 55, paste(n.pl, "persons"))
text(1899.5, 50, paste(n.d, "deaths"))
#
n.rec <- NROW(vb0[vb0$dataset != "poplink", ] )
n.pl <- with(vb0[vb0$dataset != "poplink", ], length(unique(id)))
n.d <-  with(vb0[vb0$dataset != "poplink", ], sum(event))

text(1985, 70, "Linnaeus", cex = 2, col = "blue")
text(1984, 60, paste(n.rec, "records"))
text(1985.2, 55, paste(n.pl, "persons"))
text(1985.2, 50, paste(n.d, "deaths"))
@

\end{frame}

\begin{frame}[t]{The FOB and Lisa data}

\vspace{-1.5cm}
%\begin{columns}
%\begin{column}{0.6\textwidth}
<<picklexus,fig.width=4.7,fig.height = 4.8, echo=FALSE>>=

##load("data/vb.rda")  ####################################### NOTE!!!!!!!!!!!!!!!! ############################
##load("data/vb0.rda")
tot.m <- length(unique(vb$id[vb$dataset != "poplink" & vb$sex == "male"]))
tot.f <- length(unique(vb$id[vb$dataset != "poplink" & vb$sex == "female"]))
vbd <- vb0[vb0$event == 1 & vb0$dataset != "poplink", c("birthdate", "exit", "sex")]
d.m <- sum(vbd$sex == "male")
d.f <- sum(vbd$sex == "female")
x <- vbd$birthdate + vbd$exit
y <- vbd$exit
##par(mfrow = c(1, 2))
plot(x[vbd$sex == "male"], y[vbd$sex == "male"], type = "n",
     pch = ".", col = "blue", xlab = "Year", ylab = "Age", las = 2,
     axes = FALSE, ylim = c(28, 90))
axis(1, at = c(seq(1960, 1990, by = 10), 2002, 2014), las = 1)
axis(2, at = c(40, 65, 90), las = 2)
axis(4, at = c(40, 65, 90), las = 2)
for (i in seq(1960, 1990, by = 5) + (365 - 61) / 365){
    ##abline(v = i, lwd = 2, col = "red")
    lines(c(i, i), c(40, 90), lwd = 2, col = "red")
}
for (i in seq(1960, 1985, by = 5) + (365 - 61) / 365){
    lines(c(i, i + 5), c(35, 40), col = "red", lty = 2, lwd = 2)
    lines(c(i, i), c(35, 40), col = "red", lty = 2, lwd = 2)
}
lines(c(1990, 1990) + (365 - 61)/365, c(28, 40), lwd = 2, lty = 2, col = "red")
lines(c(1990 + (365 - 61) / 365, 2002), c(28, 40), lwd = 2, lty = 2, col = "red")
for (i in 2002:2014){
    ##abline(v = i, lwd = 2, col = "blue")
    lines(c(i, i), c(40, 90), col = "blue", lwd = 2)
}
for (i in 2002:2013){
    lines(c(i, i), c(39, 40), col = "blue", lty = 2, lwd = 1.5)
    lines(c(i, i+1), c(39, 40), col = "blue", lty = 2, lwd = 1.5)
}
lines(c(2002, 2014), c(40, 40), col = "blue", lwd = 1.5)
lines(c(2002, 2014), c(90, 90), col = "blue", lwd = 1.5)
lines(c(1960, 2002), c(90, 90), lwd = 2, col = "red")
lines(c(1960, 2002), c(40, 40), lwd = 2, col = "red")
##abline(h = c(35, 40, 65, 90), lty = 2)
points(x[vbd$sex == "female"], y[vbd$sex == "female"], type = "n",
     pch = ".", col = "red")
box()
@

%\end{column}

%\begin{column}{0.4\textwidth}
%\scriptsize

%\vspace{3\baselineskip}

%\begin{tabular}{l|rr}
%      & Number & Deaths \\ \hline
%\textcolor{red}{Women} & \Sexpr{tot.f} & \textcolor{red}{\Sexpr{d.f}} \\
%\emp{Men} & \Sexpr{tot.m} & \emp{\Sexpr{d.m}} \\ \hline
%Total  & \Sexpr{tot.f + tot.m} & \Sexpr{d.f + d.m}
%\end{tabular}

%\vspace{\baselineskip}

%\emp{\underline{Note}}: Persons with \emp{missing {\sc Hisclass}} (occupation) are %\emp{excluded}.
%\end{column}
%\end{columns}
\end{frame}

\begin{frame}[fragile]{The Linnaeus data with deaths}

\begin{columns}
\begin{column}{0.6\textwidth}

<<picklexus2,fig.width=6,fig.height = 5.9,echo=FALSE>>=

##load("data/vb.rda")
tot.m <- length(unique(vb0$id[vb0$dataset != "poplink" & vb0$sex == "male"]))
tot.f <- length(unique(vb0$id[vb0$dataset != "poplink" & vb0$sex == "female"]))

vbd <- vb0[vb0$event == 1 & vb0$dataset != "poplink", c("birthdate", "exit", "sex")]
d.m <- sum(vbd$sex == "male")
d.f <- sum(vbd$sex == "female")

x <- vbd$birthdate + vbd$exit
y <- vbd$exit
##par(mfrow = c(1, 2))
plot(x[vbd$sex == "female"], y[vbd$sex == "female"], type = "p",
     pch = ".", col = "blue", xlab = "Year", ylab = "Age", las = 2,
     axes = FALSE)
axis(1, at = c(seq(1960, 1990, by = 5), 2002, 2007, 2014), las = 2)
axis(2, at = c(40, 65, 90))
axis(4, at = c(40, 65, 90))
for (i in seq(1960, 1990, by = 5) + (365 - 61) / 365){
    abline(v = i, lwd = 2)
}
for (i in 2002:2013){
    abline(v = i, lwd = 2, col = "black")
}
abline(h = c(40, 65, 90), lty = 2)
points(x[vbd$sex == "male"], y[vbd$sex == "male"], type = "p",
     pch = ".", col = "blue")
box()
@

\end{column}

\begin{column}{0.4\textwidth}
\scriptsize

\vspace{3\baselineskip}

\begin{tabular}{l|rr}
      & Number & Deaths \\ \hline
\textcolor{red}{Women} & \Sexpr{tot.f} & \textcolor{red}{\Sexpr{d.f}} \\
\emp{Men} & \Sexpr{tot.m} & \emp{\Sexpr{d.m}} \\ \hline
Total  & \Sexpr{tot.f + tot.m} & \Sexpr{d.f + d.m}
\end{tabular}

\vspace{\baselineskip}

\emp{\underline{Note}}: Persons with \emp{missing {\sc Hisclass}} (occupation) are \emp{included}.
\end{column}
\end{columns}
\end{frame}

\begin{frame}{The Linnaeus data with deaths}

\begin{columns}
\begin{column}{0.6\textwidth}
<<picklexus3,fig.width=6,fig.height = 5.9,echo=FALSE>>=
##load("data/vb.rda")
##tot.m <- length(unique(vb$id[vb$dataset != "poplink" & vb$sex == "male"]))
##tot.f <- length(unique(vb$id[vb$dataset != "poplink" & vb$sex == "female"]))
tot.m <- length(unique(vb$id[vb$from != "popum" & vb$sex == "male"]))
tot.f <- length(unique(vb$id[vb$from != "popum" & vb$sex == "female"]))

vbd <- vb[vb$event == 1 & vb$from != "popum", c("birthdate", "exit", "sex")]
d.m <- sum(vbd$sex == "male")
d.f <- sum(vbd$sex == "female")

x <- vbd$birthdate + vbd$exit
y <- vbd$exit
##par(mfrow = c(1, 2))
plot(x[vbd$sex == "female"], y[vbd$sex == "female"], type = "p",
     pch = ".", col = "blue", xlab = "Year", ylab = "Age", las = 2,
     axes = FALSE)
axis(1, at = c(seq(1960, 1990, by = 5), 2002, 2007, 2014), las = 2)
axis(2, at = c(40, 65, 90))
axis(4, at = c(40, 65, 90))
for (i in seq(1960, 1990, by = 5) + (365 - 61) / 365){
    abline(v = i, lwd = 2)
}
for (i in 2002:2013){
    abline(v = i, lwd = 2, col = "black")
}
abline(h = c(40, 65, 90), lty = 2)
points(x[vbd$sex == "male"], y[vbd$sex == "male"], type = "p",
     pch = ".", col = "blue")
box()
@

\end{column}

\begin{column}{0.4\textwidth}
\scriptsize

\vspace{3\baselineskip}

\begin{tabular}{l|rr}
      & Number & Deaths \\ \hline
\textcolor{red}{Women} & \Sexpr{tot.f} & \textcolor{red}{\Sexpr{d.f}} \\
\emp{Men} & \Sexpr{tot.m} & \emp{\Sexpr{d.m}} \\ \hline
Total  & \Sexpr{tot.f + tot.m} & \Sexpr{d.f + d.m}
\end{tabular}

\vspace{\baselineskip}

\emp{\underline{Note}}: Persons with \emp{missing {\sc Hisclass}} (occupation) are \emp{excluded}.
\end{column}
\end{columns}
\end{frame}

\begin{frame}{The sampling frame, II}

<<lexis4>>=
##load("data/vb0.rda")
vb0 <- vb0[vb0$dataset == "poplink" |
               (vb0$dataset != "poplink" & vb0$birthdate > 1895), ]
plot(c(1851, 1851, 1951, 1951, 1851), c(40, 90, 90, 40, 40),
     type = "l", xlim = c(1851, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 90))
axis(1, at = c(1851, 1901, 1951, 1961, 1981, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1961, 1961, 1986, 2014, 2014, 1961), c(40, 65, 90, 90, 40, 40),
      col = "blue", lwd = 2)
#
n.rec <- NROW(vb0[vb0$dataset == "poplink", ] )
n.pl <- with(vb0[vb0$dataset == "poplink", ], length(unique(id)))
n.d <-  with(vb0[vb0$dataset == "poplink", ], sum(event))
text(1900, 70, "PopLink", cex = 2, col = "red")
text(1899, 60, paste(n.rec, "records"))
text(1900, 55, paste(n.pl, "persons"))
text(1899.5, 50, paste(n.d, "deaths"))
#
n.rec <- NROW(vb0[vb0$dataset != "poplink", ] )
n.pl <- with(vb0[vb0$dataset != "poplink", ], length(unique(id)))
n.d <-  with(vb0[vb0$dataset != "poplink", ], sum(event))

text(1989, 70, "Linnaeus", cex = 2, col = "blue")
text(1988, 60, paste(n.rec, "records"))
text(1989.2, 55, paste(n.pl, "persons"))
text(1989.2, 50, paste(n.d, "deaths"))
@
\end{frame}

\begin{frame}{Variables}

\begin{description}

\item[sex]
\item[birth date] By quarter (15 Feb, May, Aug, Nov)
\item[death date] By day (\emp{Socialstyrelsen})
\item[cause of death] from \emp{Socialstyrelsen}
\item[parish] Urban/Rural
\item[\sc Hisclass] elite, middle, worker.
\item[income] 1990--2005, grouped by quartiles.
\item[education] 1990--2005, seven categories.
\item[period]  1851--1875, \ldots,  2008--2013 (10 periods).
\item[cohort] 1761--1771, \ldots, 1971--1974 (22 birth cohorts).

\end{description}

\end{frame}

\begin{frame}{\sc Hisclass}

\begin{enumerate}
\item \emp{\tt elite}, HISCLASS 1 and 2. Higher managers and professionals.
\item \emp{\tt middle}, HISCLASS 3, 4, 5, 6, and 8. Lower managers, lower white collar, and farmers.
%%\item \emp{\tt lowWhiteC}, HISCLASS 4, 5 and 6. Lower white collar.
\item \emp{\tt worker}, HISCLASS 7, 9, 10, 11 and 12. Workers of different skills including farm workers.
\end{enumerate}

\end{frame}

\begin{frame}[fragile]{Preliminary Cox regression}

{\scriptsize

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{library}\hlstd{(eha)}
\hlkwd{system.time}\hlstd{(fit} \hlkwb{<-} \hlkwd{coxph}\hlstd{(}\hlkwd{Surv}\hlstd{(enter, exit, event)} \hlopt{~}
                              \hlstd{hisclass} \hlopt{*} \hlstd{period} \hlopt{+} \hlstd{civst} \hlopt{+} \hlstd{urban} \hlopt{+} \hlstd{sex,}
                         \hlkwc{data} \hlstd{= vb0))}
\end{alltt}
\begin{verbatim}
   user  system elapsed
151.360   0.444 151.806
\end{verbatim}
\begin{alltt}
\hlkwd{system.time}\hlstd{(dr} \hlkwb{<-} \hlkwd{drop1}\hlstd{(fit,} \hlkwc{test} \hlstd{=} \hlstr{"Chisq"}\hlstd{))}
\end{alltt}
\begin{verbatim}
   user  system elapsed
455.236   0.944 456.188
\end{verbatim}
\end{kframe}
\end{knitrout}
}

\emp{It takes time (more than 10 minutes)!}

\end{frame}

\begin{frame}{ANOVA}

% latex table generated in R 3.3.3 by xtable 1.8-2 package
% Mon Apr 17 14:43:45 2017
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
 & Df & AIC & LRT & Pr($>$Chi) \\
  \hline
$<$none$>$ &  & 2227937.83 &  &  \\
  civst & 2 & 2228508.16 & 574.34 & 0.0000 \\
  urban & 1 & 2227962.96 & 27.13 & 0.0000 \\
  sex & 1 & 2231526.62 & 3590.79 & 0.0000 \\
  hisclass:period & 27 & 2228575.58 & 691.75 & 0.0000 \\
   \hline
\end{tabular}
\end{table}

And with so much data, everything tends to be \emp{statistically} significant.

\end{frame}

\begin{frame}[fragile]{Test of Proportional Hazards}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
    user   system  elapsed
1063.928    0.916 1064.028
\end{verbatim}
\end{kframe}
\end{knitrout}

Another \emp{19 minutes}!
\end{frame}

\begin{frame}[fragile]{Proportional hazards, results}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{#system.time(cz <- cox.zph(fit))}
\hlstd{cz}\hlopt{$}\hlstd{table[}\hlkwd{NROW}\hlstd{(cz}\hlopt{$}\hlstd{table), ]}
\end{alltt}
\begin{verbatim}
    rho   chisq         p
     NA 2064.07    0.0000
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{itemize}
\item As usual, \emp{highly significant} result (PH assumption violated).
\item Should we worry?
\begin{itemize}
\item Yes
\item No
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}[fragile]{Sufficient statistics}

Aggregate:

{\scriptsize
<<aggregate,eval = TRUE, echo = TRUE>>=
nr <- NROW(vb0Tab)
@
}

\begin{itemize}
\item Gives a table assuming \emp{piecewise constant hazards} on five-year intervals: 40--45, 45--50, \ldots, 85--90 ("age").

\item Number of cells: $2 \times 10 \times 10 \times 20 \times 2 \times 3 \times 4 = $ \emp{96000}

\item Number of \emp{non-empty} cells: \emp{\Sexpr{nr}}.

\end{itemize}

\url{http://capa.ddb.umu.se/ds/aggregate.html}

\end{frame}

\begin{frame}[fragile]{Exposure by hisclass, men}

<<exphiswom>>=
xx <- with(vb0Tab[vb0Tab$sex == "male", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
yy <- round(prop.table(xx, 1) * 100, 1)
who <- length(levels(vb0Tab$hisclass))
barplot(t(yy)[who:1, ], col = c("blue", "red", "green", "black"), las = 2, ylab = "Per cent")
@

\end{frame}

\begin{frame}[fragile]{Exposure by hisclass, women}

<<exphis>>=
xx <- with(vb0Tab[vb0Tab$sex == "female", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
yy <- round(prop.table(xx, 1) * 100, 1)
who <- length(levels(vb0Tab$hisclass))
barplot(t(yy)[who:1, ], col = c("blue", "red", "green", "black"), las = 2, ylab = "Per cent")
@

\end{frame}

\begin{frame}{Age--Period--Cohort data}

\vspace{-1.5cm}

<<moddat,fig.height=4.5,fig.width=5>>=
x <- c(1961, 2014)
y <- c(40, 40)
plot(x, y, type = "l", col = "blue", lwd = 2, xlim = c(1960, 2015), ylim = c(40, 90),
     ylab = "Age", xlab = "Year", axes = FALSE)
axis(1, at = c(1961, 1991, 2014))
axis(2, las = 1)
box()
lines(c(1961, 1961), c(40, 65), col = "blue", lwd = 2)
lines(c(2014, 2014), c(40, 90), col = "blue", lwd = 2)
lines(c(1961, 1986), c(65, 90), col = "blue", lwd = 2)
lines(c(1986, 2014), c(90, 90), col = "blue", lwd = 2)
#
lines(c(1971, 1971), c(40, 75), col = "blue", lty = 2)
lines(c(1981, 1981), c(40, 85), col = "blue", lty = 2)
lines(c(1991, 1991), c(40, 90), col = "blue", lty = 2)
lines(c(2002, 2002), c(40, 90), col = "blue", lty = 2)
lines(c(2008, 2008), c(40, 90), col = "blue", lty = 2)
#
for (age in seq(40, 65, by = 5)){
    lines(c(1961, 2014), c(age, age), col = "blue", lty = 2)
}
for(i in 0:3){
    age <- 70 + 5 * i
    lines(c(1961 + 5 * (i + 1), 2014), c(age, age), col = "blue", lty = 2)
}
#
lines(c(1961, 1961 + 30), c(60, 90), lty = 2, col = "red")
lines(c(1961, 1961 + 40), c(50, 90), lty = 2, col = "red")
lines(c(1961, 1961 + 50), c(40, 90), lty = 2, col = "red")
lines(c(1971, 2014), c(40, 40 + 43), lty = 2, col = "red")
lines(c(1981, 2014), c(40, 40 + 33), lty = 2, col = "red")
lines(c(1991, 2014), c(40, 40 + 23), lty = 2, col = "red")
lines(c(2001, 2014), c(40, 40 + 13), lty = 2, col = "red")
lines(c(2011, 2014), c(40, 40 + 3), lty = 2, col = "red")
@

\end{frame}

\begin{frame}[fragile]{Counting "cells"}

<<cremodern,echo=FALSE>>=
library(eha)
##vb0$hisclass[is.na(vb0$hisclass)] <- 4 ## NOTE!!
##mode <- cal.window(vb0, c(1961, 2014))
##modern <- mode[mode$birthdate > 1896, ]
##modern$hisclass <- factor(modern$hisclass)
##modern$period <- factor(modern$period)
##moTab <- aggregate(modern[, c("event", "exposure")],
  ##                 modern[, c("sex", "age", "period", "cohort",
    ##                          "urban", "civst", "hisclass")],
      ##             FUN = sum)
moTab <- vb0Tab[vb0Tab$period %in% levels(vb0Tab$period)[7:10], ]
moTab$period <- factor(moTab$period)
moTab$cohort <- factor(moTab$cohort)
moTab$age <- factor(moTab$age)
moTab$hisclass <- factor(moTab$hisclass)
moTab$hisclass <- relevel(moTab$hisclass, ref = "middle")
moTab$civst <- relevel(moTab$civst, ref = "married")
##
olTab <- vb0Tab[vb0Tab$period %in% levels(vb0Tab$period)[2:4], ]
olTab$period <- factor(olTab$period)
olTab$cohort <- factor(olTab$cohort)
olTab$age <- factor(olTab$age)
olTab$hisclass <- factor(olTab$hisclass)
olTab$hisclass <- relevel(olTab$hisclass, ref = "middle")
olTab$civst <- relevel(olTab$civst, ref = "married")

@

\begin{itemize}
\item There are \emp{115 cells in one figure}
\item There are $2 \times 2 \times 3 \times 4 = $ \emp{48 such figures}.
\item In total: $115 \times 48 = $ \emp{5520 combinations}.
\item Number of \emp{non-empty} combinations: \emp{4072}.
\end{itemize}

{\scriptsize
<<heada>>=
head(moTab)
@
}

An \emp{anonymous} data set!
\end{frame}

\begin{frame}[fragile]{PH regression with the table}

\begin{itemize}
\item Piecewise constant hazard model generates the same likelihood as a Poisson model.
\begin{itemize}
\item Does \emp{not} imply that the number of deaths follow a Poisson distribution!
\end{itemize}
\end{itemize}

\small

<<coxpois,echo=TRUE,cache=FALSE>>=
system.time(fitp <- glm(event ~ offset(log(exposure)) + age +
                hisclass * period + civst + urban + sex,
            data = vb0Tab, family = poisson))
system.time(dr <- drop1(fitp, test = "Chisq"))
@

\end{frame}

\begin{frame}{ANOVA with the table}

<<anovatable,echo=FALSE>>=
knitr::kable(dr)
@

\end{frame}

\begin{frame}[fragile]{The 'baseline hazard function'}

{\scriptsize
<<extab,echo=FALSE>>=
knitr::kable(summary(fitp)$coefficients[1:10, 1:2])
@
}

\begin{itemize}
\item \emp{(Intercept)} is log hazard in 40--45. \emp{Reference} interval.
\item \emp{age45} is the \emp{difference} in log hazard between 45--50 and 40--45,
\item \emp{age50} is the \emp{difference} in log hazard between 50--55 and 40--45,
\item Etcetera \ldots
\end{itemize}

\end{frame}

\begin{frame}[fragile]{Baseline hazards}

<<fixthem,fig.height = 3.4>>=
alpha <- summary(fitp)$coefficients[1:10, 1]
alpha[2:10] <- alpha[2:10] + alpha[1]
oldpar <- par(mfrow = c(1, 2))
plot(seq(40, 85, by = 5), alpha, col = "red", xlab = "Age", main = "log(hazard)", ylab = "", axes = FALSE)
axis(1)
axis(2, las = 1)
axis(1)
box()
x <- seq(40, 85, by = 5)
plot(x, exp(alpha), col = "blue", type = "b", xlab = "Age", main = "hazard", ylab = "", axes = FALSE)
axis(1)
axis(2, las = 1)
box()
abline(h = 0)
par(oldpar)
##ff <- lm(alpha ~ x)
##plot(ff)
@

Why not try:

\begin{equation*}
\log\{h(t)\} = \alpha + \beta t, \quad 40 < t \le 90,
\end{equation*}

a \emp{Gompertz} distribution?

\end{frame}

%%\begin{frame}{Period analysis, men}

%%<<coxper,eval=FALSE>>=
%%oldpar <- par(mfrow = c(2, 3))
%%men <- modern[modern$sex == "male", ]
%%##levels(men$hisclass) <- c("elite", "middle", "worker", "NA")
%%cols = c("black", "darkgreen", "red", "blue")
%%for (per in levels(men$period)){
%%    min <- men[men$period == per, ]
%%   fit <- coxreg(Surv(enter, exit, event) ~ strata(hisclass) + urban * civst, data = min)
%%   plot(fit, col = cols, main = paste("Men", per), fn = "surv", printLegend = FALSE, xlab = "Age", xlim = c(40, 90))
%%   legend("bottomleft", legend = levels(min$hisclass), col = cols, lty = 1)
%%   abline(h = 0.5, lty = 3)
%%}
%%par(oldpar)
%%@

%%\end{frame}

\begin{frame}[fragile]{Survival analysis, men 1961-2013}

<<coxper2>>=
source("R/plot_surv.R")
##oldpar <- par(mfrow = c(2, 3))
men <- moTab[moTab$sex == "male", ]
cols <- c("darkgreen", "black", "red", "blue")
##men$hisclass <- relevel(men$hisclass, ref = "middle")
plot_surv(men, sex = "Men")
@

\end{frame}

\begin{frame}[fragile]{Median life after 40, men 1961-2013}

<<coxpertab>>=
source("R/mean_pch.R")
source("R/median_pch.R")
source("R/plot_median.R")

##pop <- vb0Tab[!(vb0Tab$period %in% oldperiod) &
  ##                vb0Tab$sex == "male" &
    ##              vb0Tab$civst != "unmarried" &
      ##            !vb0Tab$urban, ]
##pop$hisclass <- relevel(pop$hisclass, ref = "middle")
##pop$civst <- relevel(pop$civst, ref = "married")
##pop$period <- factor(pop$period)
##pop$age <- factor(pop$age)
plot_median(men)
@

\end{frame}

\begin{frame}{Survival analysis, women 1961-2013}

<<coxperwom>>=
par(mfrow = c(2, 3))
women <- moTab[moTab$sex == "female", ]
##levels(men$hisclass) <- c("elite", "middle", "worker", "NA")
##cols = c("black", "green", "red", "blue")
plot_surv(women, sex = "Women")
##for (per in levels(women$period)){
##    min <- women[women$period == per, ]
##   fit <- coxreg(Surv(enter, exit, event) ~ strata(hisclass) + urban * civst, data = min)
##   plot(fit, col = cols, main = paste("Women,", per), fn = "surv", printLegend = FALSE, xlab = "Age", xlim = c(40, 90))
##   legend("bottomleft", legend = levels(min$hisclass), col = cols, lty = 1)
##}
@

\end{frame}

\begin{frame}[fragile]{Median life after 40, women 1961-2013}
<<coxpertabwom,warning = FALSE>>=
source("R/mean_pch.R")
source("R/median_pch.R")
##oldperiod <- levels(vb0Tab$period)[1:4]
##pop <- modern[modern$sex == "female", ]
##pop <- vb0Tab[!(vb0Tab$period %in% oldperiod) &
  ##                vb0Tab$sex == "male" &
    ##              vb0Tab$civst != "unmarried" &
      ##            !vb0Tab$urban, ]
##pop$hisclass <- relevel(pop$hisclass, ref = "middle")
##pop$civst <- relevel(pop$civst, ref = "married")
##pop$period <- factor(pop$period)
##pop$age <- factor(pop$age)
plot_median(women)
@

\end{frame}

\begin{frame}[fragile]{Median life after 40, men 1851-1950}

<<medaf40olmen>>=
men <- olTab[olTab$sex == "male", ]
plot_median(men)
@

\end{frame}

\begin{frame}[fragile]{Median life after 40, women 1851-1950}

<<medaf40olwomen,warning=FALSE>>=
women <- olTab[olTab$sex == "female", ]
plot_median(women)
@

\end{frame}

%\begin{frame}[frame]{Income, men, 1990-2005}

%<<incomemen>>=
%source("R/plot_income.R")
%load("data/loTab.rda")
%loTab$period <- factor(loTab$period)
%men <- loTab[loTab$sex == "male",]
%plot_income(men)
%@

%\end{frame}

% \begin{frame}{Conclusion}

% \begin{itemize}
% \item Most striking: Very bad prospects for \emp{unemployed}, both men and women, all periods.
% \item Apart from unemployed: Equal opportunities for men 1960--80, for women 1960--90, but after that a clear separation with the \emp{elite on top}.
% \end{itemize}

% \vspace{2\baselineskip}

% After this slide: \emp{Old stuff!}

% \end{frame}

\begin{frame}{Proportional hazards for women?}

\includegraphics[width=11cm,height=8cm]{figure/hazplotw-1}

\end{frame}

\begin{frame}{Proportional hazards for men?}

\includegraphics[width=11cm,height=8cm]{figure/hazplotm-1}

\end{frame}

% \begin{frame}[fragile]{Interactions}

% \begin{verbatim}
% fit <- coxph(Surv(enter - 40, exit - 40, event) ~
%     period * hisclass * sex + urban, data = vb0)
% drop1(fit, test = "Chisq")
% \end{verbatim}

% \begin{center}
% \begin{tabular}{lrrrr}
%   \hline
%  & Df & AIC & LRT & Pr($>$Chi) \\
%   \hline
% $<$none$>$ &  & 1631901.32 &  &  \\
%   urban & 1 & 1631946 & 47.04 & $7.0\times 10^{-12}$ \\
%   \emp{period:hisclass:sex} & 27 & 1631957 & 110.11 & \emp{$5.3\times 10^{-12}$} \\
%    \hline
% \end{tabular}
% \end{center}

% \begin{itemize}

% \item This model is described by \emp{80 coefficients}.

% \item We show them in \emp{graphs}.
% \end{itemize}

% \end{frame}

\begin{frame}{No proportional hazards (wrt \sc Hisclass)!}

\begin{itemize}
\item Fit each {\sc Hisclass} \emp{separately}, for each time period.
\item Use the \emp{cumulative hazard} at 65 (or 90) for comparison.
\end{itemize}

So, we fit separate PH models for each combination of

\begin{itemize}
  \item \emp{\tt sex}
  \item \emp{\tt period}: 1851--75, 1876--1900, 1901--10, etc.
  \item \emp{\tt hisclass}
  \item \emp{\tt age group}: 40--64, 65--89.
  \item \emp{\tt cause of death}
\end{itemize}
\end{frame}

% \begin{frame}{The data table}

% <<lexis3, fig.height = 4>>=
% plot(c(1851, 1851, 1951, 1951, 1851), c(40, 90, 90, 40, 40),
%      type = "l", xlim = c(1851, 2014), lwd = 2, col = "red",
%      xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(40, 90))
% axis(1, at = c(1851, 1901, 1951, 1961, 1981, 2014), las = 2)
% axis(2, las = 2, at = c(0, 40, 65, 90))
% axis(4, las = 2, at = c(0, 40, 65, 90))
% box()
% abline(h = 0)
% lines(c(1961, 1961, 2014, 2014, 1961), c(40, 90, 90, 40, 40),
%       col = "blue", lwd = 2)
% for (i in c(1851, 1876, 1901, 1926, 1971, 1981, 1991, 2002, 2007)){
%     lines(c(i, i), c(40, 90), lty = 2)
% }
% for (i in c(45, 50, 55, 60, 65, 70, 75, 80, 85)){
%     lines(c(1851, 1951), c(i, i), lty = 2)
%     lines(c(1961, 2014), c(i, i), lty = 2)
% }
% @

% \emp{Poisson regression}:

% In \emp{each cell}, and for \emp{each} combination of \emp{covariate level}s, calculate
% \begin{itemize}
% \item  the number of \emp{deaths} and
% \item the total \emp{exposure time}.
% \end{itemize}
% \end{frame}

\begin{frame}{Women and men, age 40--64}

  \begin{center}
\includegraphics[height = 7.5cm]{figure/cumsw-1}
  \end{center}
\end{frame}

\begin{frame}{Relative differences, ages 40--64 and 65--89}

  %\begin{center}
\includegraphics[height = 7.5cm]{figure/cumswstand-1}\includegraphics[height=7.5cm]{figure/cumswold-1}
   %\end{center}
\end{frame}

\end{document}
\begin{frame}{Cancer mortality, ages 40-64 and 65-89}

\includegraphics[height=7.5cm]{figure/12cumsw-1}\includegraphics[height=7.5cm]{figure/12cumswold-1}

\end{frame}

\begin{frame}{Cardiovascular mortality}

\includegraphics[height=7.5cm]{figure/41cumsw-1}\includegraphics[height=7.5cm]{figure/41cumswold-1}

\end{frame}

\begin{frame}[fragile]{Education 1990--2013}

<<educ>>=

educ <- lisaold[!(lisaold$period %in% c("1986-1987", "1988-1989")), ]
educ$period <- factor(educ$period)
educ <- educ[!is.na(educ$educ), ]
levels(educ$educ) <- c("low", "low", "middle", "middle", "middle", "high", "high")
educ$educ <- relevel(educ$educ, ref = "middle")
educ$civst <- relevel(educ$civst, ref = "married")
fitm <- coxph(Surv(enter, exit, event) ~ strata(educ) + civst + urban,
             data = educ[educ$sex == "male", ])
fitf <- coxph(Surv(enter, exit, event) ~ strata(educ) + civst + urban,
             data = educ[educ$sex == "female", ])
cols <- c("darkgreen", "red", "black")
@

<<ploteduc>>=
oldpar <- par(mfrow = c(1, 2))

plot(survfit(fitm, start.time = 40), mark.time = FALSE, main = "Men",
     col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
axis(1)
axis(2, at = c(0, 0.5, 1.0), las = 1)
abline(h = 0.5, lty = 3)
box()
legend("bottomleft", legend = levels(educ$educ), col = cols, lty = 1, lwd = 1.5)

plot(survfit(fitf, start.time = 40), mark.time = FALSE, main = "Women",
     col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
axis(1)
axis(2, at = c(0, 0.5, 1.0), las = 1)
abline(h = 0.5, lty = 3)
box()
legend("bottomleft", legend = levels(educ$educ), col = cols, lty = 1, lwd = 1.5)

par(oldpar)
@

\end{frame}

\begin{frame}{Income 1986--2005}

<<femeduc>>=
income <- lisaold[!is.na(lisaold$income), ]
income$period <- factor(income$period)
##income <- income[!is.na(income$income), ]
##levels(income$income) <- c("low", "low", "middle", "middle", "middle", "high", "high")
income$income <- relevel(income$income, ref = "low")
income$civst <- relevel(income$civst, ref = "married")
fitm <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
             data = income[income$sex == "male", ])
fitf <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
             data = income[income$sex == "female", ])
cols <- c("darkgreen", "red", "blue", "black")

@

<<plotincome>>=
oldpar <- par(mfrow = c(1, 2))

plot(survfit(fitm, start.time = 40), mark.time = FALSE, main = "Men",
     col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
axis(1)
axis(2, at = c(0, 0.5, 1.0), las = 1)
abline(h = 0.5, lty = 3)
box()
legend("bottomleft", legend = levels(income$income), col = cols, lty = 1, lwd = 1.5)

plot(survfit(fitf, start.time = 40), mark.time = FALSE, main = "Women",
     col = cols, lwd = 1.5, axes = FALSE, xlab = "Age", ylab = "Surviving fraction")
axis(1)
axis(2, at = c(0, 0.5, 1.0), las = 1)
abline(h = 0.5, lty = 3)
box()
legend("bottomleft", legend = levels(income$income), col = cols, lty = 1, lwd = 1.5)

par(oldpar)
@

\end{frame}

\begin{frame}[fragile]{Income by education, 1990--2005, ages 40--75!}

<<incomeandeduc>>=
both <- income[!is.na(income$educ), ]
levels(both$educ) <- c("low", "low", "middle", "middle", "middle", "high", "high")
both$educ <- relevel(both$educ, ref = "middle")
bmal <- both[both$sex == "male", ]
bfem <- both[both$sex == "female", ]
fitme1 <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
                data = bmal[bmal$educ == "low", ])
fitme2 <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
                data = bmal[bmal$educ == "middle", ])
fitme3 <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
                data = bmal[bmal$educ == "high", ])
##
fitfe1 <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
                data = bfem[bfem$educ == "low", ])
fitfe2 <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
                data = bfem[bfem$educ == "middle", ])
fitfe3 <- coxph(Surv(enter, exit, event) ~ strata(income) + civst + urban,
                data = bfem[bfem$educ == "high", ])

@

<<plotedin>>=
oldpar <- par(mfrow = c(2, 3))
ylim = c(0.5, 1)
plot(survfit(fitme1, start.time = 40), col = cols, main = "Men, low education", xlab = "Age", ylim = ylim)
legend("bottomleft", legend = levels(income$income), col = cols, lty = 1, lwd = 1.0)

plot(survfit(fitme2, start.time = 40), col = cols, main = "Men, average education", xlab = "Age", ylim = ylim)
plot(survfit(fitme3, start.time = 40), col = cols, main = "Men, high education", xlab = "Age", ylim = ylim)
##
plot(survfit(fitfe1, start.time = 40), col = cols, main = "Women, low education", xlab = "Age", ylim = ylim)
legend("bottomleft", legend = levels(income$income), col = cols, lty = 1, lwd = 1.0)

plot(survfit(fitfe2, start.time = 40), col = cols, main = "Women, average education", xlab = "Age", ylim = ylim)
plot(survfit(fitfe3, start.time = 40), col = cols, main = "Women, high education", xlab = "Age", ylim = ylim)
par(oldpar)
@

\end{frame}

%%\begin{frame}{Education and income, 1990--2005, all deaths}

%%\includegraphics[height=3.5cm]{figure/liswomfig-1} Women, 40--64.

%%\includegraphics[height=3.5cm]{figure/lismanfig-1} Men, 40--64.

%%\end{frame}

%%\begin{frame}{Education and income, 1990--2005, all deaths}

%%\includegraphics[height=3.5cm]{figure/liswomfig2-1} Women, 65--89.

%%\includegraphics[height=3.5cm]{figure/lismanfig2-1} Men, 65--89.

%%\end{frame}

% \begin{frame}{Education and income, 1990--2005, cancer}

% \includegraphics[height=3.5cm]{figure/12liswomfig-1} Women, 40--64.

% \includegraphics[height=3.5cm]{figure/12lismanfig-1} Men, 40--64.

% \end{frame}

% \begin{frame}{Education and income, 1990--2005, cancer}

% \includegraphics[height=3.5cm]{figure/12liswomfig2-1} Women, 65--89.

% \includegraphics[height=3.5cm]{figure/12lismanfig2-1} Men, 65--89.

% \end{frame}

% \begin{frame}{Education and income, 1990--2005, cardiovascular}

% \includegraphics[height=3.5cm]{figure/41liswomfig-1} Women, 40--64.

% \includegraphics[height=3.5cm]{figure/41lismanfig-1} Men, 40--64.

% \end{frame}

% \begin{frame}{Education and income, 1990--2005, cardiovascular}

% \includegraphics[height=3.5cm]{figure/41liswomfig2-1} Women, 65--89.

% \includegraphics[height=3.5cm]{figure/41lismanfig2-1} Men, 65--89.

% \end{frame}

%%% Stop here!

\begin{frame}{Answers(?)}

\begin{enumerate}
\item Has \emp{inequality in mortality} between \emp{social classes}
  increased?
\begin{itemize}
\item Not really
\end{itemize}
%%\item Is the mortality transition initiated by a certain social class?

\item Are there any \emp{gender differences} in the effect of social position?
\begin{itemize}
\item Yes
\end{itemize}

\item Is social position \emp{equally important} among the \emp{retired} population as
  for those in \emp{working age}?
\begin{itemize}
\item Not really
\end{itemize}

\item Is there a difference between classes in mortality from
  \emp{cardiovascular} diseases and \emp{cancers} respectively?
\begin{itemize}
\item Yes
\end{itemize}
\item Are there large differences in survival depending on level of
  \emp{education} and \emp{income} categories?
\begin{itemize}
\item Yes, definitely.
\end{itemize}

\end{enumerate}

\end{frame}

\begin{frame}{Conclusion}

%\scriptsize

\begin{itemize}
\item \emp{Improved} survival characterised all {\sc Hisclass}es.
\item There are \emp{gender differences} in the social pattern
\item No difference in the social pattern in the two studied age groups.
\item The \emp{upper} class(es) gradually take over as the \emp{most healthy} group.
\item \emp{Education} and \emp{income} are the \emp{most important}
  variables during late
  time period.
\end{itemize}

\end{frame}

% \begin{frame}{Software for report writing and analysis}

% \begin{itemize}
% \item  \R : \emp{\url{https://www.R-project.org}}.
%   \begin{itemize}
%     \item \emp{RStudio}: \url{https://www.rstudio.com}.
%   \end{itemize}
% \item \R\ \emp{packages} (\emp{\url{https://cran.R-project.org}}):
%   \begin{itemize}
%   \item \emp{skum} (Broström, 2016) data from \emp{DDB}.
%   \item \emp{rmarkdown} (RStudio Team, 2016).
%   \item \emp{knitr} (Xie, 2016).
%   \item \emp{eha} (Broström, 2016).
%   \item \emp{xtable} (Dahl, 2016).
% \end{itemize}

% \item \emp{Version control} (reproducibility)
%   \begin{itemize}
%     \item \emp{git}: \url{http://git}
%     \item \emp{gitHub}: \url{http://github.com/goranbrostrom/}

% \end{itemize}
% \end{itemize}

% \end{frame}

\end{document}