### 2018-02-28 CreData/fromPoplink.Rmd
fix hisclass

---
title: "From Poplink"
author: "Göran Broström"
date: "28 February 2018"
output:
  pdf_document: default
  html_document: default
---

## Introduction

In this version we start at January 1, 1801, and we **do not** split the *poplink* data 
into two data sets (no causes of death, yet). End at December 31, 1950

```{r setup}
knitr::opts_chunk$set(cache=FALSE,echo=TRUE,message=TRUE,warning=TRUE,comment=NA)
library(xtable)
library(knitr)
library(eha)
library(skum)
##library(dplyr)
##library(tidyr)
##library(stringr)
```

## PopLink data

The POPLINK data cover 1801--1950. The resulting output consists of two tables,
**skel** and **skume** (I avoid the natural name **skum** in order to avoid a
mix-up with the **R** package by that name).
A covariate **period** is added and used for 
splitting the data into calendar time periods of length 10 years each.

### Get data

```{r getdata}
library(skum)
##load("../../Data/adult.rda")
##adult <- cal.window(adult, c(1801, 1951))
adult <- obs
indx <- match(adult$id, per$id)
adult$mid <- per$mid[indx]
adult$fid <- per$fid[indx]
adult$event <- !is.na(adult$sluttyp) & adult$sluttyp == 2
```

## Put on hisclass and civil status

```{r hisclass}
##New Hisclasses:
ett <- with(adult, !is.na(socBranch) &
              (hisclass == 1))
               ## ((socBranch == "official" & socStatus == "high") |
                 ##    (socBranch == "business" & socStatus == "high")
                ##))

tva <- with(adult, !is.na(socBranch) &
              (hisclass %in% c(2, 4)))
                ##((socBranch == "farming" & socStatus == "high") |
                ## (socBranch == "business" & socStatus == "low") |
                ## (socBranch == "official" & socStatus == "low")
                ##))

tre <- with(adult, !is.na(socBranch) &
              (hisclass %in% 5:7)) ## Original 'fyra' ---> 'tre'
              ##  ((socBranch == "farming" & socStatus == "low") |
                ##     (socBranch == "worker"))
                ##)
fyra <- with(adult, !is.na(socBranch) &
               (hisclass == 3))

hisc <- rep(-1, NROW(adult))
hisc[ett] <- 1
hisc[tva] <- 2
hisc[tre] <- 3
hisc[fyra] <- 4
is.na(hisc) <- hisc == -1
adult$hisclass <- hisc
```

Continue ...

```{r urbanetc}
adult$urban <- with(adult, !is.na(ortnmn) & (ortnmn %in% c("UMEÅ", "SKELLEFTEÅ STAD")))
levels(adult$civst) <- c("unmarried", "married", "prev.married", "prev.married", "prev.married")
```

## Propagate

```{r propagate}
##if (file.exists("adult.rda")){
  ##load("adult.rda")
if (TRUE){
source("R/fixCovar2.R")  
    adult$hisclass40 <- with(adult, fixCovar2(id, hisclass, enter, exit, 40, 20))
    adult$hisclass40[is.na(adult$hisclass40)] <- 5
    adult$hisclass40 <- factor(adult$hisclass40, labels = c("elite", "middle", "worker", "farmer", "none"))
    adult$hisclass65 <- with(adult, fixCovar2(id, hisclass, enter, exit, 65, 45))
    adult$hisclass65[is.na(adult$hisclass65)] <- 5
    adult$hisclass65 <- factor(adult$hisclass65, labels = c("elite", "middle", "worker", "farmer", "none"))

    adult$civst40 <- with(adult, fixCovar2(id, civst, enter, exit, 40, 20))
    adult$civst65 <- with(adult, fixCovar2(id, civst, enter, exit, 65, 45))

    adult$urban40 <- with(adult, fixCovar2(id, civst, enter, exit, 40, 20))
    adult$urban65 <- with(adult, fixCovar2(id, civst, enter, exit, 65, 45))

}else{
  adult <- adult[order(adult$id, adult$enter), ]
  adult <- group_by(adult, id) %>% fill_("hisclass")
  adult$hisclass[is.na(adult$hisclass)] <- 4
  adult$hisclass <- factor(adult$hisclass, labels = c("elite", "middle", "worker", "NA"))
  ##
  adult <- group_by(adult, id) %>%  fill_("civst")
  adult$civst[is.na(adult$civst)] <- "unmarried"
  save(adult, file = "adult.rda")
}
```

## The Lexis cut

Our data covers the years 1801--1950 and the ages 40--99.

```{r lexis}
##adult <- cal.window(adult, c(1801, 1951))
adult <- age.window(adult, c(40, 100))
```

## Split and period

```{r splitper}
source("R/addPer.R")
##skel <- cal.window(adult, c(1801, 1901))
##skel <- addPer(skel, cuts = seq(1801, 1901, by = 25))
skume <- cal.window(adult, c(1801, 1951))
skume <- addPer(skume, cuts = seq(1801, 1951, by = 25))
```

### Trim skume

```{r skumefilt}
skume <- skume[, c("id", "region", "birthdate", "sex", "enter", "exit", "event",
                    "period", "urban", "urban40", "urban65", "civst", "civst40", "civst65", 
                    "hisclass", "hisclass40", "hisclass65")]
skume <- skume[order(skume$id, skume$exit), ]
skume <- rc(skume)
indx <- tapply(skume$id, skume$id)
Exit <- with(skume, tapply(exit, id, max))
Enter <- with(skume, tapply(enter, id, min))
Event <- with(skume, tapply(event, id, max))
skume$Enter <- Enter[indx]
skume$Exit <- Exit[indx]
skume$Event <- Event[indx]
skume <- skume[!duplicated(skume$id), ]
skume$enter <- skume$Enter
skume$exit <- skume$Exit
skume$event <- skume$Event
skume$id <- paste("pop", skume$id, sep = "_")
skume$Exit <- NULL
skume$Enter <- NULL
skume$Event <- NULL
```

## Save

```{r save}
##save(skel, file = "skel.rda")
save(skume, file = "skume.rda")
```