### 2018-03-03 CreData/fromFobLisa.Rmd
continuing

---
title: "fromFobLisa, Version 2"
author: "Göran Broström"
date: "March 1, 2018"
output:
  html_document:
    df_print: paged
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
options(stringsAsFactors = FALSE)
```

# Introduction

Now *fromFobLisa.Rmd* is updated to include *fob90* and also let the
follow-up for *fob85* be *five years* and the follow-up for *fob90* be *ten
years*.

This is an updated version of *showData4.Rmd*, where we used the newly-found
knowledge that the yearly *lisa* data relates to *December 31* the given
year, while the *FoB* data relate to the date *November 1* the given
year. These facts have the following implications for our plan: For *FoB*
data, we follow individuals from their appearance in a FoB until the next
FoB five years later, even if they do not appear there. The exception is if
they die within five years after the first appearance. So, *fob60* induces
a time interval *from November 1, 1960 to October 31, 1965*, and so on. 
Since we do not have birth dates but only birth year and quarter,
we decide to use as birthdate the *midpoint of the birth quarter*. 
WE work with all ages 0-110, and only as the final step, `age.window` is called.

The following FoBs are treated similarly, until we reach the Fob
in 1985. Since the first *lisa* date is December 31, 1986, this "fob
interval" starts at November 1, 1985, and ends at December 31, 1986, that
is, it is 14 months long.

# Fixed data

We start with the *inddata* file and treat it as a *fixed-data*
file (no time-varying info). 
We find the birth dates (see above), death dates (for those who are observed die), and the "date last seen".

Note that 37 individuals in *inddata* have no information at all, so they are removed.

```{r frominddata}
load("../../Data/inddata.rda")
load("../../Data/sex.rda")
indx <- match(inddata$LINNEID, sex$LINNEID)
inddata$KON <- sex$KON[indx]
per <- inddata[!is.na(inddata$KON), ]
per$birthdate <- per$FODELSEAR + 0.125  +  (per$FODELSEKVARTAL - 1) / 4
```

Now we put on *death date*, *death cause*, and *KOMMUN of death*, from *socstyr*, on *per*:

```{r styronper}
load("../../Data/socstyr.rda")
indx <- match(per$LINNEID, socstyr$LINNEID)
per$DODSDAT <- socstyr$DODSDAT[indx]
fixd <- !is.na(per$DODSDAT) & (per$DODSDAT < 1960010)
per$DODSDAT[fixd] <- per$DODSDAT[fixd] * 100 + 15
per$deathdate <- as.numeric(skum::tillTid(per$DODSDAT))
per$dodsKOMMUN <- socstyr$KOMMUN[indx]
per$dodsALDER <- socstyr$ALDER[indx]
per$ULORSAK <- socstyr$ULORSAK[indx]
```

# The FoBs

Through the *FoB*s we have snapshots of the population on *November 1* the years *1960(5)1990*, giving the dates (decimal form).

```{r days}
load("../../Data/fob.rda")
ar <- seq(1960, 1990, by = 5)
days <- as.numeric(skum::tillTid(paste(ar, "1101", sep = "")))
days
```

We have to limit the data collection to the *KOMMUN*er that are contained
in our area, that is

```{r kommuner}
ourkom <- c(2406, 2414, 2417, 2418, 2480, 2482)
ourkom <- c(ourkom, 2411, 2412, 2413, 2416, 2415, 2461, 2407, 2402, 2408)
fob <- fob[fob$KOMMUN %in% ourkom, ]
```

Get `birthdate` from `per`.

```{r getbirthdateanddathdate}
indx <- match(fob$LINNEID, per$LINNEID)
fob$birthdate <- per$birthdate[indx]
fob$deathdate <- per$deathdate[indx]
```

Create individual Fobs (1960, 1965, etc):

```{r rawfob}
source("R/extractFob.R")
n <- length(ar)
fobbar <- vector(mode = "list", length = n)
names(fobbar) <- paste("fob", (ar %% 100), sep = "")
for (i in 1:(n-1)){
    fobbar[[i]] <- extractFob(ar[i], followUp = 5, data = fob)
}
fobbar[[n]] <- extractFob(ar[n], followUp = 11 + 61 / 365, data = fob)
fobdata <- do.call(rbind, fobbar)
fobdata <- fobdata[order(fobdata$LINNEID, fobdata$enter), ]
##fobdata <- age.window(fobdata, c(40, 110))
rownames(fobdata) <- 1:NROW(fobdata)
save(fobdata, file = "fobdata2.rda")
```

# The Lisa data

The *lisa* data is slightly simpler because it is yearly with cutoff dates 12-31/01-01 each new-year. It starts off at 1987-01-01 (lisa1986) and ends at 2013-12-31 (lisa2013). So the time interval a *lisa* covers is *AR + 1*! But note that the *socstyr* data only cover deaths *before January 1, 2014, so *lisa2013* is in effect useless to us.

```{r lisadat}
load("../../Data/lisa.rda")
lisa <- lisa[lisa$KOMMUN %in% ourkom, ]
indx <- match(lisa$LINNEID, per$LINNEID)
lisa$birthdate <- per$birthdate[indx]
```

Those with *missing birthdate* in *lisa* are those born 1975 or later. They can be removed.

```{r removeyoung}
lisa <- lisa[!is.na(lisa$birthdate), ]
```

So, in *lisa*, all enter in *AR + 1*, at the exact age of *(AR + 1 - birthdate)*, and exit at the exact age *enter + 1* if they do not die in *the year AR + 1*, otherwise at their death date, or at *deathdate - birthdate*.

```{r fixlisaintervals}
lisa$enter <- lisa$AR + 1 - lisa$birthdate
indx <- match(lisa$LINNEID, per$LINNEID) # To be safe ...
lisa$deathdate <- per$deathdate[indx]
lisa$event <- !is.na(lisa$deathdate) & lisa$deathdate > lisa$AR + 1 & lisa$deathdate <= lisa$AR + 2
lisa$exit <- lisa$enter + 1 # First tip.
lisa$exit[lisa$event] <- with(lisa, deathdate[event] - birthdate[event])
```

Now some sorting  and cleaning.

```{r cleanandsort}
##lisadata <- age.window(lisa, c(40, 110))
lisadata <- lisa[order(lisa$LINNEID, lisa$enter), ]
lisadata <- cal.window(lisadata, c(1987, 2014)) # 'Remove lisa2013'
rownames(lisadata) <- 1:NROW(lisadata)
save(lisadata, file = "lisadata.rda")
```

Now, read *checkData* (maybe not).