### 2018-03-04 CreData/fromFobLisa.Rmd
last minute

---
title: "fromFobLisa, Version 2"
author: "Göran Broström"
date: "March 1, 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
options(stringsAsFactors = FALSE)
library(stringr)
library(tidyr)
library(dplyr)
```

# Introduction

Now *fromFobLisa.Rmd* is updated to include *fob90* and also let the
follow-up for *fob85* be *five years* and the follow-up for *fob90* be *ten
years*.

This is an updated version of *showData4.Rmd*, where we used the newly-found
knowledge that the yearly *lisa* data relates to *December 31* the given
year, while the *FoB* data relate to the date *November 1* the given
year. These facts have the following implications for our plan: For *FoB*
data, we follow individuals from their appearance in a FoB until the next
FoB five years later, even if they do not appear there. The exception is if
they die within five years after the first appearance. So, *fob60* induces
a time interval *from November 1, 1960 to October 31, 1965*, and so on. 
Since we do not have birth dates but only birth year and quarter,
we decide to use as birthdate the *midpoint of the birth quarter*. 
WE work with all ages 0-110, and only as the final step, `age.window` is called.

The following FoBs are treated similarly, until we reach the Fob
in 1985. Since the first *lisa* date is December 31, 1986, this "fob
interval" starts at November 1, 1985, and ends at December 31, 1986, that
is, it is 14 months long.

# Fixed data

We start with the *inddata* file and treat it as a *fixed-data*
file (no time-varying info). 
We find the birth dates (see above), death dates (for those who are observed die), and the "date last seen".

Note that 37 individuals in *inddata* have no information at all, so they are removed.

```{r frominddata}
load("../../Data/inddata.rda")
load("../../Data/sex.rda")
indx <- match(inddata$LINNEID, sex$LINNEID)
inddata$KON <- sex$KON[indx]
per <- inddata[!is.na(inddata$KON) & !is.na(inddata$FODELSEAR), ]
per$birthdate <- per$FODELSEAR + 0.125  +  (per$FODELSEKVARTAL - 1) / 4
```

Now we put on *death date*, *death cause*, and *KOMMUN of death*, from *socstyr*, on *per*:

```{r styronper}
load("../../Data/socstyr.rda")
indx <- match(per$LINNEID, socstyr$LINNEID)
per$DODSDAT <- socstyr$DODSDAT[indx]
fixd <- !is.na(per$DODSDAT) & (per$DODSDAT < 1960010)
per$DODSDAT[fixd] <- per$DODSDAT[fixd] * 100 + 15
per$deathdate <- as.numeric(skum::tillTid(per$DODSDAT))
per$dodsKOMMUN <- socstyr$KOMMUN[indx]
per$dodsALDER <- socstyr$ALDER[indx]
per$ULORSAK <- socstyr$ULORSAK[indx]
```

# The FoBs

Through the *FoB*s we have snapshots of the population on *November 1* the years *1960(5)1990*, giving the dates (decimal form).

```{r days}
load("../../Data/fob.rda")
fob <- fob[fob$LINNEID %in% per$LINNEID, ]
ar <- seq(1960, 1990, by = 5)
days <- as.numeric(skum::tillTid(paste(ar, "1101", sep = "")))
days
```

We have to limit the data collection to the *KOMMUN*er that are contained
in our area, that is

```{r kommuner}
ourkom <- c(2406, 2414, 2417, 2418, 2480, 2482)
ourkom <- c(ourkom, 2411, 2412, 2413, 2416, 2415, 2461, 2407, 2402, 2408)
fob <- fob[fob$KOMMUN %in% ourkom, ]
```

Get `birthdate` from `per`.

```{r getbirthdateanddathdate}
indx <- match(fob$LINNEID, per$LINNEID)
fob$birthdate <- per$birthdate[indx]
fob$deathdate <- per$deathdate[indx]
```

Create individual Fobs (1960, 1965, etc):

```{r rawfob}
source("R/extractFob.R")
n <- length(ar)
fobbar <- vector(mode = "list", length = n)
names(fobbar) <- paste("fob", (ar %% 100), sep = "")
for (i in 1:(n-1)){
    fobbar[[i]] <- extractFob(ar[i], followUp = 5, data = fob)
}
fobbar[[n]] <- extractFob(ar[n], followUp = 11 + 61 / 365, data = fob)
fobdata <- do.call(rbind, fobbar)
fobdata <- fobdata[order(fobdata$LINNEID, fobdata$enter), ]
##fobdata <- age.window(fobdata, c(40, 110))
rownames(fobdata) <- 1:NROW(fobdata)
save(fobdata, file = "fobdata2.rda")
```

# Fob gets hisclass and the rest

WE continue as in "Mort2/CreateData2/fromFobLisa.Rmd":

```{r syskonfob}
    load("../../Data/socstyr.rda")
    indx <- match(fobdata$LINNEID, socstyr$LINNEID)
    fobdata$ULORSAK <- str_trim(as.character(socstyr$ULORSAK[indx]))
    source("R/putONssyk.R")
    ##putONssyk
    fobb <- putONssyk(fobdata)
    source("R/ssykTOhisclass.R")
    ##ssykTOhisclass
    fobb$hisclass <- ssykTOhisclass(fobb$ssyk)
    is.na(fobb$hisclass) <- with(fobb, is.na(hisclass) |
                                       (!is.na(hisclass) & hisclass == -1))
    source("R/addPer.R")
    ##fobb <- pushHis(fobb)
    ##fobb$hisclass <- fobb$nyhisclass
    ## Instead
    ##fobb <- group_by(fobb, LINNEID) %>% fill_("hisclass") # Later!!summary(lisa)
    fobb$id <- fobb$LINNEID
    ##fobb <- addPer(fobb, cuts = c(1961, 1976, 1986, 1996, 2002)) # Wait!
    load("../../Data/sex.rda")

    indx <- match(fobb$LINNEID, sex$LINNEID)
    fobb$KON <- sex$KON[indx]
    fobb$urban <- fobb$FORSAMLING %in% c(248001, 248002, 284004, 284005,
                                         248011, 248201, 248204)
    civst <- numeric(NROW(fobb))
    civst[fobb$CIVIL %in% c(0, 5)] <- 1 # Ogift
    civst[fobb$CIVIL %in% c(2, 7)] <- 2 # Gift, sambo
    civst[fobb$CIVIL %in% c(1, 3, 4, 6, 8, 9)] <- 3 # skild, änka
    fobb$civst <- factor(civst,
                         labels = c("unmarried", "married", "prev.married"))
    fobb$sex <- fobb$KON
    levels(fobb$sex) <- c("male", "female")
    fobb <- fobb[, c("id", "AR", "birthdate", "sex", "enter", "exit", "event",
                     "urban", "civst", "hisclass", "ULORSAK")]
```

# The Lisa data

The *lisa* data is slightly simpler because it is yearly with cutoff dates 12-31/01-01 each new-year. It starts off at 1987-01-01 (lisa1986) and ends at 2013-12-31 (lisa2013). So the time interval a *lisa* covers is *AR + 1*! But note that the *socstyr* data only cover deaths *before January 1, 2014, so *lisa2013* is in effect useless to us.

```{r lisadat}
load("../../Data/lisa.rda")
lisa <- lisa[(lisa$KOMMUN %in% ourkom) & (lisa$LINNEID %in% per$LINNEID), ]
indx <- match(lisa$LINNEID, per$LINNEID)

lisa$birthdate <- per$birthdate[indx]

```

So, in *lisa*, all enter in *AR + 1*, at the exact age of *(AR + 1 - birthdate)*, and exit at the exact age *enter + 1* if they do not die in *the year AR + 1*, otherwise at their death date, or at *deathdate - birthdate*.

```{r fixlisaintervals}
lisa$enter <- lisa$AR + 1 - lisa$birthdate
indx <- match(lisa$LINNEID, per$LINNEID) # To be safe ...
lisa$deathdate <- per$deathdate[indx]
lisa$event <- !is.na(lisa$deathdate) & lisa$deathdate > lisa$AR + 1 & lisa$deathdate <= lisa$AR + 2
lisa$exit <- lisa$enter + 1 # First tip.
lisa$exit[lisa$event] <- with(lisa, deathdate[event] - birthdate[event])
```

Now some sorting  and cleaning.

```{r cleanandsort}
##lisadata <- age.window(lisa, c(40, 110))
lisadata <- lisa[order(lisa$LINNEID, lisa$enter), ]
lisadata <- cal.window(lisadata, c(1987, 2014)) # 'Remove lisa2013'
rownames(lisadata) <- 1:NROW(lisadata)
##save(lisadata, file = "lisadata2.rda")
```

## Lisa gets hisclass and the rest

```{r lisadata}
    ##load("../../../Data/lisadata.rda")
    ##load("data4/lisadata.rda")
    ## Cause of death:
    load("../../Data/socstyr.rda")
    indx <- match(lisadata$LINNEID, socstyr$LINNEID)
    lisadata$ULORSAK <- str_trim(as.character(socstyr$ULORSAK[indx]))
    civst <- lisadata$CIV
    civst[is.na(civst)] <- 0
    civst[civst == 4] <- 3
    civst[civst == 5] <- 2
    civst[civst == 6] <- 3
    is.na(civst) <- civst == 0
    lisadata$civst <- civst
    rm(civst)
    ## Fix KON:
    load("../../Data/sex.rda")
    indx <- match(lisadata$LINNEID, sex$LINNEID)
    lisadata$sex <- sex$KON[indx]
    levels(lisadata$sex) <- c("male", "female")
    lisaold <- lisadata
    lisadata <- cal.window(lisadata, c(2002, 2014))

    ##lisadata <- lisadata[lisadata$AR > 1999.5 & !is.na(lisadata$SSYK3), ]

    ## This is new: Do not remove:
    ##lisadata <- lisadata[!is.na(lisadata$SSYK3), ] ##NOTE!

    ## Fix HISCLASS:
    source("R/ssyk3TOhisclass.R")
    ##ssyk3TOhisclass
    lisadata$hisclass <- ssyk3TOhisclass(lisadata$SSYK3)
    ## Push forward:
    source("R/pushHis.R")
    ## liss <- pushHis(lisadata) # Later!
    liss <- lisadata
    ##liss$hisclass <- liss$nyhisclass
    ##liss$nyhisclass <- NULL
    ## Now:
    ##liss <- group_by(lisadata, LINNEID) %>% fill_("hisclass") # Later!!
    liss$id <- liss$LINNEID
    ##lisaold$id <- lisaold$LINNEID

    ## lisaold <- addPer(lisaold, name = "period", cuts = seq(1986, 2014, by = 2)) ## Later! 
    ## liss <- addPer(liss, name = "period", cuts = seq(2002, 2014, by = 6)) # Later!!
    ##liss <- liss[!is.na(liss$hisclass), ]
    ##liss$hisclass <- cut(liss$hisclass, c(0, 2.5, 4.5, 5.5, 8),
      ##                   labels = 1:4)
    ##liss$hisclass <- as.numeric(as.character(liss$hisclass))
    ##liss <- liss[with(liss, !is.na(civst) &
      ##                      !is.na(hisclass)), ]
    ##lisaold <- lisaold[with(lisaold, !is.na(civst)), ]
    liss$civst <- factor(liss$civst,
                         labels = c("unmarried", "married", "prev.married"))
    ##lisaold$civst <- factor(lisaold$civst,
      ##                   labels = c("unmarried", "married", "prev.married"))
    liss$urban <- liss$FORSAMLING %in% c(248001, 248002, 284004,
                                         284005, 248011, 248201, 248204)
    ##lisaold$urban <- lisaold$FORSAMLING %in% c(248001, 248002, 284004,
      ##                                   284005, 248011, 248201, 248204)
    liss$id <- liss$LINNEID
    liss <- liss[, c("id", "AR", "birthdate", "sex", "enter", "exit", "event",
                     "urban", "civst", "hisclass", "ULORSAK")]
    ##liss$dataset <- paste("lisa", as.character(liss$AR), sep = "")
    save(liss, file = "liss2.rda")
    ##save(lisaold, file = "lisaold.rda")
```

# Combine Fob and Lisa

```{r combfoblisa}
source("R/fixCovar2.R")
foli <- rbind(fobb, liss)
foli <- foli[order(foli$id, foli$enter), ]
source("R/pushForward.R")
foli$hisclass <- unlist(tapply(foli$hisclass, foli$id, pushForward))
foli$hisclass[is.na(foli$hisclass)] <- 5
foli$hisclass <- factor(foli$hisclass, labels = c("elite", "middle", "worker", "farmer", "none"))
foli$hisclass40 <- fixCovar2(foli$id, foli$hisclass, foli$enter, foli$exit, age = 40, lookBack = 20)
foli$hisclass65 <- fixCovar2(foli$id, foli$hisclass, foli$enter, foli$exit, age = 65, lookBack = 45)
save(foli, file = "foli.rda")
```