### 2018-03-05 CreData/fromPoplink.Rmd

---
title: "From Poplink"
author: "Göran Broström"
date: "5 March, 2018"
output:
  pdf_document: default
  html_document: default
---

## Introduction

In this version we start at January 1, 1801, and we **do not** split the *poplink* data 
into two data sets (no causes of death, yet). Instead, `skume` refers to Beginning at January 1, 1801 and End at December 31, 1950.

```{r setup}
knitr::opts_chunk$set(cache=FALSE,echo=TRUE,message=TRUE,warning=TRUE,comment=NA)
library(xtable)
library(knitr)
library(eha)
library(skum)
##library(dplyr)
##library(tidyr)
##library(stringr)
```

## PopLink data

The POPLINK data cover 1801--1950. The resulting output consists of two tables,
**skel** and **skume** (I avoid the natural name **skum** in order to avoid a
mix-up with the **R** package by that name).
A covariate **period** is added and used for 
splitting the data into calendar time periods of length 10 years each.

### Get data

```{r getdata}
library(skum)

adult <- obs
indx <- match(adult$id, per$id)
adult$mid <- per$mid[indx]
adult$fid <- per$fid[indx]
adult$event <- !is.na(adult$sluttyp) & adult$sluttyp == 2
```

## Put on hisclass and civil status

```{r hisclass}
## Hisclasses:
ett <- with(adult, !is.na(socBranch) &
              (hisclass == 1))

tva <- with(adult, !is.na(socBranch) &
              (hisclass %in% c(2, 4)))

tre <- with(adult, !is.na(socBranch) &
              (hisclass %in% 5:7)) ## Original 'fyra' ---> 'tre'

fyra <- with(adult, !is.na(socBranch) &
               (hisclass == 3))

hisc <- rep(-1, NROW(adult))
hisc[ett] <- 1
hisc[tva] <- 2
hisc[tre] <- 3
hisc[fyra] <- 4
is.na(hisc) <- hisc == -1
adult$hisclass <- hisc
```

Continue ...

```{r urbanetc}
adult$urban <- with(adult, !is.na(ortnmn) & (ortnmn %in% c("UMEÅ", "SKELLEFTEÅ STAD")))
levels(adult$civst) <- c("unmarried", "married", "prev.married", "prev.married", "prev.married")
```

## Propagate

```{r propagate}
    source("R/pushForward.R")
    adult <- adult[order(adult$id, adult$enter), ]
    ##adult <- group_by(adult, id) %>% fill_("hisclass")
    adult$hisclass <- unlist(tapply(adult$hisclass, adult$id, pushForward))
    adult$hisclass[is.na(adult$hisclass)] <- 5
    adult$hisclass <- factor(adult$hisclass, 
                             labels = c("elite", "middle", "worker", "farmer", "none"))
```

## The Lexis cut

Our data covers the years 1801--1950 and the ages 40--99.

```{r lexis}
skume <- age.window(adult, c(30, 100))
skume <- cal.window(skume, c(1801, 1951))
```

### Trim skume

```{r skumefilt}
skume <- skume[, c("id", "birthdate", "sex", "enter", "exit", "event",
                    "urban", "civst",  "hisclass")]
skume <- skume[order(skume$id, skume$enter), ]
summary(skume)
```

## Save

```{r save}
save(skume, file = "../data/skume.rda")
```