### 2018-09-06 compare.Rmd
first commit

---
title: "Compare foli and new"
author: "Göran Broström"
date: "September 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(eha)
```

## Intro

I compare the old data file **foli.rmd** with the new version **per_60.rds**. The latter is slightly more strict with right censoring ("last seen").

## The foli file

```{r getfoli}
load("data/foli.rda")
summary(foli[, c("enter", "exit", "event", "sex", "birthdate")])
```

## The per_60 file

```{r getper}
per60 <- readRDS("../Long_Life/mydata/per_60.rds")
per60$sex <- factor(per60$KON, labels = c("male", "female"))
summary(per60[, c("enter", "exit", "event", "sex", "birthdate")])
```

## Comparison

We see that **foli** is age-cut 30-100, while **p60** only has a lower cut, at age 60. In order to make them comparable, we cut them into the largest common age interval, 6--100:

```{r cutcomp}
ageint <- c(60, 100)
per60 <- age.window(per60, ageint)[, c("LINNEID", "enter", "exit", "event", "sex", "birthdate")]
foli <- age.window(foli, ageint)[, c("id", "enter", "exit", "event", "sex", "birthdate")]
summary(foli)
summary(per60)
```